\section{Wands}

\label{f0}
\begin{table}[]
\sffamily
\caption{Table: Wands}
\begin{tabular}{lllll}
               &                 &                & \textbf{Spell} & \textbf{Caster}\\
\textbf{Minor} & \textbf{Medium} & \textbf{Major} & \textbf{Level} & \textbf{Level}\\
 01--05 & --- & --- & 0 & 1st\\
 06--60 & --- & --- & 1st & 1st\\
 61--100 & 01--60 & --- & 2nd & 3rd\\
 --- & 61--100 & 01--60 & 3rd & 5th\\
 --- & --- & 61--100 & 4th & 7th\\
\end{tabular}
\end{table}

				
A wand is a thin baton that contains a single spell of 4th level or lower. A wand has 50 charges when created---each charge allows the use of the wand's spell one time. A wand that runs out of charges is just a stick. The price of a wand is equal to the level of the spell \mbox{$\times$} the creator's caster level \mbox{$\times$} 750 gp. If the wand has a material component cost, it is added to the base price and cost to create once for each charge (50 \mbox{$\times$} material component cost). Table: Wands gives sample prices for wands created at the lowest possible caster level for each spellcasting class. Note that some spells appear at different levels for different casters. The level of such spells depends on the caster crafting the wand.

\begin{table}
 \sffamily
 \caption{Wand Costs}
 \begin{tabular}{lllll}
               & \textbf{Cleric} & \\
\textbf{Spell} & \textbf{Druid}  &                   &               & \textbf{Paladin} \\
\textbf{Level} & \textbf{Wizard} & \textbf{Sorcerer} & \textbf{Bard} & \textbf{Ranger} \\
0 & 375 gp & 375 gp & 375 gp & ---\\
1st & 750 gp & 750 gp & 750 gp & 750 gp\\
2nd & 4,500 gp & 6,000 gp & 6,000 gp & 6,000 gp\\
3rd & 11,250 gp & 13,500 gp & 15,750 gp & 15,750 gp\\
4th & 21,000 gp & 24,000 gp & 30,000 gp & 30,000 gp\\  
 \end{tabular}

\end{table}

% </div class="table">

				
\textbf{Physical Description}: A wand is 6 to 12 inches long, 1/4 inch thick, and weighs no more than 1 ounce. Most wands are wood, but some are bone, metal, or even crystal. A typical wand has AC 7, 5 hit points, hardness 5, and a break DC of 16.
				
\textbf{Activation}: Wands use the spell trigger activation method, so casting a spell from a wand is usually a standard action that doesn't provoke attacks of opportunity. (If the spell being cast has a longer casting time than 1 action, however, it takes that long to cast the spell from a wand.) To activate a wand, a character must hold it in hand (or whatever passes for a hand, for nonhumanoid creatures) and point it in the general direction of the target or area. A wand may be used while grappling or while swallowed whole.
				
\textbf{Special Qualities}: Roll d\%. A 01--30 result indicates that something (a design, inscription, or the like) provides some clue to the wand's function, and 31--100 indicates no special qualities.        	
