\section{Weapons}

\label{f0}		

\begin{table}[]
\sffamily
\caption{Table: Weapons}
\setlength{\tabcolsep}{1pt}
\begin{tabularx}{\linewidth}{lllXl}
               &                 &                &                 & \textbf{Bonus} \\
               &                 &                &                 & \textbf{Base} \\
\textbf{Minor} & \textbf{Medium} & \textbf{Major} & \textbf{Weapon} & \textbf{Price\(^{1}\)}\\
01--70 & 01--10 & --- & +1 & 2,000 gp\\
71--85 & 11--29 & --- & +2 & 8,000 gp\\
--- & 30--58 & 01--20 & +3 & 18,000 gp\\
--- & 59--62 & 21--38 & +4 & 32,000 gp\\
--- & --- & 39--49 & +5 & 50,000 gp\\
--- & --- & --- & +6\(^{2}\) & 72,000 gp\\
--- & --- & --- & +7\(^{2}\) & 98,000 gp\\
--- & --- & --- & +8\(^{2}\) & 128,000 gp\\
--- & --- & --- & +9\(^{2}\) & 162,000 gp\\
--- & --- & --- & +10\(^{2}\) & 200,000 gp\\
86--90 & 63--68 & 50--63 & Specific weapon\(^{3}\) & ---\\
91--100 & 69--100 & 64--100 & Special ability and roll again\(^{4}\) & ---\\
\end{tabularx}
\textsuperscript{1} For ammunition, this price is for 50 arrows, bolts, or bullets. \\
\textsuperscript{2} A weapon can't have an enhancement bonus higher than +5. Use these lines to determine price when special abilities are added in. \\
\textsuperscript{3} See Table: Specific Weapons. \\
\textsuperscript{4} See Table: Melee Weapon Special Abilities for melee weapons and Table: Ranged Weapon Special Abilities for ranged weapons.\\
\end{table}	

\begin{table}[]
\sffamily
\caption{Table: Melee Weapon Special Abilities}
\setlength{\tabcolsep}{1pt}
\begin{tabularx}{\linewidth}{lllXl}
               &                 &                &                  & \textbf{Base} \\
               &                 &                & \textbf{Special} & \textbf{Price} \\
\textbf{Minor} & \textbf{Medium} & \textbf{Major} & \textbf{Ability} & \textbf{Modifier\(^{1}\)}\\
01--10 & 01--06 & 01--03 & Bane & +1 bonus\\
11--17 & 07--12 & --- & Defending & +1 bonus\\
18--27 & 13--19 & 04--06 & Flaming & +1 bonus\\
28--37 & 20--26 & 07--09 & Frost & +1 bonus\\
38--47 & 27--33 & 10--12 & Shock & +1 bonus\\
48--56 & 34--38 & 13--15 & Ghost touch & +1 bonus\\
57--67 & 39--44 & --- & Keen, 2 & +1 bonus\\
68--71 & 45--48 & 16--19 & Ki Focus & +1 bonus\\
72--75 & 49--50 & --- & Merciful & +1 bonus\\
76--82 & 51--54 & 20--21 & Mighty cleaving & +1 bonus\\
83--87 & 55--59 & 22--24 & Spell storing & +1 bonus\\
88--91 & 60--63 & 25--28 & Throwing & +1 bonus\\
92--95 & 64--65 & 29--32 & Thundering & +1 bonus\\
96--99 & 66--69 & 33--36 & Vicious & +1 bonus\\
--- & 70--72 & 37--41 & Anarchic & +2 bonus\\
--- & 73--75 & 42--46 & Axiomatic & +2 bonus\\
--- & 76--78 & 47--49 & Disruption, 3 & +2 bonus\\
--- & 79--81 & 50--54 & Flaming burst & +2 bonus\\
--- & 82--84 & 55--59 & Icy burst & +2 bonus\\
--- & 85--87 & 60--64 & Holy & +2 bonus\\
--- & 88--90 & 65--69 & Shocking burst & +2 bonus\\
--- & 91--93 & 70--74 & Unholy & +2 bonus\\
--- & 94--95 & 75--78 & Wounding & +2 bonus\\
--- & --- & 79--83 & Speed & +3 bonus\\
--- & --- & 84--86 & Brilliant energy & +4 bonus\\
--- & --- & 87--88 & Dancing & +4 bonus\\
--- & --- & 89--90 & Vorpal, 2 & +5 bonus\\
100 & 96--100 & 91--100 & Roll again twice & ---\\
\end{tabularx}\\
\textsuperscript{1} Add to enhancement bonus on Table: Weapons to determine total market price. \\
\textsuperscript{2} Piercing or slashing weapons only (slashing only for vorpal). Reroll if randomly generated for a bludgeoning weapon. \\
\textsuperscript{3} Bludgeoning weapons only. Reroll if randomly generated for a piercing or slashing weapon.\\
\textsuperscript{4} Reroll if you get a duplicate special ability, an ability incompatible with an ability that you've already rolled, or if the extra ability puts you over the +10 limit. A weapon's enhancement bonus and special ability bonus equivalents can't total more than +10.\\
\end{table}

\begin{table}[]
\sffamily
\caption{Table: Melee Weapon Special Abilities}
\setlength{\tabcolsep}{1pt}
\begin{tabular}{lllll}
               &                 &                &                  & \textbf{Base} \\
               &                 &                & \textbf{Special} & \textbf{Price} \\
\textbf{Minor} & \textbf{Medium} & \textbf{Major} & \textbf{Ability} & \textbf{Modifier\(^{1}\)}\\
01--10 & 01--06 & 01--03 & Bane & +1 bonus\\
11--17 & 07--12 & --- & Defending & +1 bonus\\
18--27 & 13--19 & 04--06 & Flaming & +1 bonus\\
28--37 & 20--26 & 07--09 & Frost & +1 bonus\\
38--47 & 27--33 & 10--12 & Shock & +1 bonus\\
48--56 & 34--38 & 13--15 & Ghost touch & +1 bonus\\
57--67 & 39--44 & --- & Keen\(^{2}\) & +1 bonus\\
68--71 & 45--48 & 16--19 & Ki Focus & +1 bonus\\
72--75 & 49--50 & --- & Merciful & +1 bonus\\
76--82 & 51--54 & 20--21 & Mighty cleaving & +1 bonus\\
83--87 & 55--59 & 22--24 & Spell storing & +1 bonus\\
88--91 & 60--63 & 25--28 & Throwing & +1 bonus\\
92--95 & 64--65 & 29--32 & Thundering & +1 bonus\\
96--99 & 66--69 & 33--36 & Vicious & +1 bonus\\
--- & 70--72 & 37--41 & Anarchic & +2 bonus\\
--- & 73--75 & 42--46 & Axiomatic & +2 bonus\\
--- & 76--78 & 47--49 & Disruption\(^{3}\) & +2 bonus\\
--- & 79--81 & 50--54 & Flaming burst & +2 bonus\\
--- & 82--84 & 55--59 & Icy burst & +2 bonus\\
--- & 85--87 & 60--64 & Holy & +2 bonus\\
--- & 88--90 & 65--69 & Shocking burst & +2 bonus\\
--- & 91--93 & 70--74 & Unholy & +2 bonus\\
--- & 94--95 & 75--78 & Wounding & +2 bonus\\
--- & --- & 79--83 & Speed & +3 bonus\\
--- & --- & 84--86 & Brilliant energy & +4 bonus\\
--- & --- & 87--88 & Dancing & +4 bonus\\
--- & --- & 89--90 & Vorpal\(^{2}\) & +5 bonus\\
100 & 96--100 & 91--100 & Roll again twice\(^{4}\) & ---\\
\end{tabular}\\
\textsuperscript{1} Add to enhancement bonus on Table: Weapons to determine total market price. \\
\textsuperscript{2} Piercing or slashing weapons only (slashing only for vorpal). Reroll if randomly generated for a bludgeoning weapon. \\
\textsuperscript{3} Bludgeoning weapons only. Reroll if randomly generated for a piercing or slashing weapon. \\
\textsuperscript{4} Reroll if you get a duplicate special ability, an ability incompatible with an ability that you've already rolled, or if the extra ability puts you over the +10 limit. A weapon's enhancement bonus and special ability bonus equivalents can't total more than +10.\\
\end{table}
		
\begin{table}[]
\sffamily
\caption{Table: Ranged Weapon Special Abilities}
\setlength{\tabcolsep}{1pt}
\begin{tabular}{lllll}
               &                 &                &                  & \textbf{Base} \\
               &                 &                & \textbf{Special} & \textbf{Price} \\
\textbf{Minor} & \textbf{Medium} & \textbf{Major} & \textbf{Ability} & \textbf{Modifier\(^{1}\)}\\
01--12 & 01--08 & 01--04 & Bane,\(^{2}\) & +1 bonus\\
13--25 & 09--16 & 05--08 & Distance & +1 bonus\\
26--40 & 17--28 & 09--12 & Flaming,\(^{2}\) & +1 bonus\\
41--55 & 29--40 & 13--16 & Frost,\(^{2}\) & +1 bonus\\
56--60 & 41--42 & --- & Merciful,\(^{2}\) & +1 bonus\\
61--68 & 43--47 & 17--21 & Returning & +1 bonus\\
69--83 & 48--59 & 22--25 & Shock,\(^{2}\) & +1 bonus\\
84--93 & 60--64 & 26--27 & Seeking & +1 bonus\\
94--99 & 65--68 & 28--29 & Thundering,\(^{2}\) & +1 bonus\\
--- & 69--71 & 30--34 & Anarchic,\(^{2}\) & +2 bonus\\
--- & 72--74 & 35--39 & Axiomatic,\(^{2}\) & +2 bonus\\
--- & 75--79 & 40--49 & Flaming burst,\(^{2}\) & +2 bonus\\
--- & 80--82 & 50--54 & Holy,\(^{2}\) & +2 bonus\\
--- & 83--87 & 55--64 & Icy burst,\(^{2}\) & +2 bonus\\
--- & 88--92 & 65--74 & Shocking burst,\(^{2}\) & +2 bonus\\
--- & 93--95 & 75--79 & Unholy,\(^{2}\) & +2 bonus\\
--- & --- & 80--84 & Speed & +3 bonus\\
--- & --- & 85--90 & Brilliant energy & +4 bonus\\
100 & 96--100 & 91--100 & Roll again twice & ---\\
\end{tabular}
\textsuperscript{1} Add to enhancement bonus on Table: Weapons to determine total market price.\\
\textsuperscript{2} Bows, crossbows, and slings crafted with this ability bestow this power upon their ammunition.\\
\textsuperscript{3} Reroll if you get a duplicate special ability, an ability incompatible with an ability that you've already rolled, or if the extra ability puts you over the +10 limit. A weapon's enhancement bonus and special ability bonus equivalents can't total more than +10.\\
\end{table}
\begin{table}[]
\sffamily
\caption{Table: Specific Weapons}
\setlength{\tabcolsep}{1pt}
\begin{tabularx}{\linewidth}{lllXl}
               &                 &                & \textbf{Specific} & \textbf{Market} \\
\textbf{Minor} & \textbf{Medium} & \textbf{Major} & \textbf{Weapon} & \textbf{Price}\\
01--15 & --- & --- & Sleep arrow & 132 gp\\
16--25 & --- & --- & Screaming bolt & 267 gp\\
26--45 & --- & --- & Silver dagger, masterwork & 322 gp\\
46--65 & --- & --- & Cold iron longsword, masterwork & 330 gp\\
66--75 & 01--09 & --- & Javelin of lightning & 1,500 gp\\
76--80 & 10--15 & --- & Slaying arrow & 2,282 gp\\
81--90 & 16--24 & --- & Adamantine dagger & 3,002 gp\\
91--100 & 25--33 & --- & Adamantine battleaxe & 3,010 gp\\
--- & 34--37 & --- & Slaying arrow & 4,057 gp\\
--- & 38--40 & --- & Shatterspike & 4,315 gp\\
--- & 41--46 & --- & Dagger of venom & 8,302 gp\\
--- & 47--51 & --- & Trident of warning & 10,115 gp\\
--- & 52--57 & 01--04 & Assassin's dagger & 10,302 gp\\
--- & 58--62 & 05--07 & Shifter's sorrow & 12,780 gp\\
--- & 63--66 & 08--09 & Trident of fish command & 18,650 gp\\
--- & 67--74 & 10--13 & Flame tongue & 20,715 gp\\
--- & 75--79 & 14--17 & Luck blade & 22,060 gp\\
--- & 80--86 & 18--24 & Sword of subtlety & 22,310 gp\\
--- & 87--91 & 25--31 & Sword of the planes & 22,315 gp\\
--- & 92--95 & 32--37 & Nine lives stealer & 23,057 gp\\
--- & 96--98 & 38--42 & Oathbow & 25,600 gp\\
--- & 99--100 & 43--46 & Sword of life stealing & 25,715 gp\\
--- & --- & 47--51 & Mace of terror & 38,552 gp\\
--- & --- & 52--57 & Life-drinker & 40,320 gp\\
--- & --- & 58--62 & Sylvan scimitar & 47,315 gp\\
--- & --- & 63--67 & Rapier of puncturing & 50,320 gp\\
--- & --- & 68--73 & Sun blade & 50,335 gp\\
--- & --- & 74--79 & Frost brand & 54,475 gp\\
--- & --- & 80--84 & Dwarven thrower & 60,312 gp\\
--- & --- & 85--91 & Luck blade, wish & 62,360 gp\\
--- & --- & 92--95 & Mace of smiting & 75,312 gp\\
--- & --- & 96--97 & Luck blade & 102,660 gp\\
--- & --- & 98--99 & Holy avenger & 120,630 gp\\
--- & --- & 100 & Luck blade & 142,960 gp\\
\end{tabularx}
\end{table}
								
A magic weapon is enhanced to strike more truly and deliver more damage. Magic weapons have enhancement bonuses ranging from +1 to +5. They apply these bonuses to both attack and damage rolls when used in combat. All magic weapons are also masterwork weapons, but their masterwork bonuses on attack rolls do not stack with their enhancement bonuses on attack rolls.
				
Weapons come in two basic categories: melee and ranged. Some of the weapons listed as melee weapons can also be used as ranged weapons. In this case, their enhancement bonuses apply to both melee and ranged attacks.
				
Some magic weapons have special abilities. Special abilities count as additional bonuses for determining the market value of the item, but do not modify attack or damage bonuses (except where specifically noted). A single weapon cannot have a modified bonus (enhancement bonus plus special ability bonus equivalents, including those from character abilities and spells) higher than +10. A weapon with a special ability must also have at least a +1 enhancement bonus. Weapons cannot possess the same special ability more than once.
				
Weapons or ammunition can be made of an unusual material. Roll d\%: 01--95 indicates that the item is of a standard sort, and 96--100 indicates that it is made of a special material (see Equipment).
				
\textbf{Caster Level for Weapons}: The caster level of a weapon with a special ability is given in the item description. For an item with only an enhancement bonus and no other abilities, the caster level is three times the enhancement bonus. If an item has both an enhancement bonus and a special ability, the higher of the two caster level requirements must be met.
				
\textbf{Additional Damage Dice}: Some magic weapons deal additional dice of damage. Unlike other modifiers to damage, additional dice of damage are not multiplied when the attacker scores a critical hit.
				
\textbf{Ranged Weapons and Ammunition}: The enhancement bonus from a ranged weapon does not stack with the enhancement bonus from ammunition. Only the higher of the two enhancement bonuses applies.
				
Ammunition fired from a projectile weapon with an enhancement bonus of +1 or higher is treated as a magic weapon for the purpose of overcoming damage reduction. Similarly, ammunition fired from a projectile weapon with an alignment gains the alignment of that projectile weapon.
				
\textbf{Magic Ammunition and Breakage}: When a magic arrow, crossbow bolt, or sling bullet misses its target, there is a 50\% chance it breaks or is otherwise rendered useless. A magic arrow, bolt, or bullet that successfully hits a target is automatically destroyed after it delivers its damage.
				
\textbf{Light Generation}: Fully 30\% of magic weapons shed light equivalent to a \textit{light }spell. These glowing weapons are quite obviously magical. Such a weapon can't be concealed when drawn, nor can its light be shut off. Some of the specific weapons detailed below always or never glow, as defined in their descriptions.
				
\textbf{Hardness and Hit Points}: Each +1 of a magic weapon's enhancement bonus adds +2 to its hardness and +10 to its hit points.
				
\textbf{Activation}: Usually a character benefits from a magic weapon in the same way a character benefits from a mundane weapon---by wielding (attacking with) it. If a weapon has a special ability that the user needs to activate, then the user usually needs to utter a command word (a standard action). A character can activate the special abilities of 50 pieces of ammunition at the same time, assuming each piece has identical abilities.
				
\textbf{Magic Weapons and Critical Hits}: Some weapon special abilities and some specific weapons have an extra effect on a critical hit. This special effect also functions against creatures not normally subject to critical hits. On a successful critical roll, apply the special effect, but do not multiply the weapon's regular damage. 
				
\textbf{Weapons for Unusually Sized Creatures}: The cost of weapons for creatures who are neither Small nor Medium varies (see Equipment). The cost of the masterwork quality and any magical enhancement remains the same.
				
\textbf{Special Qualities}: Roll d\%. A 01--30 result indicates that the item sheds light, 31--45 indicates that something (a design, inscription, or the like) provides a clue to the weapon's function, and 46--100 indicates no special qualities. 
				
\subsection{Magic Weapon Special Ability Descriptions}

				
A weapon with a special ability must also have at least a +1 enhancement bonus.
				
\textbf{Anarchic}: An \textit{anarchic weapon} is infused with the power of chaos. It makes the weapon chaotically aligned and thus bypasses the corresponding damage reduction. It deals an extra 2d6 points of damage against all creatures of lawful alignment. It bestows one permanent negative level on any lawful creature attempting to wield it. The negative level remains as long as the weapon is in hand and disappears when the weapon is no longer wielded. This negative level cannot be overcome in any way (including \textit{restoration }spells) while the weapon is wielded.
				
Moderate evocation \mbox{$[$}chaotic\mbox{$]$}\textit{; }CL 7th; Craft Magic Arms and Armor, \textit{chaos hammer, }creator must be chaotic; Price +2 bonus.
				
\textbf{Axiomatic}: An \textit{axiomatic weapon} is infused with lawful power. It makes the weapon law-aligned and thus bypasses the corresponding damage reduction. It deals an extra 2d6 points of damage against chaotic creatures. It bestows one permanent negative level on any chaotic creature attempting to wield it. The negative level remains as long as the weapon is in hand and disappears when the weapon is no longer wielded. This negative level cannot be overcome in any way (including \textit{restoration }spells) while the weapon is wielded.
				
Moderate evocation \mbox{$[$}lawful\mbox{$]$}\textit{; }CL 7th; Craft Magic Arms and Armor, \textit{order's wrath, }creator must be lawful; Price +2 bonus.
				
\textbf{Bane}: A \textit{bane weapon} excels against certain foes. Against a designated foe, the weapon's enhancement bonus is +2 better than its actual bonus. It also deals an extra 2d6 points of damage against the foe. To randomly determine a weapon's designated foe, roll on the following table.

\begin{tabular}{ll}
\textbf{d\%} & \textbf{Designated Foe}      \\
01--05        & Aberrations                  \\
06--09        & Animals                      \\
10--16        & Constructs                   \\
17--22        & Dragons                      \\
23--27        & Fey                          \\
28--60        & Humanoids (pick one subtype) \\
61--65        & Magical beasts               \\
66--70        & Monstrous humanoids          \\
71--72        & Oozes                        \\
73--88        & Outsiders (pick one subtype) \\
89--90        & Plants                       \\
91--98        & Undead                       \\
99--100       & Vermin                      \\
\end{tabular}
				
Moderate conjuration; CL 8th; Craft Magic Arms and Armor, \textit{summon monster I}; Price +1 bonus.
				
\textbf{Brilliant Energy}: A \textit{brilliant energy weapon} has its significant portion transformed into light, although this does not modify the item's weight. It always gives off light as a torch (20-foot radius). A \textit{brilliant energy weapon} ignores nonliving matter. Armor and shield bonuses to AC (including any enhancement bonuses to that armor) do not count against it because the weapon passes through armor. (Dexterity, deflection, dodge, natural armor, and other such bonuses still apply.) A \textit{brilliant energy weapon} cannot harm undead, constructs, and objects. This property can only be applied to melee weapons, thrown weapons, and ammunition.
				
Strong transmutation\textit{; }CL 16th; Craft Magic Arms and Armor, \textit{gaseous form, continual flame; }Price +4 bonus.
				
\textbf{Dancing}: As a standard action, a \textit{dancing weapon }can be loosed to attack on its own. It fights for 4 rounds using the base attack bonus of the one who loosed it and then drops. While dancing, it cannot make attacks of opportunity, and the person who activated it is not considered armed with the weapon. The weapon is considered wielded or attended by the creature for all maneuvers and effects that target items. While dancing, the weapon shares the same space as the activating character and can attack adjacent foes (weapons with reach can attack opponents up to 10 feet away). The dancing weapon accompanies the person who activated it everywhere, whether she moves by physical or magical means. If the wielder who loosed it has an unoccupied hand, she can grasp it while it is attacking on its own as a free action; when so retrieved, the weapon can't dance (attack on its own) again for 4 rounds.
				
Strong transmutation\textit{; }CL 15th; Craft Magic Arms and Armor, \textit{animate objects; }Price +4 bonus.
				
\textbf{Defending}: A \textit{defending weapon} allows the wielder to transfer some or all of the weapon's enhancement bonus to his AC as a bonus that stacks with all others. As a free action, the wielder chooses how to allocate the weapon's enhancement bonus at the start of his turn before using the weapon, and the bonus to AC lasts until his next turn.
				
Moderate abjuration\textit{; }CL 8th; Craft Magic Arms and Armor, \textit{shield }or \textit{shield of faith; }Price +1 bonus.
				
\textbf{Disruption}: A \textit{disruption weapon} is the bane of all undead. Any undead creature struck in combat must succeed on a DC 14 Will save or be destroyed. A \textit{disruption weapon} must be a bludgeoning melee weapon.
				
Strong conjuration; CL 14th; Craft Magic Arms and Armor, \textit{heal; }Price +2 bonus.
				
\textbf{Distance}: This special ability can only be placed on a ranged weapon. A \textit{distance weapon} has double the range increment of other weapons of its kind.
				
Moderate divination; CL 6th; Craft Magic Arms and Armor, \textit{clairaudience/clairvoyance; }Price +1 bonus.
				
\textbf{Flaming}: Upon command, a \textit{flaming weapon} is sheathed in fire that deals an extra 1d6 points of fire damage on a successful hit. The fire does not harm the wielder. The effect remains until another command is given. 
				
Moderate evocation\textit{; }CL 10th; Craft Magic Arms and Armor and \textit{flame blade, flame strike, }or \textit{fireball; }Price +1 bonus.
				
\textbf{Flaming Burst}: A \textit{flaming burst weapon} functions as a \textit{flaming weapon} that also explodes with flame upon striking a successful critical hit. The fire does not harm the wielder. In addition to the extra fire damage from the \textit{flaming} ability (see above), a \textit{flaming burst weapon} deals an extra 1d10 points of fire damage on a successful critical hit. If the weapon's critical multiplier is \mbox{$\times$}3, add an extra 2d10 points of fire damage instead, and if the multiplier is \mbox{$\times$}4, add an extra 3d10 points of fire damage. 
				
Even if the \textit{flaming} ability is not active, the weapon still deals its extra fire damage on a successful critical hit.
				
Strong evocation\textit{; }CL 12th; Craft Magic Arms and Armor and \textit{flame blade, flame strike, }or \textit{fireball; }Price +2 bonus.
				
\textbf{Frost}: Upon command, a \textit{frost} \textit{weapon} is sheathed in icy cold that deals an extra 1d6 points of cold damage on a successful hit. The cold does not harm the wielder. The effect remains until another command is given.
				
Moderate evocation\textit{; }CL 8th; Craft Magic Arms and Armor, \textit{chill metal }or \textit{ice storm; }Price +1 bonus.
				
\textbf{Ghost Touch}: A \textit{ghost touch weapon} deals damage normally against incorporeal creatures, regardless of its bonus. An incorporeal creature's 50\% reduction in damage from corporeal sources does not apply to attacks made against it with \textit{ghost touch weapons}. The weapon can be picked up and moved by an incorporeal creature at any time. A manifesting ghost can wield the weapon against corporeal foes. Essentially, a \textit{ghost touch weapon} counts as both corporeal or incorporeal.
				
Moderate conjuration; CL 9th; Craft Magic Arms and Armor, \textit{plane shift; }Price +1 bonus.
				
\textbf{Holy}: A \textit{holy weapon} is imbued with holy power. This power makes the weapon good-aligned and thus bypasses the corresponding damage reduction. It deals an extra 2d6 points of damage against all creatures of evil alignment. It bestows one permanent negative level on any evil creature attempting to wield it. The negative level remains as long as the weapon is in hand and disappears when the weapon is no longer wielded. This negative level cannot be overcome in any way (including by \textit{restoration }spells) while the weapon is wielded.
				
Moderate evocation \mbox{$[$}good\mbox{$]$}; CL 7th; Craft Magic Arms and Armor, \textit{holy smite, }creator must be good; Price +2 bonus.
				
\textbf{Icy Burst}: An \textit{icy burst weapon} functions as a \textit{frost weapon} that also explodes with frost upon striking a successful critical hit. The frost does not harm the wielder. In addition to the extra damage from the \textit{frost} ability, an \textit{icy burst weapon} deals an extra 1d10 points of cold damage on a successful critical hit. If the weapon's critical multiplier is \mbox{$\times$}3, add an extra 2d10 points of cold damage instead, and if the multiplier is \mbox{$\times$}4, add an extra 3d10 points. 
				
Even if the \textit{frost} ability is not active, the weapon still deals its extra cold damage on a successful critical hit.
				
Moderate evocation\textit{; }CL 10th; Craft Magic Arms and Armor, \textit{chill metal }or \textit{ice storm; }Price +2 bonus.
				
\textbf{Keen}: This ability doubles the threat range of a weapon. Only piercing or slashing melee weapons can be \textit{keen}. If you roll this property randomly for an inappropriate weapon, reroll. This benefit doesn't stack with any other effect that expands the threat range of a weapon (such as the \textit{keen edge }spell or the Improved Critical feat).
				
Moderate transmutation\textit{; }CL 10th; Craft Magic Arms and Armor, \textit{keen edge; }Price +1 bonus.
				
\textbf{Ki Focus}: The magic weapon serves as a channel for the wielder's ki, allowing her to use her special \textit{ki }attacks through the weapon as if they were unarmed attacks. These attacks include the monk's \textit{ki }strike, quivering palm, and the Stunning Fist feat (including any condition that the monk can apply using this feat). Only melee weapons can have the \textit{ki focus} ability.
				
Moderate transmutation; CL 8th; Craft Magic Arms and Armor, creator must be a monk; Price +1 bonus.
				
\textbf{Merciful}: The weapon deals an extra 1d6 points of damage, and all damage it deals is nonlethal damage. On command, the weapon suppresses this ability until told to resume it (allowing it to deal lethal damage, but without any bonus damage from this ability).
				
Faint conjuration; CL 5th; Craft Magic Arms and Armor, \textit{cure light wounds}; Price +1 bonus.
				
\textbf{Mighty Cleaving}: A \textit{mighty cleaving weapon} allows a wielder using the Cleave feat to make one additional attack if the first attack hits, as long as the next foe is adjacent to the first and also within reach. This additional attack cannot be against the first foe. Only melee weapons can be \textit{mighty cleaving weapons}.
				
Moderate evocation\textit{; }CL 8th; Craft Magic Arms and Armor, \textit{divine power; }Price +1 bonus.
				
\textbf{Returning}: This special ability can only be placed on a weapon that can be thrown. A \textit{returning weapon} flies through the air back to the creature that threw it. It returns to the thrower just before the creature's next turn (and is therefore ready to use again in that turn). Catching a \textit{returning weapon} when it comes back is a free action. If the character can't catch it, or if the character has moved since throwing it, the weapon drops to the ground in the square from which it was thrown.
				
Moderate transmutation\textit{; }CL 7th; Craft Magic Arms and Armor, \textit{telekinesis; }Price +1 bonus.
				
\textbf{Seeking}: Only ranged weapons can have the \textit{seeking }ability. The weapon veers toward its target, negating any miss chances that would otherwise apply, such as from concealment. The wielder still has to aim the weapon at the right square. Arrows mistakenly shot into an empty space, for example, do not veer and hit invisible enemies, even if they are nearby.
				
Strong divination; CL 12th; Craft Magic Arms and Armor, \textit{true seeing}; Price +1 bonus.
				
\textbf{Shock}: Upon command, a \textit{shock weapon} is sheathed in crackling electricity that deals an extra 1d6 points of electricity damage on a successful hit. The electricity does not harm the wielder. The effect remains until another command is given. 
				
Moderate evocation\textit{; }CL 8th; Craft Magic Arms and Armor, \textit{call lightning }or \textit{lightning bolt; }Price +1 bonus.
				
\textbf{Shocking Burst}: A \textit{shocking burst weapon} functions as a \textit{shock weapon} that explodes with electricity upon striking a successful critical hit. The electricity does not harm the wielder. In addition to the extra electricity damage from the \textit{shock} ability, a \textit{shocking burst weapon} deals an extra 1d10 points of electricity damage on a successful critical hit. If the weapon's critical multiplier is \mbox{$\times$}3, add an extra 2d10 points of electricity damage instead, and if the multiplier is \mbox{$\times$}4, add an extra 3d10 points.
				
Even if the \textit{shock} ability is not active, the weapon still deals its extra electricity damage on a successful critical hit.
				
Moderate evocation\textit{; }CL 10th; Craft Magic Arms and Armor, \textit{call lightning }or \textit{lightning bolt; }Price +2 bonus.
				
\textbf{Speed}: When making a full-attack action, the wielder of a \textit{speed weapon} may make one extra attack with it. The attack uses the wielder's full base attack bonus, plus any modifiers appropriate to the situation. (This benefit is not cumulative with similar effects, such as a \textit{haste }spell.)
				
Moderate transmutation\textit{; }CL 7th; Craft Magic Arms and Armor, \textit{haste; }Price +3 bonus.
				
\textbf{Spell Storing}: A \textit{spell storing weapon} allows a spellcaster to store a single targeted spell of up to 3rd level in the weapon. (The spell must have a casting time of 1 standard action.) Anytime the weapon strikes a creature and the creature takes damage from it, the weapon can immediately cast the spell on that creature as a free action if the wielder desires. (This special ability is an exception to the general rule that casting a spell from an item takes at least as long as casting that spell normally.) Once the spell has been cast from the weapon, a spellcaster can cast any other targeted spell of up to 3rd level into it. The weapon magically imparts to the wielder the name of the spell currently stored within it. A randomly rolled \textit{spell storing weapon} has a 50\% chance to have a spell stored in it already.
				
Strong evocation (plus aura of stored spell)\textit{; }CL 12th; Craft Magic Arms and Armor, creator must be a caster of at least 12th level; Price +1 bonus.
				
\textbf{Throwing}: This ability can only be placed on a melee weapon. A melee weapon crafted with this ability gains a range increment of 10 feet and can be thrown by a wielder proficient in its normal use.
				
Faint transmutation\textit{; }CL 5th; Craft Magic Arms and Armor, \textit{magic stone; }Price +1 bonus.
				
\textbf{Thundering}: A \textit{thundering weapon} creates a cacophonous roar like thunder upon striking a successful critical hit. The sonic energy does not harm the wielder. A \textit{thundering weapon} deals an extra 1d8 points of sonic damage on a successful critical hit. If the weapon's critical multiplier is \mbox{$\times$}3, add an extra 2d8 points of sonic damage instead, and if the multiplier is \mbox{$\times$}4, add an extra 3d8 points of sonic damage. Subjects dealt critical hits by a \textit{thundering weapon} must make a DC 14 Fortitude save or be deafened permanently.
				
Faint necromancy\textit{; }CL 5th; Craft Magic Arms and Armor, \textit{blindness/deafness; }Price +1 bonus.
				
\textbf{Unholy}: An \textit{unholy weapon} is imbued with unholy power. This power makes the weapon evil-aligned and thus bypasses the corresponding damage reduction. It deals an extra 2d6 points of damage against all creatures of good alignment. It bestows one permanent negative level on any good creature attempting to wield it. The negative level remains as long as the weapon is in hand and disappears when the weapon is no longer wielded. This negative level cannot be overcome in any way (including \textit{restoration }spells) while the weapon is wielded.
				
Moderate evocation \mbox{$[$}evil\mbox{$]$}\textit{; }CL 7th; Craft Magic Arms and Armor, \textit{unholy blight, }creator must be evil; Price +2 bonus.
				
\textbf{Vicious}: When a \textit{vicious weapon} strikes an opponent, it creates a flash of disruptive energy that resonates between the opponent and the wielder. This energy deals an extra 2d6 points of damage to the opponent and 1d6 points of damage to the wielder. Only melee weapons can be \textit{vicious}.
				
Moderate necromancy; CL 9th; Craft Magic Arms and Armor, \textit{enervation}; Price +1 bonus.
				
\textbf{Vorpal}: This potent and feared ability allows the weapon to sever the heads of those it strikes. Upon a roll of natural 20 (followed by a successful roll to confirm the critical hit), the weapon severs the opponent's head (if it has one) from its body. Some creatures, such as many aberrations and all oozes, have no heads. Others, such as golems and undead creatures other than vampires, are not affected by the loss of their heads. Most other creatures, however, die when their heads are cut off. A \textit{vorpal weapon} must be a slashing melee weapon. If you roll this property randomly for an inappropriate weapon, reroll.
				
Strong necromancy and transmutation\textit{; }CL 18th; Craft Magic Arms and Armor\textit{, circle of death}, \textit{keen edge; }Price +5 bonus.
				
\textbf{Wounding}: A \textit{wounding weapon} deals 1 point of bleed damage when it hits a creature. Multiple hits from a wounding weapon increase the bleed damage. Bleeding creatures take the bleed damage at the start of their turns. Bleeding can be stopped by a DC 15 Heal check or through the application of any spell that cures hit point damage. A critical hit does not multiply the bleed damage. Creatures immune to critical hits are immune to the bleed damage dealt by this weapon.
				
Moderate evocation; CL 10th; Craft Magic Arms and Armor, \textit{bleed; }Price +2 bonus.
				
\subsection{Specific Weapons}

				
\textbf{Adamantine Battleaxe}
				
\textbf{Aura} no aura (nonmagical);\textbf{ CL }---
				
\textbf{Slot} none; \textbf{Price} 3,010 gp; \textbf{Weight} 6 lbs.
				
DESCRIPTION
				
This nonmagical axe is made out of adamantine. As a masterwork weapon, it has a +1 enhancement bonus on attack rolls. 
				
\textbf{Adamantine Dagger}
				
\textbf{Aura} no aura (nonmagical); \textbf{CL} ---
				
\textbf{Slot} none; \textbf{Price} 3,002 gp; \textbf{Weight} 1 lb.
				
DESCRIPTION
				
This nonmagical dagger is made out of adamantine. As a masterwork weapon, it has a +1 enhancement bonus on attack rolls. 
				
\textbf{Assassin's Dagger}
				
\textbf{Aura} moderate necromancy;\textbf{ CL }9th
				
\textbf{Slot} none; \textbf{Price} 10,302 gp; \textbf{Weight} 1 lb.
				
DESCRIPTION
				
This wicked-looking, curved \textit{+2 dagger} provides a +1 bonus to the DC of a Fortitude save forced by the death attack of an assassin. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{slay living}; \textbf{Cost }5,302 gp
				
\textbf{Dagger of Venom}
				
\textbf{Aura} faint necromancy; \textbf{CL} 5th
				
\textbf{Slot} none; \textbf{Price} 8,302 gp; \textbf{Weight} 1 lb.
				
DESCRIPTION
				
This black \textit{+1 dagger} has a serrated edge. It allows the wielder to use a \textit{poison} effect (as the spell, save DC 14) upon a creature struck by the blade once per day. The wielder can decide to use the power after he has struck. Doing so is a free action, but the \textit{poison} effect must be invoked in the same round that the dagger strikes. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{poison}; \textbf{Cost }4,302 gp
				
\textbf{Dwarven Thrower}
				
\textbf{Aura} moderate evocation; \textbf{CL} 10th
				
\textbf{Slot} none; \textbf{Price} 60,312 gp; \textbf{Weight} 5 lbs.
				
DESCRIPTION
				
This weapon functions as a \textit{+2 warhammer} in the hands of most users. Yet in the hands of a dwarf, the warhammer gains an additional +1 enhancement bonus (for a total enhancement bonus of +3) and gains the \textit{returning} special ability. It can be hurled with a 30-foot range increment. When hurled, a \textit{dwarven thrower} deals an extra 2d8 points of damage against creatures of the giant subtype or an extra 1d8 points of damage against any other target. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, creator must be a dwarf of at least 10th level; \textbf{Cost }30,312 gp
				
\textbf{Flame Tongue}
				
\textbf{Aura} strong evocation; \textbf{CL} 12th
				
\textbf{Slot} none; \textbf{Price} 20,715 gp; \textbf{Weight} 4 lbs.
				
DESCRIPTION
				
This is a \textit{+1 flaming burst longsword}. Once per day, the sword can blast forth a fiery ray at any target within 30 feet as a ranged touch attack. The ray deals 4d6 points of fire damage on a successful hit. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{scorching ray }and \textit{fireball}, \textit{flame blade}, or \textit{flame strike}; \textbf{Cost }10,515 gp
				
\textbf{Frost Brand}
				
\textbf{Aura} strong evocation; \textbf{CL} 14th
				
\textbf{Slot} none; \textbf{Price} 54,475 gp; \textbf{Weight} 8 lbs.
				
DESCRIPTION
				
This \textit{+3 frost greatsword} sheds light as a torch when the temperature drops below 0\^A\mbox{${}^\circ$} F. At such times it cannot be concealed when drawn, nor can its light be shut off. Its wielder is protected from fire; the sword absorbs the first 10 points of fire damage each round that the wielder would otherwise take.
				
A \textit{frost brand} extinguishes all nonmagical fires in a 20-foot radius. As a standard action, it can also dispel lasting fire spells, but not instantaneous effects. You must succeed on a dispel check (1d20 +14) against each spell to dispel it. The DC to dispel such spells is 11 + the caster level of the fire spell. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{ice storm}, \textit{dispel magic}, \textit{protection from energy}; \textbf{Cost }27,375 gp and 5 sp
				
\textbf{Holy Avenger}
				
\textbf{Aura} strong abjuration; \textbf{CL} 18th
				
\textbf{Slot} none; \textbf{Price} 120,630 gp; \textbf{Weight} 4 lbs.
				
DESCRIPTION
				
This \textit{+2 cold iron longsword} becomes a \textit{+5 holy cold iron longsword} in the hands of a paladin.
				
This sacred weapon provides spell resistance of 5 + the paladin's level to the wielder and anyone adjacent to her. It also enables the paladin to use \textit{greater dispel magic} (once per round as a standard action) at the class level of the paladin. Only the area dispel is possible, not the targeted dispel or counterspell versions of \textit{greater dispel magic}.
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{holy aura}, creator must be good; \textbf{Cost }60,630 gp
				
\textbf{Javelin of Lightning}
				
\textbf{Aura} faint evocation; \textbf{CL} 5th
				
\textbf{Slot} none; \textbf{Price} 1,500 gp; \textbf{Weight} 2 lbs.
				
DESCRIPTION
				
This javelin becomes a 5d6 \textit{lightning bolt} when thrown (Reflex DC 14 half). It is consumed in the attack. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{lightning bolt}; \textbf{Cost }750 gp
				
\textbf{Life-Drinker}
				
\textbf{Aura} strong necromancy;\textbf{ CL }13th
				
\textbf{Slot} none; \textbf{Price} 40,320 gp; \textbf{Weight} 12 lbs.
				
DESCRIPTION
				
This \textit{+1 greataxe} is favored by undead and constructs, who do not suffer its drawback. A life-drinker bestows two negative levels on its target whenever it deals damage, just as if its target had been struck by an undead creature. One day after being struck, subjects must make a DC 16 Fortitude save for each negative level or the negative levels become permanent.
				
Each time a \textit{life-drinker} deals damage to a foe, it also bestows one negative level on the wielder. Any negative levels gained by the wielder in this fashion lasts for 1 hour. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{enervation}; \textbf{Cost }20,320 gp
				
\textbf{Longsword, Cold Iron Masterwork }
				
\textbf{Aura} no aura (nonmagical); \textbf{CL} ---
				
\textbf{Slot} none; \textbf{Price} 330 gp; \textbf{Weight} 4 lbs.
				
DESCRIPTION
				
This nonmagical longsword is crafted out of cold iron. As a masterwork weapon, it has a +1 enhancement bonus on attack rolls. 
				
\textbf{Luck Blade}
				
\textbf{Aura} strong evocation; \textbf{CL} 17th
				
\textbf{Slot} none; \textbf{Price} 22,060 gp (0 \textit{wishes}), 62,360 gp (1 \textit{wish}), 102,660 gp (2 \textit{wishes}), 142,960 gp (3 \textit{wishes}); \textbf{Weight} 2 lbs.
				
DESCRIPTION
				
This \textit{+2 short sword} gives its possessor a +1 luck bonus on all saving throws. Its possessor also gains the power of good fortune, usable once per day. This extraordinary ability allows its possessor to reroll one roll that she just made, before the results are revealed. She must take the result of the reroll, even if it's worse than the original roll. In addition, a luck blade may contain up to three \textit{wishes} (when randomly rolled, a luck blade holds 1d4--1 \textit{wishes}, minimum 0). When the last \textit{wish} is used, the sword remains a \textit{+2 short sword}, still grants the +1 luck bonus, and still grants its reroll power. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{wish} or \textit{miracle}; \textbf{Cost }11,185 gp (0 wishes), 43,835 gp (1 wish), 76,485 gp (2 wishes), 109,135 gp (3 wishes).
				
\textbf{Mace of Smiting}
				
\textbf{Aura} moderate transmutation;\textbf{ CL }11th
				
\textbf{Slot} none; \textbf{Price} 75,312 gp; \textbf{Weight} 8 lbs.
				
DESCRIPTION
				
This \textit{+3 adamantine heavy mace} has a +5 enhancement bonus against constructs, and a successful critical hit dealt to a construct completely destroys the construct (no saving throw). A critical hit dealt to an outsider deals \mbox{$\times$}4 damage rather than \mbox{$\times$}2. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{disintegrate}; \textbf{Cost }39,312 gp
				
\textbf{Mace of Terror}
				
\textbf{Aura} strong necromancy;\textbf{ CL }13th
				
\textbf{Slot} none; \textbf{Price} 38,552 gp; \textbf{Weight} 8 lbs.
				
DESCRIPTION
				
This weapon usually appears to be a particularly frightening-looking iron or steel mace. On command, this \textit{+2 heavy mace} causes the wielder's clothes and appearance to transform into an illusion of darkest horror such that living creatures in a 30-foot cone become panicked as if by a \textit{fear} spell (Will DC 16 partial). Those who fail take a --2 morale penalty on saving throws, and they flee from the wielder. The wielder may use this ability up to three times per day. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{fear}; \textbf{Cost }19,432 gp
				
\textbf{Nine Lives Stealer}
				
\textbf{Aura} strong necromancy \mbox{$[$}evil\mbox{$]$}; \textbf{CL} 13th
				
\textbf{Slot} none; \textbf{Price} 23,057 gp; \textbf{Weight} 4 lbs.
				
DESCRIPTION
				
This longsword always performs as a \textit{+2 longsword}, but it also has the power to draw the life force from an opponent. It can do this nine times before the ability is lost. At that point, the sword becomes a simple \textit{+2 longsword} (with a faint evil aura). A critical hit must be dealt for the sword's death-dealing ability to function, and this weapon has no effect on creatures not subject to critical hits. The victim is entitled to a DC 20 Fortitude save to avoid death. If the save is successful, the sword's death-dealing ability does not function, no use of the ability is expended, and normal critical damage is determined. This sword is evil, and any good character attempting to wield it gains two negative levels. These negative levels remain as long as the sword is in hand and disappear when the sword is no longer wielded. These negative levels never result in actual level loss, but they cannot be overcome in any way (including \textit{restoration} spells) while the sword is wielded. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{finger of death}; \textbf{Cost }11,528 gp 5 sp.
				
\textbf{Oathbow}
				
\textbf{Aura} strong evocation;\textbf{ CL }15th
				
\textbf{Slot} none; \textbf{Price} 25,600 gp; \textbf{Weight} 3 lbs.
				
DESCRIPTION
				
Of elven make, this white \textit{+2 composite longbow} (+2 Str bonus) whispers, \texttt{{}"{}}Swift defeat to my enemies\texttt{{}"{}} in Elven when nocked and pulled. Once per day, if the archer swears aloud to slay her target (a free action), the bow's whisper becomes the shout \texttt{{}"{}}Death to those who have wronged me!\texttt{{}"{}} Against such a sworn enemy, the bow has a +5 enhancement bonus, and arrows launched from it deal an additional 2d6 points of damage (and \mbox{$\times$}4 on a critical hit instead of the normal \mbox{$\times$}3). After an enemy has been sworn, the bow is treated as only a masterwork weapon against all foes other than the sworn enemy, and the archer takes a --1 penalty on attack rolls with any weapon other than the \textit{oathbow}. These bonuses and penalties last for 7 days or until the sworn enemy is slain or destroyed by the wielder of the \textit{oathbow}, whichever comes first.
				
The \textit{oathbow} may only have one sworn enemy at a time. Once the wielder swears to slay a target, he cannot make a new oath until he has slain that target or 7 days have passed. Even if the wielder slays the sworn enemy on the same day that he makes the oath, he cannot activate the \textit{oathbow's} special power again until 24 hours have passed from the time he made the oath. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, creator must be an elf; \textbf{Cost }13,100 gp
				
\textbf{Rapier of Puncturing}
				
\textbf{Aura} strong necromancy;\textbf{ CL }13th
				
\textbf{Slot} none; \textbf{Price} 50,320 gp; \textbf{Weight} 2 lbs.
				
DESCRIPTION
				
Three times per day, this \textit{+2 wounding rapier} allows the wielder to make a touch attack with the weapon that deals 1d6 points of Constitution damage by draining blood. Creatures immune to critical hits are immune to the Constitution damage dealt by this weapon. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{harm}; \textbf{Cost }25,320 gp
				
\textbf{Screaming Bolt}
				
\textbf{Aura} faint enchantment;\textbf{ CL }5th
				
\textbf{Slot} none; \textbf{Price} 267 gp; \textbf{Weight} 1/10 lb.
				
DESCRIPTION
				
These \textit{+2 bolts} scream when fired, forcing all enemies of the wielder within 20 feet of the path of the bolt to succeed on a DC 14 Will save or become shaken. This is a mind-affecting fear effect. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{doom}; \textbf{Cost }137 gp
				
\textbf{Shatterspike}
				
\textbf{Aura} strong evocation;\textbf{ CL }13th
				
\textbf{Slot} none; \textbf{Price} 4,315 gp; \textbf{Weight} 4 lbs.
				
DESCRIPTION
				
This intimidating weapon appears to be a longsword with multiple hooks, barbs, and serrations along the blade, excellent for catching and sundering a foe's weapon. Wielders without the Improved Sunder feat use a \textit{shatterspike} as a \textit{+1 longsword} only. Wielders with the Improved Sunder feat instead use \textit{shatterspike} as a \textit{+4 longsword} when attempting to sunder an opponent's weapon. \textit{Shatterspike} can damage weapons with an enhancement bonus of +4 or lower.
				
Construction
				
\textbf{Requirements} Str 13, Craft Magic Arms and Armor, Improved Sunder, Power Attack, \textit{shatter}; \textbf{Cost }2,315 gp
				
\textbf{Shifter's Sorrow}
				
\textbf{Aura} strong transmutation;\textbf{ CL }15th
				
\textbf{Slot} none; \textbf{Price} 12,780 gp; \textbf{Weight} 10 lbs.
				
DESCRIPTION
				
This \textit{+1/+1 two-bladed sword} has blades of alchemical silver. The weapon deals an extra 2d6 points of damage against any creature with the shapechanger subtype. When a shapechanger or a creature in an alternate form (such as a druid using wild shape) is struck by the weapon, it must make a DC 15 Will save or return to its natural form. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{baleful polymorph}; \textbf{Cost }6,780 gp
				
\textbf{Silver Dagger, Masterwork}
				
\textbf{Aura} no aura (nonmagical); \textbf{CL} ---
				
\textbf{Slot} none; \textbf{Price} 322 gp; \textbf{Weight} 1 lb.
				
DESCRIPTION
				
As a masterwork weapon, this alchemical silver dagger has a +1 enhancement bonus on attack rolls (but not to damage rolls).
				
\textbf{Slaying Arrow}
				
\textbf{Aura} strong necromancy;\textbf{ CL }13th
				
\textbf{Slot} none; \textbf{Price} 2,282 gp (\textit{slaying arrow}) or 4,057 gp (\textit{greater slaying arrow}); \textbf{Weight} 1/10 lb.
				
DESCRIPTION
				
This \textit{+1 arrow} is keyed to a particular type or subtype of creature. If it strikes such a creature, the target must make a DC 20 Fortitude save or take 50 points of damage. Note that even creatures normally exempt from Fortitude saves (undead and constructs) are subject to this attack. When keyed to a living creature, this is a death effect (and thus \textit{death ward} protects a target). To determine the type or subtype of creature the arrow is keyed to, roll on the table below.
				
A \textit{greater slaying arrow} functions just like a normal \textit{slaying arrow}, but the DC to avoid the death effect is 23 and the arrow deals 100 points of damage if the saving throw is failed. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{finger of death} (\textit{slaying arrow}) or heightened\textit{ finger of death} (\textit{greater slaying arrow}); \textbf{Cost }1,144 gp 5 sp (\textit{slaying arrow}) or 2,032 gp (\textit{greater slaying arrow})
% <thead href="../spells/fingerOfDeath.html#finger-of-death">
\begin{tabular}{ll}
\textbf{d\%} & \textbf{Designated Type or Subtype} \\
01--05        & Aberrations                         \\
06--09        & Animals                             \\
10--16        & Constructs                          \\
17--27        & Dragons                             \\
28--32        & Fey                                 \\
33           & Humanoids, aquatic                  \\
34--35        & Humanoids, dwarf                    \\
36--37        & Humanoids, elf                      \\
38--44        & Humanoids, giant                    \\
45           & Humanoids, gnoll                    \\
46           & Humanoids, gnome                    \\
47--49        & Humanoids, goblinoid                \\
50           & Humanoids, halfling                 \\
51--54        & Humanoids, human                    \\
55--57        & Humanoids, reptilian                \\
58--60        & Humanoids, orc                      \\
61--65        & Magical beasts                      \\
66--70        & Monstrous humanoids                 \\
71--72        & Oozes                               \\
73           & Outsiders, air                      \\
74--76        & Outsiders, chaotic                  \\
77           & Outsiders, earth                    \\
78--80        & Outsiders, evil                     \\
81           & Outsiders, fire                     \\
82--84        & Outsiders, good                     \\
85--87        & Outsiders, lawful                   \\
88           & Outsiders, water                    \\
89--90        & Plants                              \\
91--98        & Undead                              \\
99--100       & Vermin                             
\end{tabular}

\textbf{Sleep Arrow}
				
\textbf{Aura} faint enchantment;\textbf{ CL }5th
				
\textbf{Slot} none; \textbf{Price} 132 gp; \textbf{Weight} 1/10 lb.
				
DESCRIPTION
				
This \textit{+1 arrow} is painted white and has white fletching. If it strikes a foe so that it would normally deal damage, it instead bursts into magical energy that deals nonlethal damage (in the same amount as would lethal damage) and forces the target to make a DC 11 Will save or fall asleep. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{sleep}; \textbf{Cost }70 gp
				
\textbf{Sun Blade}
				
\textbf{Aura} moderate evocation;\textbf{ CL }10th
				
\textbf{Slot} none; \textbf{Price} 50,335 gp; \textbf{Weight} 2 lbs.
				
DESCRIPTION
				
This sword is the size of a bastard sword. However, a \textit{sun blade} is wielded as if it were a short sword with respect to weight and ease of use. In other words, the weapon appears to all viewers to be a bastard sword, and deals bastard sword damage, but the wielder feels and reacts as if the weapon were a short sword. Any individual able to use either a bastard sword or a short sword with proficiency is proficient in the use of a \textit{sun blade}. Likewise, Weapon Focus and Weapon Specialization in short sword and bastard sword apply equally, but the benefits of those feats do not stack.
				
In normal combat, the glowing golden blade of the weapon is equal to a \textit{+2 bastard sword}. Against evil creatures, its enhancement bonus is +4. Against Negative Energy Plane creatures or undead creatures, the sword deals double damage (and \mbox{$\times$}3 on a critical hit instead of the usual \mbox{$\times$}2).
				
The blade also has a special sunlight power. Once per day, the wielder can swing the blade vigorously above his head while speaking a command word. The \textit{sun blade} then sheds a bright yellow radiance that acts like bright light and affects creatures susceptible to light as if it were natural sunlight. The radiance begins shining in a 10-foot radius around the sword wielder and extends outward at 5 feet per round for 10 rounds thereafter, to create a globe of light with a 60-foot radius. When the wielder stops swinging, the radiance fades to a dim glow that persists for another minute before disappearing entirely. All \textit{sun blades} are of good alignment, and any evil creature attempting to wield one gains one negative level. The negative level remains as long as the sword is in hand and disappears when the sword is no longer wielded. This negative level cannot be overcome in any way (including by \textit{restoration} spells) while the sword is wielded. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{daylight}, creator must be good; \textbf{Cost }25,335 gp
				
\textbf{Sword of Life Stealing}
				
\textbf{Aura} strong necromancy;\textbf{ CL }17th
				
\textbf{Slot} none; \textbf{Price} 25,715 gp; \textbf{Weight} 4 lbs.
				
DESCRIPTION
				
This black iron \textit{+2 longsword} bestows a negative level when it deals a critical hit. The sword wielder gains 1d6 temporary hit points each time a negative level is bestowed on another. These temporary hit points last for 24 hours. One day after being struck, subjects must make a DC 16 Fortitude save for each negative level gained or they become permanent. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{enervation}; \textbf{Cost }13,015 gp
				
\textbf{Sword of the Planes}
				
\textbf{Aura} strong evocation;\textbf{ CL }15th
				
\textbf{Slot} none; \textbf{Price} 22,315 gp; \textbf{Weight} 4 lbs.
				
DESCRIPTION
				
This longsword has an enhancement bonus of +1 on the Material Plane, but on any Elemental Plane its enhancement bonus increases to +2. The +2 enhancement bonus also applies whenever the weapon is used against creatures native to the Elemental Plane. It operates as a \textit{+3 longsword} on the Astral Plane and the Ethereal Plane, or when used against opponents native to either of those planes. On any other plane, or against any outsider, it functions as a \textit{+4 longsword}. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{plane shift}; \textbf{Cost }11,315 gp
				
\textbf{Sword of Subtlety}
				
\textbf{Aura} moderate illusion;\textbf{ CL }7th
				
\textbf{Slot} none; \textbf{Price} 22,310 gp; \textbf{Weight} 2 lbs.
				
DESCRIPTION
				
A \textit{+1 short sword} with a thin, dull gray blade, this weapon provides a +4 bonus on its wielder's attack and damage rolls when he is making a sneak attack with it. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{blur}; \textbf{Cost }11,310 gp
				
\textbf{Sylvan Scimitar}
				
\textbf{Aura} moderate evocation;\textbf{ CL }11th
				
\textbf{Slot} none; \textbf{Price} 47,315 gp; \textbf{Weight} 4 lbs.
				
DESCRIPTION
				
This \textit{+3 scimitar}, when used outdoors in a temperate climate, grants its wielder the use of the Cleave feat and deals an extra 1d6 points of damage. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{divine power} or creator must be a 7th-level druid; \textbf{Cost }23,815 gp
				
\textbf{Trident of Fish Command}
				
\textbf{Aura} moderate enchantment;\textbf{ CL }7th
				
\textbf{Slot} none; \textbf{Price} 18,650 gp; \textbf{Weight} 4 lbs.
				
DESCRIPTION
				
The magical properties of this \textit{+1 trident} with a 6-foot-long haft enable its wielder to charm up to 14 HD of aquatic animals as per the spell \textit{charm animals} (Will DC 16 negates, animals get a +5 bonus if currently under attack by the wielder or his allies), no two of which can be more than 30 feet apart. The wielder can use this effect up to three times per day. The wielder can communicate with the animals as if using a \textit{speak with animals} spell. Animals making their saving throws are free of control, but they will not approach within 10 feet of the trident. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{charm animals, speak with animals}; \textbf{Cost }9,482 gp and 5 sp
				
\textbf{Trident of Warning}
				
\textbf{Aura} moderate divination;\textbf{ CL }7th
				
\textbf{Slot} none; \textbf{Price} 10,115 gp; \textbf{Weight} 4 lbs.
				
DESCRIPTION
				
A weapon of this type enables its wielder to determine the location, depth, kind, and number of aquatic predators within 680 feet. A \textit{trident of warning} must be grasped and pointed in order for the character using it to gain such information, and it requires 1 round to scan a hemisphere with a radius of 680 feet. The weapon is otherwise a \textit{+2 trident}. 
				
Construction
				
\textbf{Requirements} Craft Magic Arms and Armor, \textit{locate creature}; \textbf{Cost }5,215 gp	
