\section{Duelist}

\label{f0}				
Duelists represent the pinnacle of elegant swordplay. They move with a grace unmatched by most foes, parrying blows and countering attacks with swift thrusts of their blades. They may wear armor, but generally eschew such bulky protection as their grace allows them to dodge their opponents with ease. While others flounder on treacherous terrain, duelists charge nimbly across the battlefield, leaping and tumbling into the fray. They thrive in melee, where their skill with the blade allows them to make sudden attacks against clumsy foes and to cripple opponents with particularly well-placed thrusts of the blade.
				
The path to the duelist is natural for rogues and bards, as those classes do not rely on armor for defense, although nearly as many duelists come from the ranks of fighters and rangers. They are often found in those regions that possess elaborate rules and etiquette for battle. 
				
\textbf{Role}: The abilities of duelists complement those rogues or bards who wish to accentuate their fighting prowess but, because of their lack of heavy armor, are afraid to leap into combat. Duelists fight in the forefront alongside fighters, barbarians, and other melee combatants, deftly avoiding the blades of their opponents while expertly targeting their vulnerabilities.
				
\textbf{Alignment}: Duelists can be of any alignment, although since most hail from backgrounds as rogues or bards, they tend to eschew lawful behavior. Lawful duelists are not unheard of, however, and such duelists often adhere to a strict code of honor, refusing to attack unarmed or obviously inferior opponents.
				
\textbf{Hit Die}: d10.
				
\subsection{Requirements}

				
To qualify to become a duelist, a character must fulfill all the following criteria.
				
\textbf{Base Attack Bonus}: +6.
				
\textbf{Skills}: Acrobatics 2 ranks, Perform 2 ranks.
				
\textbf{Feats}: Dodge, Mobility, Weapon Finesse.
				
\subsection{Class Skills}

				
The duelist's class skills (and the key ability for each skill) are Acrobatics (Dex), Bluff (Cha), Escape Artist (Dex), Perception (Wis), Perform (Cha), and Sense Motive (Wis).
				
\textbf{ Skill Ranks at Each Level}: 4 + Int modifier.
% <div class="table">

% <

\begin{table*}[]  % In the CRB, the picture overlaps the table
\sffamily
\caption{Table: Duelist}
\begin{tabular}{llllll}
               &\textbf{Base} \\
               &\textbf{Attack} & \textbf{Fort} & \textbf{Ref}  & \textbf{Will}\\
\textbf{Level} &\textbf{Base}   & \textbf{Save} & \textbf{Save} & \textbf{Save} & \textbf{Special}\\
1st & +1 & +0 & +1 & +0 & Canny defense, precise strike\\
2nd & +2 & +1 & +1 & +1 & Improved reaction +2, parry\\
3rd & +3 & +1 & +2 & +1 & Enhanced mobility\\
4th & +4 & +1 & +2 & +1 & Combat Reflexes, grace\\
5th & +5 & +2 & +3 & +2 & Riposte\\
6th & +6 & +2 & +3 & +2 & Acrobatic charge\\
7th & +7 & +2 & +4 & +2 & Elaborate defense\\
8th & +8 & +3 & +4 & +3 & Improved reaction +4\\
9th & +9 & +3 & +5 & +3 & Deflect Arrows, no retreat\\
10th & +10 & +3 & +5 & +3 & Crippling critical\\
\end{tabular}
\end{table*}
				
\subsection{Class Features}

				
All of the following are class features of the duelist prestige class.
				
\textbf{Weapon and Armor Proficiency}: The duelist is proficient with all simple and martial weapons. Duelists are proficient with light armor but not with shields.
				
\textbf{Canny Defense (Ex)}: When wearing light or no armor and not using a shield, a duelist adds 1 point of Intelligence bonus (if any) per duelist class level as a dodge bonus to her Armor Class while wielding a melee weapon. If a duelist is caught flat-footed or otherwise denied her Dexterity bonus, she also loses this bonus.
				
\textbf{Precise Strike (Ex)}: A duelist gains the ability to strike precisely with a light or one-handed piercing weapon, adding her duelist level to her damage roll.
				
When making a precise strike, a duelist cannot attack with a weapon in her other hand or use a shield. A duelist's precise strike only works against living creatures with discernible anatomies. Any creature that is immune to critical hits is also immune to a precise strike, and any item or ability that protects a creature from critical hits also protects a creature from a precise strike.
				
\textbf{Improved Reaction (Ex)}: At 2nd level, a duelist gains a +2 bonus on initiative checks. At 8th level, the bonus increases to +4. This bonus stacks with the benefit provided by the Improved Initiative feat.
				
\textbf{Parry (Ex)}: At 2nd level, a duelist learns to parry the attacks of other creatures, causing them to miss. Whenever the duelist takes a full attack action with a light or one-handed piercing weapon, she can elect not to take one of her attacks. At any time before her next turn, she can attempt to parry an attack against her or an adjacent ally as an immediate action. To parry the attack, the duelist makes an attack roll, using the same bonuses as the attack she chose to forego during her previous action. If her attack roll is greater than the roll of the attacking creature, the attack automatically misses. For each size category that the attacking creature is larger than the duelist, the duelist takes a --4 penalty on her attack roll. The duelist also takes a --4 penalty when attempting to parry an attack made against an adjacent ally. The duelist must declare the use of this ability after the attack is announced, but before the roll is made.
				
\textbf{Enhanced Mobility (Ex)}: Starting at 3rd level, when wearing light or no armor and not using a shield, a duelist gains an additional +4 bonus to AC against attacks of opportunity caused when she moves out of a threatened square.
				
\textbf{Combat Reflexes}: At 4th level, a duelist gains the benefit of the Combat Reflexes feat when using a light or one-handed piercing weapon.
				
\textbf{Grace (Ex)}: At 4th level, a duelist gains an additional +2 competence bonus on all Reflex saving throws. This ability functions for a duelist only when she is wearing light or no armor and not using a shield.
				
\textbf{Riposte (Ex)}: Starting at 5th level, a duelist can make an attack of opportunity against any creature whose attack she successfully parries, so long as the creature she is attacking is within reach.
				
\textbf{Acrobatic Charge (Ex)}: At 6th level, a duelist gains the ability to charge in situations where others cannot. She may charge over difficult terrain that normally slows movement. Depending on the circumstance, she may still need to make appropriate checks to successfully move over the terrain.
				
\textbf{Elaborate Defense (Ex)}: At 7th level and higher, if a duelist chooses to fight defensively or use total defense in melee combat, she gains an additional +1 dodge bonus to AC for every 3 levels of duelist she has attained.
				
\textbf{Deflect Arrows}: At 9th level, a duelist gains the benefit of the Deflect Arrows feat when using a light or one-handed piercing weapon. The duelist does not need a free hand to use this feat.
				
\textbf{No Retreat (Ex)}: At 9th level, enemies adjacent to the duelist that take a withdraw action provoke an attack of opportunity from the duelist.
				
\textbf{Crippling Critical (Ex)}: When you confirm a critical hit using a light or one-handed piercing weapon, you can apply one of the following penalties in addition to the damage dealt: reduce all of the target's speeds by 10 feet (minimum 5 feet), 1d4 points of Strength or Dexterity damage, --4 penalty on all saving throws, --4 penalty to Armor Class, or 2d6 points of bleed damage. These penalties last for 1 minute, except for ability damage, which must be healed normally, and bleed damage, which continues until the target receives magic healing or a DC 15 Heal skill check.
        	
