\section{Eldritch Knight}

\label{f0}				
Fearsome warriors and spellcasters, eldritch knights are rare among magic-users in their ability to wade into battle alongside fighters, barbarians, and other martial classes. Those who must face eldritch knights in combat fear them greatly, for their versatility on the battlefield is tremendous; against heavily armed and armored opponents they may level crippling spells, while opposing spellcasters meet their ends on an eldritch knight's blade.
				
Because the road to becoming an eldritch knight requires both martial prowess and arcane power, eldritch knights almost always begin their paths as multiclassed characters, such as fighter/wizards or ranger/sorcerers. They may be found wherever studies of the arcane are as prevalent as martial training.
				
\textbf{Role}: Eldritch knights blend the abilities of fighting classes and spellcasters, hurling magic at the enemy one moment and hewing down their opponents with steel the next. They are just as comfortable fighting in the thick of combat as they are casting spells at foes while remaining safely behind their compatriots. Their versatility makes them valuable allies when the nature of an upcoming battle is unclear.
				
\textbf{Alignment}: The road to becoming an eldritch knight is as varied as the paths leading to apprenticeship under a wizard or a career as a soldier, and eldritch knights can therefore be of any alignment. Maintaining a balance between studies of arcane lore and martial techniques requires great discipline, however, and for that reason many favor lawful alignments. 
				
\textbf{Hit Die}: d10.
				
\subsection{Requirements}

				
To qualify to become an eldritch knight, a character must fulfill all the following criteria.
				
\textbf{Weapon Proficiency}: Must be proficient with all martial weapons.
				
\textbf{Spells}: Able to cast 3rd-level arcane spells.
				
\subsection{Class Skills}

				
The eldritch knight's class skills (and the key ability for each skill) are Climb (Str), Knowledge (arcana) (Int), Knowledge (nobility) (Int), Linguistics (Int), Ride (Dex), Sense Motive (Wis), Spellcraft (Int), and Swim (Str). 
				
\textbf{ Skill Ranks at Each Level}: 2 + Int modifier.
% <div class="table">

\begin{table}[]
\sffamily
\caption{Table: Eldritch Knight}
\begin{tabular}{lllllll}
      & \textbf{Base} & & & & & \\ 
      & \textbf{Attack} & \textbf{Fort} & \textbf{Ref} & \textbf{Will} & & \\
\textbf{Level} & \textbf{Bonus }& \textbf{Save }&\textbf{ Save }& \textbf{Save }& \textbf{Special }& \textbf{Spells per Day}\\
1st & +1 & +1 & +0 & +0 & Bonus combat feat, diverse training & -\\
2nd & +2 & +1 & +1 & +1 & - & +1 level of existing arcane spellcasting class\\
3rd & +3 & +2 & +1 & +1 & - & +1 level of existing arcane spellcasting class\\
4th & +4 & +2 & +1 & +1 & - & +1 level of existing arcane spellcasting class\\
5th & +5 & +3 & +2 & +2 & Bonus combat feat & +1 level of existing arcane spellcasting class\\
6th & +6 & +3 & +2 & +2 & - & +1 level of existing arcane spellcasting class\\
7th & +7 & +4 & +2 & +2 & - & +1 level of existing arcane spellcasting class\\
8th & +8 & +4 & +3 & +3 & - & +1 level of existing arcane spellcasting class\\
9th & +9 & +5 & +3 & +3 & Bonus combat feat & +1 level of existing arcane spellcasting class\\
10th & +10 & +5 & +3 & +3 & Spell critical & +1 level of existing arcane spellcasting class\\
\end{tabular}
\end{table}
				
\subsection{Class Features}

				
All of the following are features of the eldritch knight prestige class.
				
\textbf{Weapon and Armor Proficiency}: Eldritch knights gain no proficiency with any weapon or armor.
				
\textbf{Bonus Feat}: At 1st level, an eldritch knight may choose a bonus combat feat. This is in addition to the feats that a character of any class normally gets from advancing levels. The character must still meet any prerequisites for these bonus feats. An eldritch knight gains an additional combat feat at 5th and 9th level.
				
\textbf{Diverse Training}: An eldritch knight adds his level to any levels of fighter he might have for the purpose of meeting the prerequisites for feats (if he has no fighter levels, treat his eldritch knight levels as levels of fighter). He also adds his level to any levels in an arcane spellcasting class for the purpose of meeting the prerequisites for feats.
				
\textbf{Spells per Day}: At the indicated levels, an eldritch knight gains new spells per day as if he had also gained a level in an arcane spellcasting class he belonged to before adding the prestige class. He does not, however, gain any other benefit a character of that class would have gained, except for additional spells per day, spells known (if he is a spontaneous spellcaster), and an increased effective level of spellcasting. If a character had more than one arcane spellcasting class before becoming an eldritch knight, he must decide to which class he adds the new level for purposes of determining spells per day.
				
\textbf{Spell Critical (Su)}: At 10th level, whenever an eldritch knight successfully confirms a critical hit, he can cast a spell as a swift action. The spell must include the target of the attack as one of its targets or in its area of effect. Casting this spell does not provoke an attack of opportunity. The caster must still meet all of the spell's components and must roll for arcane spell failure if necessary.
        	
