\chapter{Feats}

\label{f0}        				
Some abilities are not tied to your race, class, or skill---things like particularly quick reflexes that allow you to react to danger more swiftly, the ability to craft magic items, the training to deliver powerful strikes with melee weapons, or the knack for deflecting arrows fired at you. These abilities are represented as feats. While some feats are more useful to certain types of characters than others, and many of them have special prerequisites that must be met before they are selected, as a general rule feats represent abilities outside of the normal scope of your character's race and class. Many of them alter or enhance class abilities or soften class restrictions, while others might apply bonuses to your statistics or grant you the ability to take actions otherwise prohibited to you. By selecting feats, you can customize and adapt your character to be uniquely yours. 
				
\section{Prerequisites}

				
Some feats have prerequisites. Your character must have the indicated ability score, class feature, feat, skill, base attack bonus, or other quality designated in order to select or use that feat. A character can gain a feat at the same level at which he gains the prerequisite.
				
A character can't use a feat if he loses a prerequisite, but he does not lose the feat itself. If, at a later time, he regains the lost prerequisite, he immediately regains full use of the feat that prerequisite enables.
				
\section{Types of Feats}

				
Some feats are general, meaning that no special rules govern them as a group. Others are item creation feats, which allow characters to create magic items of all sorts. A metamagic feat lets a spellcaster prepare and cast a spell with greater effect, albeit as if the spell were of a higher spell level than it actually is.
				
\subsection{Combat Feats}

				
Any feat designated as a combat feat can be selected as a fighter's bonus feat. This designation does not restrict characters of other classes from selecting these feats, assuming that they meet the prerequisites.
				
\subsection{Critical Feats}

				
Critical feats modify the effects of a critical hit by inflicting an additional condition on the victim of the critical hit. Characters without the Critical Mastery feat can only apply the effects of one critical feat to an individual critical hit. Characters with multiple critical feats can decide which feat to apply after the critical hit has been confirmed.
								
\subsection{Item Creation Feats}

				
An item creation feat lets a character create a magic item of a certain type. Regardless of the type of item each involves, the various item creation feats all have certain features in common.
				
\textbf{Raw Materials Cost}: The cost of creating a magic item equals half the base price of the item.
				
Using an item creation feat also requires access to a laboratory or magical workshop, special tools, and so on. A character generally has access to what he needs unless unusual circumstances apply.
				
\textbf{Time}: The time to create a magic item depends on the feat and the cost of the item.
				
\textbf{Item Cost}: Brew Potion, Craft Staff, Craft Wand, and Scribe Scroll create items that directly reproduce spell effects, and the power of these items depends on their caster level---that is, a spell from such an item has the power it would have if cast by a spellcaster of that level. The price of these items (and thus the cost of the raw materials) also depends on the caster level. The caster level must be low enough that the spellcaster creating the item can cast the spell at that level. To find the final price in each case, multiply the caster level by the spell level, then multiply the result by a constant, as shown below:

\textbf{\textit{Scrolls:} Base price = spell level \mbox{$\times$} caster level \mbox{$\times$} 25 gp.}\newline
\textbf{\textit{Potions:} Base price = spell level \mbox{$\times$} caster level \mbox{$\times$} 50 gp.}\newline
\textbf{\textit{Wands:} Base price = spell level \mbox{$\times$} caster level \mbox{$\times$} 750 gp.}\newline
\textbf{\textit{Staves:} The price for staves is calculated using more complex formulas (see Magic Items).}
				
A 0-level spell is considered to have a spell level of 1/2 for the purpose of this calculation.
				
\textbf{Extra Costs}: Any potion, scroll, or wand that stores a spell with a costly material component also carries a commensurate cost. For potions and scrolls, the creator must expend the material component cost when creating the item. For a wand, the creator must expend 50 units of the material component. Some magic items similarly incur extra costs in material components, as noted in their descriptions.
				
\textbf{Skill Check}: Successfully creating a magic item requires a Spellcraft check with a DC equal to 5 + the item's caster level. Alternatively, you can use an associated Craft or Profession skill to attempt this check instead, depending upon the item being crafted. See Magic Item Creation for more details on which Craft and Profession checks may be substituted in this manner. The DC of this check can increase if the crafter is rushed or does not meet all of the prerequisites. A failed check ruins the materials used, while a check that fails by 5 or more results in a cursed item. See Magic Items for more details.
								
\subsection{Metamagic Feats}

				
As a spellcaster's knowledge of magic grows, he can learn to cast spells in ways slightly different from the norm. Preparing and casting a spell in such a way is harder than normal but, thanks to metamagic feats, is at least possible. Spells modified by a metamagic feat use a spell slot higher than normal. This does not change the level of the spell, so the DC for saving throws against it does not go up. Metamagic feats do not affect spell-like abilities.
				
\textbf{Wizards and Divine Spellcasters}: Wizards and divine spellcasters must prepare their spells in advance. During preparation, the character chooses which spells to prepare with metamagic feats (and thus which ones take up higher-level spell slots than normal).
				
\textbf{Sorcerers and Bards}: Sorcerers and bards choose spells as they cast them. They can choose when they cast their spells whether to apply their metamagic feats to improve them. As with other spellcasters, the improved spell uses up a higher-level spell slot. Because the sorcerer or bard has not prepared the spell in a metamagic form in advance, he must apply the metamagic feat on the spot. Therefore, such a character must also take more time to cast a metamagic spell (one enhanced by a metamagic feat) than he does to cast a regular spell. If the spell's normal casting time is a standard action, casting a metamagic version is a full-round action for a sorcerer or bard. (This isn't the same as a 1-round casting time.) The only exception is for spells modified by the Quicken Spell metamagic feat, which can be cast as normal using the feat.
				
For a spell with a longer casting time, it takes an extra full-round action to cast the spell.
				
\textbf{Spontaneous Casting and Metamagic Feats}: A cleric spontaneously casting a cure or inflict spell, or a druid spontaneously casting a \textit{summon nature's ally }spell, can cast a metamagic version of it instead. Extra time is also required in this case. Casting a standard action metamagic spell spontaneously is a full-round action, and a spell with a longer casting time takes an extra full-round action to cast. The only exception is for spells modified by the Quicken Spell feat, which can be cast as a swift action.
				
\textbf{Effects of Metamagic Feats on a Spell}: In all ways, a metamagic spell operates at its original spell level, even though it is prepared and cast using a higher-level spell slot. Saving throw modifications are not changed unless stated otherwise in the feat description.
				
The modifications made by these feats only apply to spells cast directly by the feat user. A spellcaster can't use a metamagic feat to alter a spell being cast from a wand, scroll, or other device.
				
Metamagic feats that eliminate components of a spell don't eliminate the attack of opportunity provoked by casting a spell while threatened. Casting a spell modified by Quicken Spell does not provoke an attack of opportunity.
				
Metamagic feats cannot be used with all spells. See the specific feat descriptions for the spells that a particular feat can't modify.
				
\textbf{Multiple Metamagic Feats on a Spell}: A spellcaster can apply multiple metamagic feats to a single spell. Changes to its level are cumulative. You can't apply the same metamagic feat more than once to a single spell.
				
\textbf{Magic Items and Metamagic Spells}: With the right item creation feat, you can store a metamagic version of a spell in a scroll, potion, or wand. Level limits for potions and wands apply to the spell's higher spell level (after the application of the metamagic feat). A character doesn't need the metamagic feat to activate an item storing a metamagic version of a spell.
				
\textbf{Counterspelling Metamagic Spells}: Whether or not a spell has been enhanced by a metamagic feat does not affect its vulnerability to counterspelling or its ability to counterspell another spell (see Magic).
				
\section{Feat Descriptions}

				
Feats are summarized on Table: Feats below. Note that the prerequisites and benefits of the feats on this table are abbreviated for ease of reference. See the feats description for full details.
				
The following format is used for all feat descriptions.
				
\textbf{Feat Name}: The feat's name also indicates what subcategory, if any, the feat belongs to, and is followed by a basic description of what the feat does.
				
\textbf{Prerequisite}: A minimum ability score, another feat or feats, a minimum base attack bonus, a minimum number of ranks in one or more skills, or anything else required in order to take the feat. This entry is absent if a feat has no prerequisite. A feat may have more than one prerequisite.
				
\textbf{Benefit}: What the feat enables the character (\texttt{{}"{}}you\texttt{{}"{}} in the feat description) to do. If a character has the same feat more than once, its benefits do not stack unless indicated otherwise in the description.
				
\textbf{Normal:} What a character who does not have this feat is limited to or restricted from doing. If not having the feat causes no particular drawback, this entry is absent.
				
\textbf{Special:} Additional unusual facts about the feat.

\begin{table*}[]
\sffamily
\setlength{\tabcolsep}{1pt}
\caption{Table: Feats}
\begin{tabularx}{\linewidth}{lXl}
Feats & Prerequisites & Benefits\\
Acrobatic & --- & +2 bonus on Acrobatics and Fly checks \\
Agile Maneuvers & --- & Use your Dex bonus when calculating your CMB \\
Alertness & --- & +2 bonus on Perception and Sense Motive checks\\
Alignment Channel & Channel energy class feature & Channel energy can heal or harm outsiders\\
Animal Affinity & --- & +2 bonus on Handle Animal and Ride checks\\
Arcane Armor Training & Armor Proficiency, Light, caster level 3rd & Reduce your arcane spell failure chance by 10\%\\
\quad Arcane Armor Mastery & Arcane Armor Training, Armor Proficiency, Medium, caster level 7th & Reduce your arcane spell failure chance by 20\%\\
Arcane Strike & Ability to cast arcane spells & +1 damage and weapons are considered magic\\
Armor Proficiency, Light & --- & No penalties on attack rolls while wearing light armor\\
\quad Armor Proficiency, Medium & Armor Proficiency, Light & No penalties on attack rolls while wearing medium armor\\
\quad \quad Armor Proficiency, Heavy & Armor Proficiency, Medium & No penalties on attack rolls while wearing heavy armor\\
Athletic & --- & +2 bonus on Climb and Swim checks\\
Augment Summoning & Spell Focus & Summoned creatures gain +4 Str and Con\\
Blind-Fight & --- & Reroll miss chances for concealment\\
Catch Off-Guard & --- & No penalties for improvised melee weapons\\
Channel Smite & Channel energy class feature & Channel energy through your attack\\
Combat Casting & --- & +4 bonus on concentration checks for casting defensively\\
Combat Expertise & Int 13 & Trade attack bonus for AC bonus\\
\quad Improved Disarm & Combat Expertise & +2 bonus on disarm attempts, no attack of opportunity\\
\quad \quad Greater Disarm & Improved Disarm, base attack bonus +6 & Disarmed weapons are knocked away from your enemy\\
\quad Improved Feint & Combat Expertise & Feint as a move action\\
\quad \quad Greater Feint & Improved Feint, base attack bonus +6 & Enemies you feint lose their Dex bonus for 1 round \\
\quad Improved Trip & Combat Expertise & +2 bonus on trip attempts, no attack of opportunity\\
\quad \quad Greater Trip & Improved Trip, base attack bonus +6 & Enemies you provoke attacks of opportunity\\
\quad Whirlwind Attack & Dex, Combat Expertise, Spring Attack, base attack bonus +4 & Make one melee attack against all foes within reach\\
Combat Reflexes & --- & Make additional attacks of opportunity\\
\quad Stand Still & Combat Reflexes & Stop enemies from moving past you\\
Command Undead & Channel negative energy class feature & Channel energy can be used to control undead\\
Critical Focus & Base attack bonus +9 & +4 bonus on attack rolls made to confirm critical hits\\
\quad Bleeding Critical & Critical Focus, base attack bonus +11 & Whenever you score a critical hit, the target takes 2d6 bleed \\
\quad Blinding Critical & Critical Focus, base attack bonus +15 & Whenever you score a critical hit, the target is blinded\\
\quad Critical Mastery & Any two Critical feats, 14th level fighter  & Apply two effects to your critical hits\\
\quad Deafening Critical & Critical Focus, base attack bonus +13 & Whenever you score a critical hit, the target is deafened\\
\quad Sickening Critical & Critical Focus, base attack bonus +11 & Whenever you score a critical hit, the target is sickened\\
\quad Staggering Critical & Critical Focus, base attack bonus +13 & Whenever you score a critical hit, the target is staggered\\
\quad Stunning Critical & Staggering Critical, base attack bonus +17 & Whenever you score a critical hit, the target is stunned\\
\quad Tiring Critical & Critical Focus, base attack bonus +13 & Whenever you score a critical hit, the target is fatigued\\
\quad \quad Exhausting Critical & Tiring Critical, base attack bonus +15 & Whenever you score a critical hit, the target is exhausted\\\
Deadly Aim & Dex 13, base attack bonus +1 & Trade ranged attack bonus for damage\\
Deceitful & --- & +2 bonus on Bluff and Disguise checks\\
Defensive Combat Training & --- & Use your total Hit Dice as your base attack bonus for \\
Deft Hands & --- & +2 bonus on Disable Device and Sleight of Hand checks\\
Disruptive & 6th-level fighter & Increases the DC to cast spells adjacent to you\\
\quad Spellbreaker & Disruptive, 10th-level fighter & Enemies provoke attacks if their spells fail\\
Dodge & Dex 13 & +1 dodge bonus to AC\\
\quad Mobility & Dodge & +4 AC against attacks of opportunity from movement\\
\quad \quad Spring Attack & Mobility, base attack bonus +4 & Move before and after melee attack\\
\quad Wind Stance & Dex 15, Dodge, base attack bonus +6 & Gain 20\% concealment if you move\\
\quad \quad Lightning Stance & Dex 17, Wind Stance, base attack bonus +11 & Gain 50\% concealment if you move\\
\end{tabularx}
\end{table*}
\begin{table*}
\sffamily
  \setlength{\tabcolsep}{1pt}
\begin{tabularx}{\linewidth}{lXl}
Elemental Channel & Channel energy class feature & Channel energy can harm or heal elementals\\
Endurance & --- & +4 bonus on checks to avoid nonlethal damage\\
\quad Diehard & Endurance & Automatically stabilize and remain conscious below 0 hp\\
Eschew Materials & --- & Cast spells without material components\\
Exotic Weapon Proficiency & Base attack bonus +1 & No penalty on attacks made with one exotic weapon\\
Extra Channel & Channel energy class feature & Channel energy two additional times per day\\
Extra \textit{Ki} & \textit{Ki} pool class feature & Increase your ki pool by 2 points\\
Extra Lay On Hands & Lay on hands class feature & Use lay on hands two additional times per day\\
Extra Mercy & Mercy class feature & Your lay on hands benefits from one additional mercy\\
Extra Performance & Bardic performance class feature & Use bardic performance for 6 additional rounds per day\\
Extra Rage & Rage class feature & Use rage for 6 additional rounds per day\\
Fleet & --- & Your base speed increases by 5 feet\\
Great Fortitude & --- & +2 on Fortitude saves\\
\quad Improved Great Fortitude & Great Fortitude & Once per day, you may reroll a Fortitude Save\\
Improved Channel & Channel energy class feature & +2 bonus on channel energy DC\\
Improved Counterspell & --- & Counterspell with spell of the same school\\
Improved Critical & Proficiency with weapon, base attack bonus +8 & Double the threat range of one weapon\\
Improved Familiar & Ability to acquire a familiar, see feat & Gain a more powerful familiar\\
Improved Initiative & --- & +4 bonus on initiative checks\\
Improved Unarmed Strike & --- & Always considered armed\\
\quad Deflect Arrows & Dex 13,  Improved Unarmed Strike & Avoid one ranged attack per round\\
\quad \quad Snatch Arrows & Dex 15, Deflect Arrows & Catch one ranged attack per round\\
\quad Improved Grapple & Dex 13, Improved Unarmed Strike & +2 bonus on grapple attempts, no attack of opportunity\\
\quad \quad Greater Grapple & Improved Grapple, base attack bonus +6 & Maintain your grapple as a move action\\
\quad Scorpion Style & Improved Unarmed Strike & Reduce target's speed to 5 ft.\\
\quad \quad Gorgon's Fist & Scorpion Style, base attack bonus +6 & Stagger a foe whose speed is reduced\\
\quad \quad \quad Medusa's Wrath & Gorgon's Fist, base attack bonus +11 & Make 2 extra attacks against a hindered foe\\
\quad Stunning Fist & Dex 13, Wis 13, Improved Unarmed Strike, base attack bonus +8 & Stun opponent with an unarmed strike\\
Improvised Weapon Mastery & Catch Off-Guard, Throw Anything, base attack bonus +8 & Make an improvised weapon deadly\\
Intimidating Prowess & --- & Add Str to Intimidate in addition to Cha \\
Iron Will & --- & +2 bonus on Will saves\\
\quad Improved Iron Will & Iron Will & Once per day, you may reroll a Will save\\
Leadership & Character level 7th & Gain a cohort and followers\\
Lightning Reflexes & --- & +2 bonus on Reflex saves\\
\quad Improved Lightning Reflexes & Lightning Reflexes & Once per day, you may reroll a Reflex save\\
Lunge & Base attack bonus +6 & Take a –2 penalty to your AC to attack with reach\\
Magical Aptitude & --- & +2 bonus on Spellcraft and Use Magic Device checks\\
Martial Weapon Proficiency & --- & No penalty on attacks made with one martial weapon\\
Master Craftsman & 5 ranks in any Craft or Profession skill & You can craft magic items without being a spellcaster\\
Mounted Combat & Ride & Avoid attacks on mount with Ride check\\
\quad Mounted Archery & Mounted Combat & Halve the penalty for ranged attacks while mounted\\
\quad Ride-By Attack & Mounted Combat & Move before and after a charge attack while mounted\\
\quad \quad Spirited Charge & Ride-By Attack & Double damage on a mounted charge\\
\quad Trample & Mounted Combat & Overrun targets while mounted\\
\quad Unseat & Improved Bull Rush, Mounted Combat & Knock opponents from their mounts\\
Natural Spell & Wis & Cast spells while using wild shape\\
Nimble Moves & Dex & Ignore 5 feet of difficult terrain when you move\\
\quad Acrobatic Steps & Dex, Nimble Moves & Ignore 20 feet of difficult terrain when you move\\
Persuasive & --- &  +2 bonus on Diplomacy and Intimidate checks\\
\end{tabularx}
\end{table*}
\begin{table*}
  \sffamily
  \setlength{\tabcolsep}{1pt}
\begin{tabularx}{\linewidth}{lXl}
Point-Blank Shot & --- & +1 attack and damage on targets within 30 feet\\
\quad Far Shot & Point-Blank Shot & Decrease ranged penalties by half\\
\quad Precise Shot & Point-Blank Shot & No penalty for shooting into melee\\
\quad \quad Improved PreciseShot & Dex, Precise Shot & No cover or concealment chance on ranged attacks\\
\quad  Pinpoint Targeting & Improved Precise Shot & No armor or shield bonus on one ranged attack\\
\quad Shot on the Run & Dex, Mobility, Point-Blank Shot & Make ranged attack at any point during movement\\
\quad Rapid Shot & Dex, Point-Blank Shot & Make one extra ranged attack\\
\quad \quad Manyshot & Dex, Rapid Shot & Shoot two arrows simultaneously\\
Power Attack & Str & Trade melee attack bonus for damage\\
\quad Cleave & Power Attack & Make an additional attack if the first one hits\\
\quad \quad Great Cleave & Cleave & Make an additional attack after each attack hits\\
\quad Improved Bull Rush & Power Attack & +2 bonus on Bull Rush attempts, no attack of opportunity\\
\quad \quad Greater Bull Rush & Improved Bull Rush & Enemies you bull rush provoke attacks of opportunity\\
\quad Improved Overrun & Power Attack & +2 bonus on overrun attempts, no attack of opportunity\\
\quad \quad Greater Overrun & Improved Overrun & Enemies you overrun provoke attacks of opportunity \\
\quad Improved Sunder & Power Attack & +2 bonus on sunder attempts, no attack of opportunity\\
\quad \quad Greater Sunder & Improved Sunder & Damage from sunder attempts transfers to your enemy\\
Quick Draw & Base attack bonus +1 & Draw weapon as a free action\\
Rapid Reload & Weapon proficiency (crossbow) & Reload crossbow quickly\\
Run & --- & Run at 5 times your normal speed\\
Selective Channeling & Cha 13, channel energy class feature & Choose whom to affect with channel energy\\
Self-Sufficient & --- & +2 bonus on Heal and Survival checks\\
Shield Proficiency & --- & No penalties on attack rolls when using a shield\\
\quad Improved Shield Bash & Shield Proficiency & Keep your shield bonus when shield bashing\\
\quad \quad Shield Slam & Improved Shield Bash, Two-Weapon Fighting & Free \\
\quad \quad \quad Shield Master & Shield Slam & No two-weapon penalties when attacking with a shield\\
\quad Shield Focus & Shield Proficiency & Gain a +1 bonus to your AC when using a shield\\
\quad \quad Greater Shield Focus & Shield Focus & Gain a +1 bonus to your AC when using a shield\\
\quad Tower Shield Proficiency & Shield Proficiency & No penalties on attack rolls when using a tower shield\\
Simple Weapon Proficiency & --- & No penalty on attacks made with simple weapons\\
Skill Focus & --- & +3 bonus on one skill (+6 at 10 ranks)\\
Spell Focus & --- & +1 bonus on save DCs for one school\\
\quad Greater Spell Focus & Spell Focus & +1 bonus on save DCs for one school\\
Spell Mastery & 1st-level Wizard & Prepare some spells without a spellbook\\
Spell Penetration & --- & +2 bonus on level checks to beat spell resistance\\
\quad Greater Spell Penetration & Spell Penetration & +2 bonus on level checks to beat spell resistance\\
Stealthy & --- & +2 bonus on Escape Artist and Stealth checks\\
Step Up & Base attack bonus +1 & Take a 5-foot step as an immediate action\\
Strike Back & Base attack bonus +11 & Attack foes that strike you while using reach\\
Throw Anything & --- & No penalties for improvised ranged weapons\\
Toughness & --- & +3 hit points, +1 per Hit Die beyond 3\\
Turn Undead & Channel positive energy class feature & Channel energy can be used to make undead flee\\
Two-Weapon Fighting & Dex & Reduce two-weapon fighting penalties\\
\quad Double Slice & Two-Weapon Fighting & Add your Str bonus to off-hand damage rolls\\
\quad \quad Two-Weapon Rend & Double Slice & Rend a foe hit by both your weapons\\
\quad Improved Two-Weapon Fighting & Dex, Two-Weapon Fighting & Gain additional off-hand attack\\
\quad \quad Greater Two-Weapon Fighting & Dex, Improved Two-Weapon Fighting & Gain a third off-hand attack\\
\quad Two-Weapon Defense & Two-Weapon Fighting & Gain +1 shield bonus when fighting with two weapons\\
\end{tabularx}
\end{table*}
\begin{table*}
  \sffamily
  \setlength{\tabcolsep}{1pt}
\begin{tabularx}{\linewidth}{lXl}
Vital Strike & Base attack bonus +6 & Deal twice the normal damage on a single attack\\
\quad Improved Vital Strike & Vital Strike & Deal three times the normal damage on a single attack\\
\quad \quad Greater Vital Strike & Improved Vital Strike & Deal four times the normal damage on a single attack\\
Weapon Finesse & --- & Use Dex instead of Str on attack rolls with light weapons\\
Weapon Focus & Proficiency with weapon, base attack bonus +1 & +1 bonus on attack rolls with one weapon\\
\quad Dazzling Display & Weapon Focus & Intimidate all foes within 30 feet\\
\quad \quad Shatter Defenses & Dazzling Display & Hindered foes are flat-footed \\
\quad \quad \quad Deadly Stroke & Greater Weapon Focus, Shatter Defenses & Deal double damage plus 1 Con bleed \\
\quad Greater Weapon Focus & Weapon Focus & +1 bonus on attack rolls with one weapon\\
\quad Penetrating Strike & Weapon Focus & Your attacks ignore 5 points of damage reduction\\
\quad \quad Greater Penetrating Strike & Penetrating Strike & Your attacks ignore 10 points of damage reduction\\
\quad Weapon Specialization & Weapon Focus & +2 bonus on damage rolls with one weapon\\
\quad \quad Greater Weapon Specialization & Weapon Specialization & +2 bonus on damage rolls with one weapon\\
\textbf{Item Creation Feats} & \textbf{Prerequisites} & \textbf{Benefits}\\
Brew Potion & Caster level 3rd & Create magic potions\\
Craft Magic Arms and Armor & Caster level 5th & Create magic armors, shields, and weapons\\
Craft Rod & Caster level 9th & Create magic rods\\
Craft Staff & Caster level 11th & Create magic staves\\
Craft Wand & Caster level 5th & Create magic wands\\
Craft Wondrous Item & Caster level 3rd & Create magic wondrous items\\
Forge Ring & Caster level 7th & Create magic rings\\
Scribe Scroll & Caster level 1st & Create magic scrolls\\
\textbf{Metamagic Feats} & \textbf{Prerequisites} & \textbf{Benefits}\\
Empower Spell & --- & Increase spell variables by 50\%\\
Enlarge Spell & --- & Double spell range\\
Extend Spell & --- & Double spell duration\\
Heighten Spell & --- & Treat spell as a higher level\\
Maximize Spell & --- & Maximize spell variables\\
Quicken Spell & --- & Cast spell as a swift action\\
Silent Spell & --- & Cast spell without verbal components\\
Still Spell & --- & Cast spell without somatic components\\
Widen Spell & --- & Double spell area\\
\end{tabularx}
* This is a combat feat and can be selected as a fighter bonus feat
\end{table*}

				
\subsection{Acrobatic}

				
You are skilled at leaping, jumping, and flying.
				
\textbf{Benefit:} You get a +2 bonus on all Acrobatics and Fly skill checks. If you have 10 or more ranks in one of these skills, the bonus increases to +4 for that skill.
				
\subsection{Acrobatic Steps}

				
You can easily move over and through obstacles.
				
\textbf{Prerequisites:} Dex 15, Nimble Moves.
				
\textbf{Benefit:} Whenever you move, you may move through up to 15 feet of difficult terrain each round as if it were normal terrain. The effects of this feat stack with those provided by Nimble Moves (allowing you to move normally through a total of 20 feet of difficult terrain each round).
				
\subsection{Agile Maneuvers (Combat)}

				
You've learned to use your quickness in place of brute force when performing combat maneuvers.
				
\textbf{Benefit:} You add your Dexterity bonus to your base attack bonus and size bonus when determining your Combat Maneuver Bonus (see Combat) instead of your Strength bonus.
				
\textbf{Normal:} You add your Strength bonus to your base attack bonus and size bonus when determining your Combat Maneuver Bonus.
				
\subsection{Alertness}

				
You often notice things that others might miss.
				
\textbf{Benefit:} You get a +2 bonus on Perception and Sense Motive skill checks. If you have 10 or more ranks in one of these skills, the bonus increases to +4 for that skill.
				
\subsection{Alignment Channel}

				
Choose chaos, evil, good, or law. You can channel divine energy to affect outsiders that possess this subtype.
				
\textbf{Prerequisites:} Ability to channel energy.
				
\textbf{Benefit:} Instead of its normal effect, you can choose to have your ability to channel energy heal or harm outsiders of the chosen alignment subtype. You must make this choice each time you channel energy. If you choose to heal or harm creatures of the chosen alignment subtype, your channel energy has no effect on other creatures. The amount of damage healed or dealt and the DC to halve the damage is otherwise unchanged.
				
\textbf{Special:} You can gain this feat multiple times. Its effects do not stack. Each time you take this feat, it applies to a new alignment subtype. Whenever you channel energy, you must choose which type to effect.
				
\subsection{Animal Affinity}

				
You are skilled at working with animals and mounts.
				
\textbf{Benefit:} You get a +2 bonus on all Handle Animal and Ride skill checks. If you have 10 or more ranks in one of these skills, the bonus increases to +4 for that skill.
				
\subsection{Arcane Armor Mastery (Combat)}

				
You have mastered the ability to cast spells while wearing armor.
				
\textbf{Prerequisites:} Arcane Armor Training, Medium Armor Proficiency, caster level 7th.
				
\textbf{Benefit:} As a swift action, reduce the arcane spell failure chance due to the armor you are wearing by 20\% for any spells you cast this round. This bonus replaces, and does not stack with, the bonus granted by Arcane Armor Training.
				
\subsection{Arcane Armor Training (Combat)}

				
You have learned how to cast spells while wearing armor.
				
\textbf{Prerequisites:} Light Armor Proficiency, caster level 3rd.
				
\textbf{Benefit:} As a swift action, reduce the arcane spell failure chance due to the armor you are wearing by 10\% for any spells you cast this round.
				
\subsection{Arcane Strike (Combat)}

				
You draw upon your arcane power to enhance your weapons with magical energy.
				
\textbf{Prerequisite:} Ability to cast arcane spells.
				
\textbf{Benefit:} As a swift action, you can imbue your weapons with a fraction of your power. For 1 round, your weapons deal +1 damage and are treated as magic for the purpose of overcoming damage reduction. For every five caster levels you possess, this bonus increases by +1, to a maximum of +5 at 20th level.
				
\subsection{Armor Proficiency, Heavy (Combat)}

				
You are skilled at wearing heavy armor.
				
\textbf{Prerequisites:} Light Armor Proficiency, Medium Armor Proficiency.
				
\textbf{Benefit:} See Armor Proficiency, Light.
				
\textbf{Normal:} See Armor Proficiency, Light.
				
\textbf{Special:} Fighters and paladins automatically have Heavy Armor Proficiency as a bonus feat. They need not select it.
				
\subsection{Armor Proficiency, Light (Combat)}

				
You are skilled at wearing light armor.
				
\textbf{Benefit:} When you wear a type of armor with which you are proficient, the armor check penalty for that armor applies only to Dexterity- and Strength-based skill checks.
				
\textbf{Normal:} A character who is wearing armor with which he is not proficient applies its armor check penalty to attack rolls and to all skill checks that involve moving.
				
\textbf{Special:} All characters except monks, sorcerers, and wizards automatically have Light Armor Proficiency as a bonus feat. They need not select it.
				
\subsection{Armor Proficiency, Medium (Combat)}

				
You are skilled at wearing medium armor.
				
\textbf{Prerequisite:} Light Armor Proficiency.
				
\textbf{Benefit:} See Armor Proficiency, Light.
				
\textbf{Normal:} See Armor Proficiency, Light.
				
\textbf{Special:} Barbarians, clerics, druids, fighters, paladins, and rangers automatically have Medium Armor Proficiency as a bonus feat. They need not select it.
				
\subsection{Athletic}

				
You possess inherent physical prowess.
				
\textbf{Benefit:} You get a +2 bonus on Climb and Swim skill checks. If you have 10 or more ranks in one of these skills, the bonus increases to +4 for that skill.
				
\subsection{Augment Summoning}

				
Your summoned creatures are more powerful and robust.
				
\textbf{Prerequisite:} Spell Focus (conjuration).
				
\textbf{Benefit:} Each creature you conjure with any \textit{summon }spell gains a +4 enhancement bonus to Strength and Constitution for the duration of the spell that summoned it.
				
\subsection{Bleeding Critical (Combat, Critical)}

				
Your critical hits cause opponents to bleed profusely.
				
\textbf{Prerequisites:} Critical Focus, base attack bonus +11.
				
\textbf{Benefit:} Whenever you score a critical hit with a slashing or piercing weapon, your opponent takes 2d6 points of bleed damage (see Conditions) each round on his turn, in addition to the damage dealt by the critical hit. Bleed damage can be stopped by a DC 15 Heal skill check or through any magical healing. The effects of this feat stack.
				
\textbf{Special:} You can only apply the effects of one critical feat to a given critical hit unless you possess Critical Mastery.
				
\subsection{Blind-Fight (Combat)}

				
You are skilled at attacking opponents that you cannot clearly perceive.
				
\textbf{Benefit:} In melee, every time you miss because of concealment (see Combat), you can reroll your miss chance percentile roll one time to see if you actually hit.

An invisible attacker gets no advantages related to hitting you in melee. That is, you don't lose your Dexterity bonus to Armor Class, and the attacker doesn't get the usual +2 bonus for being invisible. The invisible attacker's bonuses do still apply for ranged attacks, however.

You do not need to make Acrobatics skill checks to move at full speed while blinded.
				
\textbf{Normal:} Regular attack roll modifiers for invisible attackers trying to hit you apply, and you lose your Dexterity bonus to AC. The speed reduction for darkness and poor visibility also applies.
				
\textbf{Special:} The Blind-Fight feat is of no use against a character who is the subject of a \textit{blink }spell.
				
\subsection{Blinding Critical (Combat, Critical)}

				
Your critical hits blind your opponents.
				
\textbf{Prerequisites:} Critical Focus, base attack bonus +15.
				
\textbf{Benefit:} Whenever you score a critical hit, your opponent is permanently blinded. A successful Fortitude save reduces this to dazzled for 1d4 rounds. The DC of this Fortitude save is equal to 10 + your base attack bonus. This feat has no effect on creatures that do not rely on eyes for sight or creatures with more than two eyes (although multiple critical hits might cause blindness, at the GM's discretion). Blindness can be cured by \textit{heal, regeneration, remove blindness, }or similar abilities.
				
\textbf{Special:} You can only apply the effects of one critical feat to a given critical hit unless you possess Critical Mastery.
				
\subsection{Brew Potion (Item Creation)}

				
You can create magic potions.
				
\textbf{Prerequisite:} Caster level 3rd.
				
\textbf{Benefit:} You can create a potion of any 3rd-level or lower spell that you know and that targets one or more creatures or objects. Brewing a potion takes 2 hours if its base price is 250 gp or less, otherwise brewing a potion takes 1 day for each 1,000 gp in its base price. When you create a potion, you set the caster level, which must be sufficient to cast the spell in question and no higher than your own level. To brew a potion, you must use up raw materials costing one half this base price. See the magic item creation rules in Magic Items for more information.

When you create a potion, you make any choices that you would normally make when casting the spell. Whoever drinks the potion is the target of the spell.
				
\subsection{Catch Off-Guard (Combat)}

				
Foes are surprised by your skilled use of unorthodox and improvised weapons.
				
\textbf{Benefit:} You do not suffer any penalties for using an improvised melee weapon. Unarmed opponents are flat-footed against any attacks you make with an improvised melee weapon.
				
\textbf{Normal:} You take a --4 penalty on attack rolls made with an improvised weapon.
				
\subsection{Channel Smite (Combat)}

				
You can channel your divine energy through a melee weapon you wield.
				
\textbf{Prerequisite:} Channel energy class feature.
				
\textbf{Benefit:} Before you make a melee attack roll, you can choose to spend one use of your channel energy ability as a swift action. If you channel positive energy and you hit an undead creature, that creature takes an amount of additional damage equal to the damage dealt by your channel positive energy ability. If you channel negative energy and you hit a living creature, that creature takes an amount of additional damage equal to the damage dealt by your channel negative energy ability. Your target can make a Will save, as normal, to halve this additional damage. If your attack misses, the channel energy ability is still expended with no effect.
				
\subsection{Cleave (Combat)}

				
You can strike two adjacent foes with a single swing.
				
\textbf{Prerequisites:} Str 13, Power Attack, base attack bonus +1.
				
\textbf{Benefit:} As a standard action, you can make a single attack at your full base attack bonus against a foe within reach. If you hit, you deal damage normally and can make an additional attack (using your full base attack bonus) against a foe that is adjacent to the first and also within reach. You can only make one additional attack per round with this feat. When you use this feat, you take a --2 penalty to your Armor Class until your next turn.
				
\subsection{Combat Casting}

				
You are adept at spellcasting when threatened or distracted.
				
\textbf{Benefit:} You get a +4 bonus on concentration checks made to cast a spell or use a spell-like ability when casting on the defensive or while grappled.
				
\subsection{Combat Expertise (Combat)}

				
You can increase your defense at the expense of your accuracy.
				
\textbf{Prerequisite:} Int 13.
				
\textbf{Benefit:} You can choose to take a --1 penalty on melee attack rolls and combat maneuver checks to gain a +1 dodge bonus to your Armor Class. When your base attack bonus reaches +4, and every +4 thereafter, the penalty increases by --1 and the dodge bonus increases by +1. You can only choose to use this feat when you declare that you are making an attack or a full-attack action with a melee weapon. The effects of this feat last until your next turn.
				
\subsection{Combat Reflexes (Combat)}

				
You can make additional attacks of opportunity.
				
\textbf{Benefit:} You may make a number of additional attacks of opportunity per round equal to your Dexterity bonus. With this feat, you may also make attacks of opportunity while flat-footed.
				
\textbf{Normal:} A character without this feat can make only one attack of opportunity per round and can't make attacks of opportunity while flat-footed.
				
\textbf{Special:} The Combat Reflexes feat does not allow a rogue to use her opportunist ability more than once per round.
				
\subsection{Command Undead}

				
Using foul powers of necromancy, you can command undead creatures, making them into your servants.
				
\textbf{Prerequisites:} Channel negative energy class feature.
				
\textbf{Benefit:} As a standard action, you can use one of your uses of channel negative energy to enslave undead within 30 feet. Undead receive a Will save to negate the effect. The DC for this Will save is equal to 10 + 1/2 your cleric level + your Charisma modifier. Undead that fail their saves fall under your control, obeying your commands to the best of their ability, as if under the effects of \textit{control undead}. Intelligent undead receive a new saving throw each day to resist your command. You can control any number of undead, so long as their total Hit Dice do not exceed your cleric level. If you use channel energy in this way, it has no other effect (it does not heal or harm nearby creatures). If an undead creature is under the control of another creature, you must make an opposed Charisma check whenever your orders conflict.
				
\subsection{Craft Magic Arms and Armor (Item Creation)}

				
You can create magic armor, shields, and weapons.
				
\textbf{Prerequisite:} Caster level 5th.
				
\textbf{Benefit:} You can create magic weapons, armor, or shields. Enhancing a weapon, suit of armor, or shield takes 1 day for each 1,000 gp in the price of its magical features. To enhance a weapon, suit of armor, or shield, you must use up raw materials costing half of this total price. See the magic item creation rules in Magic Items for more information.

The weapon, armor, or shield to be enhanced must be a masterwork item that you provide. Its cost is not included in the above cost.

You can also mend a broken magic weapon, suit of armor, or shield if it is one that you could make. Doing so costs half the raw materials and half the time it would take to craft that item in the first place.
				
\subsection{Craft Rod (Item Creation)}

				
You can create magic rods.
				
\textbf{Prerequisite:} Caster level 9th.
				
\textbf{Benefit:} You can create magic rods. Crafting a rod takes 1 day for each 1,000 gp in its base price. To craft a rod, you must use up raw materials costing half of its base price. See the magic item creation rules in Magic Items for more information.
				
\subsection{Craft Staff (Item Creation)}

				
You can create magic staves.
				
\textbf{Prerequisite:} Caster level 11th.
				
\textbf{Benefit:} You can create any staff whose prerequisites you meet. Crafting a staff takes 1 day for each 1,000 gp in its base price. To craft a staff, you must use up raw materials costing half of its base price. A newly created staff has 10 charges. See the magic item creation rules in Magic Items for more information.
				
\subsection{Craft Wand (Item Creation)}

				
You can create magic wands.
				
\textbf{Prerequisite:} Caster level 5th.
				
\textbf{Benefit:} You can create a wand of any 4th-level or lower spell that you know. Crafting a wand takes 1 day for each 1,000 gp in its base price. To craft a wand, you must use up raw materials costing half of this base price. A newly created wand has 50 charges. See the magic item creation rules in Magic Items for more information.
				
\subsection{Craft Wondrous Item (Item Creation)}

				
You can create wondrous items, a type of magic item.
				
\textbf{Prerequisite:} Caster level 3rd.
				
\textbf{Benefit:} You can create a wide variety of magic wondrous items. Crafting a wondrous item takes 1 day for each 1,000 gp in its price. To create a wondrous item, you must use up raw materials costing half of its base price. See the magic item creation rules in Magic Items for more information.

You can also mend a broken wondrous item if it is one that you could make. Doing so costs half the raw materials and half the time it would take to craft that item.
				
\subsection{Critical Focus (Combat)}

				
You are trained in the art of causing pain.
				
\textbf{Prerequisites:} Base attack bonus +9.
				
\textbf{Benefit:} You receive a +4 circumstance bonus on attack rolls made to confirm critical hits.
				
\subsection{Critical Mastery (Combat)}

				
Your critical hits cause two additional effects.
				
\textbf{Prerequisites:} Critical Focus, any two critical feats, 14th-level fighter.
				
\textbf{Benefit:} When you score a critical hit, you can apply the effects of two critical feats in addition to the damage dealt.
				
\textbf{Normal:} You can only apply the effects of one critical feat to a given critical hit in addition to the damage dealt.
				
\subsection{Dazzling Display (Combat)}

				
Your skill with your favored weapon can frighten enemies.
				
\textbf{Prerequisite:} Weapon Focus, proficiency with the selected weapon.
				
\textbf{Benefit:} While wielding the weapon in which you have Weapon Focus, you can perform a bewildering show of prowess as a full-round action. Make an Intimidate check to demoralize all foes within 30 feet who can see your display. 
				
\subsection{Deadly Aim (Combat)}

				
You can make exceptionally deadly ranged attacks by pinpointing a foe's weak spot, at the expense of making the attack less likely to succeed.
				
\textbf{Prerequisites:} Dex 13, base attack bonus +1.
				
\textbf{Benefit:} You can choose to take a --1 penalty on all ranged attack rolls to gain a +2 bonus on all ranged damage rolls. When your base attack bonus reaches +4, and every +4 thereafter, the penalty increases by --1 and the bonus to damage increases by +2. You must choose to use this feat before making an attack roll and its effects last until your next turn. The bonus damage does not apply to touch attacks or effects that do not deal hit point damage.
				
\subsection{Deadly Stroke (Combat)}

				
With a well-placed strike, you can bring a swift and painful end to most foes.
				
\textbf{Prerequisites:} Dazzling Display, Greater Weapon Focus, Shatter Defenses, Weapon Focus, proficiency with the selected weapon, base attack bonus +11.
				
\textbf{Benefit:} As a standard action, make a single attack with the weapon for which you have Greater Weapon Focus against a stunned or flat-footed opponent. If you hit, you deal double the normal damage and the target takes 1 point of Constitution bleed (see Conditions). The additional damage and bleed is not multiplied on a critical hit.
				
\subsection{Deafening Critical (Combat, Critical)}

				
Your critical hits cause enemies to lose their hearing.
				
\textbf{Prerequisites:} Critical Focus, base attack bonus +13.
				
\textbf{Benefit:} Whenever you score a critical hit against an opponent, the victim is permanently deafened. A successful Fortitude save reduces the deafness to 1 round. The DC of this Fortitude save is equal to 10 + your base attack bonus. This feat has no effect on deaf creatures. This deafness can be cured by \textit{heal, regeneration, remove deafness, }or a similar ability.
				
\textbf{Special:} You can only apply the effects of one critical feat to a given critical hit unless you possess Critical Mastery.
				
\subsection{Deceitful}

				
You are skilled at deceiving others, both with the spoken word and with physical disguises.
				
\textbf{Benefit:} You get a +2 bonus on all Bluff and Disguise skill checks. If you have 10 or more ranks in one of these skills, the bonus increases to +4 for that skill.
				
\subsection{Defensive Combat Training (Combat)}

				
You excel at defending yourself from all manner of combat maneuvers.
				
\textbf{Benefit:} You treat your total Hit Dice as your base attack bonus when calculating your Combat Maneuver Defense (see Combat). 
				
\subsection{Deflect Arrows (Combat)}

				
You can knock arrows and other projectiles off course, preventing them from hitting you.
				
\textbf{Prerequisites:} Dex 13, Improved Unarmed Strike.
				
\textbf{Benefit:} You must have at least one hand free (holding nothing) to use this feat. Once per round when you would normally be hit with an attack from a ranged weapon, you may deflect it so that you take no damage from it. You must be aware of the attack and not flat-footed. Attempting to deflect a ranged attack doesn't count as an action. Unusually massive ranged weapons (such as boulders or ballista bolts) and ranged attacks generated by natural attacks or spell effects can't be deflected.
				
\subsection{Deft Hands}

				
You have exceptional manual dexterity.
				
\textbf{Benefit:} You get a +2 bonus on Disable Device and Sleight of Hand skill checks. If you have 10 or more ranks in one of these skills, the bonus increases to +4 for that skill.
				
\subsection{Diehard}

				
You are especially hard to kill. Not only do your wounds automatically stabilize when grievously injured, but you can remain conscious and continue to act even at death's door.
				
\textbf{Prerequisite:} Endurance.
				
\textbf{Benefit:} When your hit point total is below 0, but you are not dead, you automatically stabilize. You do not need to make a Constitution check each round to avoid losing additional hit points. You may choose to act as if you were disabled, rather than dying. You must make this decision as soon as you are reduced to negative hit points (even if it isn't your turn). If you do not choose to act as if you were disabled, you immediately fall unconscious.

When using this feat, you are staggered. You can take a move action without further injuring yourself, but if you perform any standard action (or any other action deemed as strenuous, including some swift actions, such as casting a quickened spell) you take 1 point of damage after completing the act. If your negative hit points are equal to or greater than your Constitution score, you immediately die.
				
\textbf{Normal:} A character without this feat who is reduced to negative hit points is unconscious and dying.
				
\subsection{Disruptive (Combat)}

				
Your training makes it difficult for enemy spellcasters to safely cast spells near you.
				
\textbf{Prerequisites:} 6th-level fighter.
				
\textbf{Benefit:} The DC to cast spells defensively increases by +4 for all enemies that are within your threatened area. This increase to casting spells defensively only applies if you are aware of the enemy's location and are capable of taking an attack of opportunity. If you can only take one attack of opportunity per round and have already used that attack, this increase does not apply.
				
\subsection{Dodge (Combat)}

				
Your training and reflexes allow you to react swiftly to avoid an opponents' attacks.
				
\textbf{Prerequisite:} Dex 13.
				
\textbf{Benefit:} You gain a +1 dodge bonus to your AC. A condition that makes you lose your Dex bonus to AC also makes you lose the benefits of this feat.
				
\subsection{Double Slice (Combat)}

				
Your off-hand weapon while dual-wielding strikes with greater power.
				
\textbf{Prerequisite:} Dex 15, Two-Weapon Fighting.
				
\textbf{Benefit:} Add your Strength bonus to damage rolls made with your off-hand weapon.
				
\textbf{Normal:} You normally add only half of your Strength modifier to damage rolls made with a weapon wielded in your off-hand.
				
\subsection{Elemental Channel}

				
Choose one elemental subtype, such as air, earth, fire, or water. You can channel your divine energy to harm or heal outsiders that possess your chosen elemental subtype.
				
\textbf{Prerequisites:} Channel energy class feature.
				
\textbf{Benefit:} Instead of its normal effect, you can choose to have your ability to channel energy heal or harm outsiders of your chosen elemental subtype. You must make this choice each time you channel energy. If you choose to heal or harm creatures of your elemental subtype, your channel energy has no effect on other creatures. The amount of damage healed or dealt and the DC to halve the damage is otherwise unchanged.
				
\textbf{Special:} You can gain this feat multiple times. Its effects do not stack. Each time you take this feat, it applies to a new elemental subtype.
				
\subsection{Empower Spell (Metamagic)}

				
You can increase the power of your spells, causing them to deal more damage.
				
\textbf{Benefit:} All variable, numeric effects of an empowered spell are increased by half, including bonuses to those dice rolls.

Saving throws and opposed rolls are not affected, nor are spells without random variables. An empowered spell uses up a spell slot two levels higher than the spell's actual level.
				
\subsection{Endurance}

				
Harsh conditions or long exertions do not easily tire you.
				
\textbf{Benefit:} You gain a +4 bonus on the following checks and saves: Swim checks made to resist nonlethal damage from exhaustion; Constitution checks made to continue running; Constitution checks made to avoid nonlethal damage from a forced march; Constitution checks made to hold your breath; Constitution checks made to avoid nonlethal damage from starvation or thirst; Fortitude saves made to avoid nonlethal damage from hot or cold environments; and Fortitude saves made to resist damage from suffocation.

You may sleep in light or medium armor without becoming fatigued.
				
\textbf{Normal:} A character without this feat who sleeps in medium or heavier armor is fatigued the next day.
				
\subsection{Enlarge Spell (Metamagic)}

				
You can increase the range of your spells.
				
\textbf{Benefit:} You can alter a spell with a range of close, medium, or long to increase its range by 100\%. An enlarged spell with a range of close now has a range of 50 ft. + 5 ft./level, while medium-range spells have a range of 200 ft. + 20 ft./level and long-range spells have a range of 800 ft. + 80 ft./level. An enlarged spell uses up a spell slot one level higher than the spell's actual level.

Spells whose ranges are not defined by distance, as well as spells whose ranges are not close, medium, or long, do not benefit from this feat.
				
\subsection{Eschew Materials}

				
You can cast many spells without needing to utilize minor material components.
				
\textbf{Benefit:} You can cast any spell with a material component costing 1 gp or less without needing that component. The casting of the spell still provokes attacks of opportunity as normal. If the spell requires a material component that costs more than 1 gp, you must have the material component on hand to cast the spell, as normal.
				
\subsection{Exhausting Critical (Combat, Critical)}

				
Your critical hits cause opponents to become exhausted.
				
\textbf{Prerequisites:} Critical Focus, Tiring Critical, base attack bonus +15.
				
\textbf{Benefit:} When you score a critical hit on a foe, your target immediately becomes exhausted. This feat has no effect on exhausted creatures.
				
\textbf{Special:} You can only apply the effects of one critical feat to a given critical hit unless you possess the Critical Mastery feat.
				
\subsection{Exotic Weapon Proficiency (Combat)}

				
Choose one type of exotic weapon, such as the spiked chain or whip. You understand how to use that type of exotic weapon in combat, and can utilize any special tricks or qualities that exotic weapon might allow.
				
\textbf{Prerequisite:} Base attack bonus +1.
				
\textbf{Benefit:} You make attack rolls with the weapon normally.
				
\textbf{Normal:} A character who uses a weapon with which he is not proficient takes a --4 penalty on attack rolls.
				
\textbf{Special:} You can gain Exotic Weapon Proficiency multiple times. Each time you take the feat, it applies to a new type of exotic weapon. 
				
\subsection{Extend Spell (Metamagic)}

				
You can make your spells last twice as long.
				
\textbf{Benefit:} An extended spell lasts twice as long as normal. A spell with a duration of concentration, instantaneous, or permanent is not affected by this feat. An extended spell uses up a spell slot one level higher than the spell's actual level.
				
\subsection{Extra Channel}

				
You can channel divine energy more often.
				
\textbf{Prerequisite:} Channel energy class feature.
				
\textbf{Benefit:} You can channel energy two additional times per day.
				
\textbf{Special:} If a paladin with the ability to channel positive energy takes this feat, she can use lay on hands four additional times a day, but only to channel positive energy.
				
\subsection{Extra Ki}

				
You can use your ki pool more times per day than most.
				
\textbf{Prerequisite:} Ki pool class feature.
				
\textbf{Benefit:} Your \textit{ki }pool increases by 2. 
				
\textbf{Special:} You can gain Extra Ki multiple times. Its effects stack.
				
\subsection{Extra Lay On Hands}

				
You can use your lay on hands ability more often.
				
\textbf{Prerequisite:} Lay on hands class feature.
				
\textbf{Benefit:} You can use your lay on hands ability two additional times per day. 
				
\textbf{Special:} You can gain Extra Lay On Hands multiple times. Its effects stack.
				
\subsection{Extra Mercy}

				
Your lay on hands ability adds an additional mercy.
				
\textbf{Prerequisites:} Lay on hands class feature, mercy class feature.
				
\textbf{Benefit:} Select one additional mercy for which you qualify. When you use lay on hands to heal damage to one target, it also receives the additional effects of this mercy.
				
\textbf{Special:} You can gain this feat multiple times. Its effects do not stack. Each time you take this feat, select a new mercy. 
				
\subsection{Extra Performance}

				
You can use your bardic performance ability more often than normal.
				
\textbf{Prerequisite:} Bardic performance class feature.
				
\textbf{Benefit:} You can use bardic performance for 6 additional rounds per day.
				
\textbf{Special:} You can gain Extra Performance multiple times. Its effects stack.
				
\subsection{Extra Rage}

				
You can use your rage ability more than normal.
				
\textbf{Prerequisite:} Rage class feature.
				
\textbf{Benefit:} You can rage for 6 additional rounds per day.
				
\textbf{Special:} You can gain Extra Rage multiple times. Its effects stack.
				
\subsection{Far Shot (Combat)}

				
You are more accurate at longer ranges.
				
\textbf{Prerequisites:} Point-Blank Shot.
				
\textbf{Benefit:} You only suffer a --1 penalty per full range increment between you and your target when using a ranged weapon.
				
\textbf{Normal:} You suffer a --2 penalty per full range increment between you and your target.
				
\subsection{Fleet}

				
You are faster than most.
				
\textbf{Benefit:} While you are wearing light or no armor, your base speed increases by 5 feet. You lose the benefits of this feat if you carry a medium or heavy load.
				
\textbf{Special:} You can take this feat multiple times. The effects stack. 
				
\subsection{Forge Ring (Item Creation)}

				
You can create magic rings.
				
\textbf{Prerequisite:} Caster level 7th.
				
\textbf{Benefit:} You can create magic rings. Crafting a ring takes 1 day for each 1,000 gp in its base price. To craft a ring, you must use up raw materials costing half of the base price. See the magic item creation rules in Magic Items for more information.

You can also mend a broken ring if it is one that you could make. Doing so costs half the raw materials and half the time it would take to forge that ring in the first place.
				
\subsection{Gorgon's Fist (Combat)}

				
With one well-placed blow, you leave your target reeling.
				
\textbf{Prerequisites:} Improved Unarmed Strike, Scorpion Style, base attack bonus +6.
				
\textbf{Benefit:} As a standard action, make a single unarmed melee attack against a foe whose speed is reduced (such as from Scorpion Style). If the attack hits, you deal damage normally and the target is staggered until the end of your next turn unless it makes a Fortitude saving throw (DC 10 + 1/2 your character level + your Wis modifier). This feat has no effect on targets that are staggered.
				
\subsection{Great Cleave (Combat)}

				
You can strike many adjacent foes with a single blow.
				
\textbf{Prerequisites:} Str 13, Cleave, Power Attack, base attack bonus +4.
				
\textbf{Benefit:} As a standard action, you can make a single attack at your full base attack bonus against a foe within reach. If you hit, you deal damage normally and can make an additional attack (using your full base attack bonus) against a foe that is adjacent to the previous foe and also within reach. If you hit, you can continue to make attacks against foes adjacent to the previous foe, so long as they are within your reach. You cannot attack an individual foe more than once during this attack action. When you use this feat, you take a --2 penalty to your Armor Class until your next turn.
				
\subsection{Great Fortitude}

				
You are resistant to poisons, diseases, and other maladies.
				
\textbf{Benefit:} You get a +2 bonus on all Fortitude saving throws.
				
\subsection{Greater Bull Rush (Combat)}

				
Your bull rush attacks throw enemies off balance.
				
\textbf{Prerequisites:} Improved Bull Rush, Power Attack, base attack bonus +6, Str 13.
				
\textbf{Benefit:} You receive a +2 bonus on checks made to bull rush a foe. This bonus stacks with the bonus granted by Improved Bull Rush. Whenever you bull rush an opponent, his movement provokes attacks of opportunity from all of your allies (but not you).
				
\textbf{Normal:} Creatures moved by bull rush do not provoke attacks of opportunity.
				
\subsection{Greater Disarm (Combat)}

				
You can knock weapons far from an enemy's grasp.
				
\textbf{Prerequisites:} Combat Expertise, Improved Disarm, base attack bonus +6, Int 13.
				
\textbf{Benefit:} You receive a +2 bonus on checks made to disarm a foe. This bonus stacks with the bonus granted by Improved Disarm. Whenever you successfully disarm an opponent, the weapon lands 15 feet away from its previous wielder, in a random direction.
				
\textbf{Normal:} Disarmed weapons and gear land at the feet of the disarmed creature.
				
\subsection{Greater Feint (Combat)}

				
You are skilled at making foes overreact to your attacks.
				
\textbf{Prerequisites:} Combat Expertise, Improved Feint, base attack bonus +6, Int 13.
				
\textbf{Benefit:} Whenever you use feint to cause an opponent to lose his Dexterity bonus, he loses that bonus until the beginning of your next turn, in addition to losing his Dexterity bonus against your next attack.
				
\textbf{Normal:} A creature you feint loses its Dexterity bonus against your next attack.
				
\subsection{Greater Grapple (Combat)}

				
Maintaining a grapple is second nature to you.
				
\textbf{Prerequisites:} Improved Grapple, Improved Unarmed Strike, base attack bonus +6, Dex 13.
				
\textbf{Benefit:} You receive a +2 bonus on checks made to grapple a foe. This bonus stacks with the bonus granted by Improved Grapple. Once you have grappled a creature, maintaining the grapple is a move action. This feat allows you to make two grapple checks each round (to move, harm, or pin your opponent), but you are not required to make two checks. You only need to succeed at one of these checks to maintain the grapple.
				
\textbf{Normal:} Maintaining a grapple is a standard action.
				
\subsection{Greater Overrun (Combat)}

				
Enemies must dive to avoid your dangerous move.
				
\textbf{Prerequisites:} Improved Overrun, Power Attack, base attack bonus +6, Str 13.
				
\textbf{Benefit:} You receive a +2 bonus on checks made to overrun a foe. This bonus stacks with the bonus granted by Improved Overrun. Whenever you overrun opponents, they provoke attacks of opportunity if they are knocked prone by your overrun.
				
\textbf{Normal:} Creatures knocked prone by your overrun do not provoke an attack of opportunity.
				
\subsection{Greater Penetrating Strike (Combat)}

				
Your attacks penetrate the defenses of most foes. 
				
\textbf{Prerequisites:} Penetrating Strike, Weapon Focus, 16th-level fighter.
				
\textbf{Benefit:} Your attacks with weapons selected with Weapon Focus ignore up to 10 points of damage reduction. This amount is reduced to 5 points for damage reduction without a type (such as DR 10/---).
				
\subsection{Greater Shield Focus (Combat)}

				
You are skilled at deflecting blows with your shield.
				
\textbf{Prerequisites:} Shield Focus, Shield Proficiency, base attack bonus +1, 8th-level fighter.
				
\textbf{Benefit:} Increase the AC bonus granted by any shield you are using by 1. This bonus stacks with the bonus granted by Shield Focus.
				
\subsection{Greater Spell Focus}

				
Choose a school of magic to which you have already applied the Spell Focus feat. Any spells you cast of this school are very hard to resist. 
				
\textbf{Prerequisite:} Spell Focus.
				
\textbf{Benefit:} Add +1 to the Difficulty Class for all saving throws against spells from the school of magic you select. This bonus stacks with the bonus from Spell Focus.
				
\textbf{Special:} You can gain this feat multiple times. Its effects do not stack. Each time you take the feat, it applies to a new school to which you already have applied the Spell Focus feat.
				
\subsection{Greater Spell Penetration}

				
Your spells break through spell resistance much more easily than most.
				
\textbf{Prerequisite:} Spell Penetration.
				
\textbf{Benefit:} You get a +2 bonus on caster level checks (1d20 + caster level) made to overcome a creature's spell resistance. This bonus stacks with the one from Spell Penetration.
				
\subsection{Greater Sunder (Combat)}

				
Your devastating strikes cleave through weapons and armor and into their wielders, damaging both item and wielder alike in a single terrific strike.
				
\textbf{Prerequisites:} Improved Sunder, Power Attack, base attack bonus +6, Str 13.
				
\textbf{Benefit:} You receive a +2 bonus on checks made to sunder an item. This bonus stacks with the bonus granted by Improved Sunder. Whenever you sunder to destroy a weapon, shield, or suit of armor, any excess damage is applied to the item's wielder. No damage is transferred if you decide to leave the item with 1 hit point.
				
\subsection{Greater Trip (Combat)}

				
You can make free attacks on foes that you knock down.
				
\textbf{Prerequisites:} Combat Expertise, Improved Trip, base attack bonus +6, Int 13.
				
\textbf{Benefit:} You receive a +2 bonus on checks made to trip a foe. This bonus stacks with the bonus granted by Improved Trip. Whenever you successfully trip an opponent, that opponent provokes attacks of opportunity.
				
\textbf{Normal:} Creatures do not provoke attacks of opportunity from being tripped.
				
\subsection{Greater Two-Weapon Fighting (Combat)}

				
You are incredibly skilled at fighting with two weapons at the same time.
				
\textbf{Prerequisites:} Dex 19, Improved Two-Weapon Fighting, Two-Weapon Fighting, base attack bonus +11.
				
\textbf{Benefit:} You get a third attack with your off-hand weapon, albeit at a --10 penalty.
				
\subsection{Greater Vital Strike (Combat)}

				
You can make a single attack that deals incredible damage.
				
\textbf{Prerequisites:} Improved Vital Strike, Vital Strike, base attack bonus +16.
				
\textbf{Benefit:} When you use the attack action, you can make one attack at your highest base attack bonus that deals additional damage. Roll the weapon's damage dice for the attack four times and add the results together before adding bonuses from Strength, weapon abilities (such as \textit{flaming}), precision based damage, and other damage bonuses. These extra weapon damage dice are not multiplied on a critical hit, but are added to the total.
				
\subsection{Greater Weapon Focus (Combat)}

				
Choose one type of weapon (including unarmed strike or grapple) for which you have already selected Weapon Focus. You are a master at your chosen weapon.

\textbf{ Prerequisites}: Proficiency with selected weapon, Weapon Focus with selected weapon, base attack bonus +1, 8th-level fighter.
				
\textbf{Benefit:} You gain a +1 bonus on attack rolls you make using the selected weapon. This bonus stacks with other bonuses on attack rolls, including those from Weapon Focus.
				
\textbf{Special:} You can gain Greater Weapon Focus multiple times. Its effects do not stack. Each time you take the feat, it applies to a new type of weapon.
				
\subsection{Greater Weapon Specialization (Combat)}

				
Choose one type of weapon (including unarmed strike or grapple) for which you possess the Weapon Specialization feat. Your attacks with the chosen weapon are more devastating than normal.

\textbf{ Prerequisites}: Proficiency with selected weapon, Greater Weapon Focus with selected weapon, Weapon Focus with selected weapon, Weapon Specialization with selected weapon, 12th-level fighter.
				
\textbf{Benefit:} You gain a +2 bonus on all damage rolls you make using the selected weapon. This bonus to damage stacks with other damage roll bonuses, including any you gain from Weapon Specialization.
				
\textbf{Special:} You can gain Greater Weapon Specialization multiple times. Its effects do not stack. Each time you take the feat, it applies to a new type of weapon.
				
\subsection{Heighten Spell (Metamagic)}

				
You can cast spells as if they were a higher level.
				
\textbf{Benefit:} A heightened spell has a higher spell level than normal (up to a maximum of 9th level). Unlike other metamagic feats, Heighten Spell actually increases the effective level of the spell that it modifies. All effects dependent on spell level (such as saving throw DCs and ability to penetrate a \textit{lesser globe of invulnerability}) are calculated according to the heightened level. The heightened spell is as difficult to prepare and cast as a spell of its effective level. 
				
\subsection{Improved Bull Rush (Combat)}

				
You are skilled at pushing your foes around.
				
\textbf{Prerequisite:} Str 13, Power Attack, base attack bonus +1.
				
\textbf{Benefit:} You do not provoke an attack of opportunity when performing a bull rush combat maneuver. In addition, you receive a +2 bonus on checks made to bull rush a foe. You also receive a +2 bonus to your Combat Maneuver Defense whenever an opponent tries to bull rush you.
				
\textbf{Normal:} You provoke an attack of opportunity when performing a bull rush combat maneuver.
				
\subsection{Improved Channel}

				
Your channeled energy is harder to resist.
				
\textbf{Prerequisite:} Channel energy class feature.
				
\textbf{Benefit:} Add 2 to the DC of saving throws made to resist the effects of your channel energy ability. 
				
\subsection{Improved Counterspell}

				
You are skilled at countering the spells of others using similar spells.
				
\textbf{Benefit:} When counterspelling, you may use a spell of the same school that is one or more spell levels higher than the target spell.
				
\textbf{Normal:} Without this feat, you may counter a spell only with the same spell or with a spell specifically designated as countering the target spell.
				
\subsection{Improved Critical (Combat)}

				
Attacks made with your chosen weapon are quite deadly.
				
\textbf{Prerequisite:} Proficient with weapon, base attack bonus +8.
				
\textbf{Benefit:} When using the weapon you selected, your threat range is doubled.
				
\textbf{Special:} You can gain Improved Critical multiple times. The effects do not stack. Each time you take the feat, it applies to a new type of weapon.

This effect doesn't stack with any other effect that expands the threat range of a weapon.
				
\subsection{Improved Disarm (Combat)}

				
You are skilled at knocking weapons from a foe's grasp.
				
\textbf{Prerequisite:} Int 13, Combat Expertise.
				
\textbf{Benefit:} You do not provoke an attack of opportunity when performing a disarm combat maneuver. In addition, you receive a +2 bonus on checks made to disarm a foe. You also receive a +2 bonus to your Combat Maneuver Defense whenever an opponent tries to disarm you.
				
\textbf{Normal:} You provoke an attack of opportunity when performing a disarm combat maneuver.
				
\subsection{Improved Familiar}

				
This feat allows you to acquire a powerful familiar, but only when you could normally acquire a new familiar.
				
\textbf{Prerequisites:} Ability to acquire a new familiar, compatible alignment, sufficiently high level (see below).
				
\textbf{Benefit:} When choosing a familiar, the creatures listed below are also available to you. You may choose a familiar with an alignment up to one step away on each alignment axis (lawful through chaotic, good through evil).
\begin{table}
\sffamily
\begin{tabular}{lll}
                  &                    & \textbf{Arcane} \\
                  &                    & \textbf{Spellcaster} \\
\textbf{Familiar} & \textbf{Alignment} & \textbf{Level}\\

Celestial hawk\(^{1}\) & Neutral good & 3rd\\
Dire rat & Neutral & 3rd\\
Fiendish viper\(^{2}\) & Neutral evil & 3rd\\
Elemental, Small (any type) & Neutral & 5th\\
Stirge& Neutral & 5th\\
Homunculus\(^{3}\) & Any & 7th\\
Imp & Lawful evil & 7th\\
Mephit (any type) & Neutral & 7th\\
Pseudodragon & Neutral good & 7th\\
Quasit & Chaotic evil & 7th\\
\end{tabular}
\textsuperscript{1} Or other celestial animal from the standard familiar list. \\
\textsuperscript{2} Or other fiendish animal from the standard familiar list. \\
\textsuperscript{3} The master must first create the homunculus.

\end{table}

				
Improved familiars otherwise use the rules for regular familiars\textit{, }with two exceptions: if the creature's type is something other than animal, its type does not change; and improved familiars do not gain the ability to speak with other creatures of their kind (although many of them already have the ability to communicate).
				
\subsection{Improved Feint (Combat)}

				
You are skilled at fooling your opponents in combat.
				
\textbf{Prerequisites:} Int 13, Combat Expertise.
				
\textbf{Benefit:} You can make a Bluff check to feint in combat as a move action.
				
\textbf{Normal:} Feinting in combat is a standard action.
				
\subsection{Improved Grapple (Combat)}

				
You are skilled at grappling opponents.
				
\textbf{Prerequisite:} Dex 13, Improved Unarmed Strike.
				
\textbf{Benefit:} You do not provoke an attack of opportunity when performing a grapple combat maneuver. In addition, you receive a +2 bonus on checks made to grapple a foe. You also receive a +2 bonus to your Combat Maneuver Defense whenever an opponent tries to grapple you.
				
\textbf{Normal:} You provoke an attack of opportunity when performing a grapple combat maneuver.
				
\subsection{Improved Great Fortitude}

				
You can draw upon an inner reserve to resist diseases, poisons, and other grievous harm.
				
\textbf{Prerequisites:} Great Fortitude.
				
\textbf{Benefit:} Once per day, you may reroll a Fortitude save. You must decide to use this ability before the results are revealed. You must take the second roll, even if it is worse.
				
\subsection{Improved Initiative (Combat)}

				
Your quick reflexes allow you to react rapidly to danger.
				
\textbf{Benefit:} You get a +4 bonus on initiative checks.
				
\subsection{Improved Iron Will}

				
Your clarity of thought allows you to resist mental attacks.
				
\textbf{Prerequisites:} Iron Will.
				
\textbf{Benefit:} Once per day, you may reroll a Will save. You must decide to use this ability before the results are revealed. You must take the second roll, even if it is worse.
				
\subsection{Improved Lightning Reflexes}

				
You have a knack for avoiding danger all around you.
				
\textbf{Prerequisites:} Lightning Reflexes.
				
\textbf{Benefit:} Once per day, you may reroll a Reflex save. You must decide to use this ability before the results are revealed. You must take the second roll, even if it is worse.
				
\subsection{Improved Overrun (Combat)}

				
You are skilled at running down your foes.
				
\textbf{Prerequisite:} Str 13, Power Attack, base attack bonus +1.
				
\textbf{Benefit:} You do not provoke an attack of opportunity when performing an overrun combat maneuver. In addition, you receive a +2 bonus on checks made to overrun a foe. You also receive a +2 bonus to your Combat Maneuver Defense whenever an opponent tries to overrun you. Targets of your overrun attempt may not chose to avoid you.
				
\textbf{Normal:} You provoke an attack of opportunity when performing an overrun combat maneuver.
				
\subsection{Improved Precise Shot (Combat)}

				
Your ranged attacks ignore anything but total concealment and cover.
				
\textbf{Prerequisites:} Dex 19, Point-Blank Shot, Precise Shot, base attack bonus +11.
				
\textbf{Benefit:} Your ranged attacks ignore the AC bonus granted to targets by anything less than total cover, and the miss chance granted to targets by anything less than total concealment. Total cover and total concealment provide their normal benefits against your ranged attacks.
				
\textbf{Normal:} See the normal rules on the effects of cover and concealment in Combat. 
				
\subsection{Improved Shield Bash (Combat)}

				
You can protect yourself with your shield, even if you use it to attack.
				
\textbf{Prerequisite:} Shield Proficiency.
				
\textbf{Benefit:} When you perform a shield bash, you may still apply the shield's shield bonus to your AC.
				
\textbf{Normal:} Without this feat, a character that performs a shield bash loses the shield's shield bonus to AC until his next turn (see Equipment).
				
\subsection{Improved Sunder (Combat)}

				
You are skilled at damaging your foes' weapons and armor.
				
\textbf{Prerequisite:} Str 13, Power Attack, base attack bonus +1.
				
\textbf{Benefit:} You do not provoke an attack of opportunity when performing a sunder combat maneuver. In addition, you receive a +2 bonus on checks made to sunder an item. You also receive a +2 bonus to your Combat Maneuver Defense whenever an opponent tries to sunder your gear. 
				
\textbf{Normal:} You provoke an attack of opportunity when performing a sunder combat maneuver.
				
\subsection{Improved Trip (Combat)}

				
You are skilled at sending your opponents to the ground.
				
\textbf{Prerequisite:} Int 13, Combat Expertise.
				
\textbf{Benefit:} You do not provoke an attack of opportunity when performing a trip combat maneuver. In addition, you receive a +2 bonus on checks made to trip a foe. You also receive a +2 bonus to your Combat Maneuver Defense whenever an opponent tries to trip you. 
				
\textbf{Normal:} You provoke an attack of opportunity when performing a trip combat maneuver.
				
\subsection{Improved Two-Weapon Fighting (Combat)}

				
You are skilled at fighting with two weapons.
				
\textbf{Prerequisites:} Dex 17, Two-Weapon Fighting, base attack bonus +6.
				
\textbf{Benefit:} In addition to the standard single extra attack you get with an off-hand weapon, you get a second attack with it, albeit at a --5 penalty.
				
\textbf{Normal:} Without this feat, you can only get a single extra attack with an off-hand weapon.
				
\subsection{Improved Unarmed Strike (Combat)}

				
You are skilled at fighting while unarmed.
				
\textbf{Benefit:} You are considered to be armed even when unarmed---you do not provoke attacks of opportunity when you attack foes while unarmed. Your unarmed strikes can deal lethal or nonlethal damage, at your choice.
				
\textbf{Normal:} Without this feat, you are considered unarmed when attacking with an unarmed strike, and you can deal only nonlethal damage with such an attack.
				
\subsection{Improved Vital Strike (Combat)}

				
You can make a single attack that deals a large amount of damage.
				
\textbf{Prerequisites:} Vital Strike, base attack bonus +11.
				
\textbf{Benefit:} When you use the attack action, you can make one attack at your highest base attack bonus that deals additional damage. Roll the weapon's damage dice for the attack three times and add the results together before adding bonuses from Strength, weapon special abilities (such as \textit{flaming}), precision based damage, and other damage bonuses. These extra weapon damage dice are not multiplied on a critical hit, but are added to the total.
				
\subsection{Improvised Weapon Mastery (Combat)}

				
You can turn nearly any object into a deadly weapon, from a razor-sharp chair leg to a sack of flour.
				
\textbf{Prerequisites:} Catch Off-Guard or Throw Anything, base attack bonus +8.
				
\textbf{Benefit:} You do not suffer any penalties for using an improvised weapon. Increase the amount of damage dealt by the improvised weapon by one step (for example, 1d4 becomes 1d6) to a maximum of 1d8 (2d6 if the improvised weapon is two-handed). The improvised weapon has a critical threat range of 19--20, with a critical multiplier of \mbox{$\times$}2.
				
\subsection{Intimidating Prowess (Combat)}

				
Your physical might is intimidating to others.
				
\textbf{Benefit:} Add your Strength modifier to Intimidate skill checks in addition to your Charisma modifier.
				
\subsection{Iron Will}

				
You are more resistant to mental effects.
				
\textbf{Benefit:} You get a +2 bonus on all Will saving throws.
				
\subsection{Leadership}

\begin{table}[htb]
\sffamily
\caption{Effects of Leadership}
\begin{tabular}{lc*{6}{c}}
\textbf{Leadership} & \textbf{Cohort} & \multicolumn{6}{c}{Number of Followers by Level} \\
\textbf{Score} & \textbf{Level} & \textbf{1st} & \textbf{2nd} & \textbf{3rd} & \textbf{4th} & \textbf{5th} & \textbf{6th}\\
1 or lower & -- & -- & -- & -- & -- & -- & --\\
2 & 1st & -- & -- & -- & -- & -- & --\\
3 & 2nd & -- & -- & -- & -- & -- & --\\
4 & 3rd & -- & -- & -- & -- & -- & --\\
5 & 3rd & -- & -- & -- & -- & -- & --\\
6 & 4th & -- & -- & -- & -- & -- & --\\
7 & 5th & -- & -- & -- & -- & -- & --\\
8 & 5th & -- & -- & -- & -- & -- & --\\
9 & 6th & -- & -- & -- & -- & -- & --\\
10 & 7th & 5 & -- & -- & -- & -- & --\\
11 & 7th & 6 & -- & -- & -- & -- & --\\
12 & 8th & 8 & -- & -- & -- & -- & --\\
13 & 9th & 10 & 1 & -- & -- & -- & --\\
14 & 10th & 15 & 1 & -- & -- & -- & --\\
15 & 10th & 20 & 2 & 1 & -- & -- & --\\
16 & 11th & 25 & 2 & 1 & -- & -- & --\\
17 & 12th & 30 & 3 & 1 & 1 & -- & --\\
18 & 12th & 35 & 3 & 1 & 1 & -- & --\\
19 & 13th & 40 & 4 & 2 & 1 & 1 & --\\
20 & 14th & 50 & 5 & 3 & 2 & 1 & --\\
21 & 15th & 60 & 6 & 3 & 2 & 1 & 1\\
22 & 15th & 75 & 7 & 4 & 2 & 2 & 1\\
23 & 16th & 90 & 9 & 5 & 3 & 2 & 1\\
24 & 17th & 110 & 11 & 6 & 3 & 2 & 1\\
25 & 17th & 135 & 13 & 7 & 4 & 2 & 2\\
\end{tabular}
\end{table}

				
You attract followers to your cause and a companion to join you on your adventures.
				
\textbf{Prerequisite:} Character level 7th.
				
\textbf{Benefits:} This feat enables you to attract a loyal cohort and a number of devoted subordinates who assist you. A cohort is generally an NPC with class levels, while followers are typically lower level NPCs. See Table: Leadership for what level of cohort and how many followers you can recruit.
				
\textbf{Leadership Modifiers:} Several factors can affect your Leadership score, causing it to vary from the base score (character level + Cha modifier). Your reputation (from the point of view of the cohort or follower you are trying to attract) raises or lowers your Leadership score:
				
\begin{table}[htb]
\sffamily
\rowcolors{1}{white}{offyellow}
\caption{Leadership Reputation Modifiers}
\centering
\begin{tabular}{lccc}
\textbf{Leader's Reputation} & \textbf{Modifier}\\
Great renown & +2\\
Fairness and generosity & +1\\
Special Power & +1\\
Failure & -1 \\
Aloofness & -1\\
Cruelty & -2\\
\end{tabular}
\end{table}
				
Other modifiers may apply when you try to attract a cohort, as listed below.\newline

\begin{table}[htb]
\sffamily
\rowcolors{1}{white}{offyellow}
\caption{Cohort Modifiers}
\centering
\begin{tabular}{lc}
\textbf{The Leader\ldots{}} & \textbf{Modifier}\\
Has a familiar, special mount, or animal companion & -2\\
Recruits a cohort of a different alignment & -1\\
Caused the death of a cohort & -2\textsuperscript{1}\\
\multicolumn{2}{l}{\textsuperscript{1} Cumulative per cohort killed.}\\
\end{tabular}
\end{table}

				
Followers have different priorities from cohorts. When you try to attract a follower, use the following modifiers.\newline

\begin{table}[htb]
\sffamily
\rowcolors{1}{white}{offyellow}
\caption{Follower Modifiers}
\centering
\begin{tabularx}{\linewidth}{Xc}
\textbf{The Leader\ldots{}} & \textbf{Modifier}\\
Has a stronghold, base of operations, guildhouse, or the like & +2\\
Moves around a lot & -1\\
Caused the death of other followes & -1\\
\end{tabularx}
\end{table}
				
\textit{Leadership Score}: Your base Leadership score equals your level plus your Charisma modifier. In order to take into account negative Charisma modifiers, this table allows for very low Leadership scores, but you must still be 7th level or higher in order to gain the Leadership feat. Outside factors can affect your Leadership score, as detailed above.
				
\textit{Cohort Level}: You can attract a cohort of up to this level. Regardless of your Leadership score, you can only recruit a cohort who is two or more levels lower than yourself. The cohort should be equipped with gear appropriate for its level (see Creating NPCs). A cohort can be of any race or class. The cohort's alignment may not be opposed to your alignment on either the law/chaos or good/evil axis, and you take a --1 penalty to your Leadership score if you recruit a cohort of an alignment different from your own.\newline
				
A cohort does not count as a party member when determining the party's XP. Instead, divide the cohort's level by your level. Multiply this result by the total XP awarded to you, then add that number of experience points to the cohort's total.

If a cohort gains enough XP to bring it to a level one lower than your level, the cohort does not gain the new level---its new XP total is 1 less than the amount needed to attain the next level. 
				
\textit{Number of Followers by Level}: You can lead up to the indicated number of characters of each level. Followers are similar to cohorts, except they're generally low-level NPCs. Because they're usually 5 or more levels behind you, they're rarely effective in combat.
                
Followers don't earn experience and thus don't gain levels. When you gain a new level, consult Table: Leadership to determine if you acquire more followers, some of whom may be higher level than the existing followers. Don't consult the table to see if your cohort gains levels, however, because cohorts earn experience on their own.
				
\subsection{Lightning Reflexes}

				
You have faster reflexes than normal.
				
\textbf{Benefit:} You get a +2 bonus on all Reflex saving throws.
				
\subsection{Lightning Stance (Combat)}

				
The speed at which you move makes it nearly impossible for opponents to strike you.
				
\textbf{Prerequisites:} Dex 17, Dodge, Wind Stance, base attack bonus +11.
				
\textbf{Benefit:} If you take two actions to move or a withdraw action in a turn, you gain 50\% concealment for 1 round.
				
\subsection{Lunge (Combat)}

				
You can strike foes that would normally be out of reach.
				
\textbf{Prerequisites:} Base attack bonus +6.
				
\textbf{Benefit:} You can increase the reach of your melee attacks by 5 feet until the end of your turn by taking a --2 penalty to your AC until your next turn. You must decide to use this ability before any attacks are made.
				
\subsection{Magical Aptitude}

				
You are skilled at spellcasting and using magic items.
				
\textbf{Benefit:} You get a +2 bonus on all Spellcraft checks and Use Magic Device checks. If you have 10 or more ranks in one of these skills, the bonus increases to +4 for that skill.
				
\subsection{Manyshot (Combat)}

				
You can fire multiple arrows at a single target.
				
\textbf{Prerequisites:} Dex 17, Point-Blank Shot, Rapid Shot, base attack bonus +6.
				
\textbf{Benefit:} When making a full-attack action with a bow, your first attack fires two arrows. If the attack hits, both arrows hit. Apply precision-based damage (such as sneak attack) and critical hit damage only once for this attack. Damage bonuses from using a composite bow with a high Strength bonus apply to each arrow, as do other damage bonuses, such as a ranger's favored enemy bonus. Damage reduction and resistances apply separately to each arrow.
				
\subsection{Martial Weapon Proficiency (Combat)}

				
Choose a type of martial weapon. You understand how to use that type of martial weapon in combat.
				
\textbf{Benefit:} You make attack rolls with the selected weapon normally (without the non-proficient penalty).
				
\textbf{Normal:} When using a weapon with which you are not proficient, you take a --4 penalty on attack rolls.
				
\textbf{Special:} Barbarians, fighters, paladins, and rangers are proficient with all martial weapons. They need not select this feat.

You can gain Martial Weapon Proficiency multiple times. Each time you take the feat, it applies to a new type of weapon. 
				
\subsection{Master Craftsman}

				
Your superior crafting skills allow you to create simple magic items.
				
\textbf{Prerequisites:} 5 ranks in any Craft or Profession skill.
				
\textbf{Benefit:} Choose one Craft or Profession skill in which you possess at least 5 ranks. You receive a +2 bonus on your chosen Craft or Profession skill. Ranks in your chosen skill count as your caster level for the purposes of qualifying for the Craft Magic Arms and Armor and Craft Wondrous Item feats. You can create magic items using these feats, substituting your ranks in the chosen skill for your total caster level. You must use the chosen skill for the check to create the item. The DC to create the item still increases for any necessary spell requirements (see the magic item creation rules in Magic Items). You cannot use this feat to create any spell-trigger or spell-activation item.
				
\textbf{Normal:} Only spellcasters can qualify for the Craft Magic Arms and Armor and Craft Wondrous Item feats.
				
\subsection{Maximize Spell (Metamagic)}

				
Your spells have the maximum possible effect.
				
\textbf{Benefit:} All variable, numeric effects of a spell modified by this feat are maximized. Saving throws and opposed rolls are not affected, nor are spells without random variables. A maximized spell uses up a spell slot three levels higher than the spell's actual level.

An empowered, maximized spell gains the separate benefits of each feat: the maximum result plus half the normally rolled result.
				
\subsection{Medusa's Wrath (Combat)}

				
You can take advantage of your opponent's confusion, delivering multiple blows.
				
\textbf{Prerequisites:} Improved Unarmed Strike, Gorgon's Fist, Scorpion Style, base attack bonus +11.
				
\textbf{Benefit:} Whenever you use the full-attack action and make at least one unarmed strike, you can make two additional unarmed strikes at your highest base attack bonus. These bonus attacks must be made against a dazed, flat-footed, paralyzed, staggered, stunned, or unconscious foe.
				
\subsection{Mobility (Combat)}

				
You can easily move through a dangerous melee.
				
\textbf{Prerequisites:} Dex 13, Dodge.
				
\textbf{Benefit:} You get a +4 dodge bonus to Armor Class against attacks of opportunity caused when you move out of or within a threatened area. A condition that makes you lose your Dexterity bonus to Armor Class (if any) also makes you lose dodge bonuses.

Dodge bonuses stack with each other, unlike most types of bonuses.
				
\subsection{Mounted Archery (Combat)}

				
You are skilled at making ranged attacks while mounted.
				
\textbf{Prerequisites:} Ride 1 rank, Mounted Combat.
				
\textbf{Benefit:} The penalty you take when using a ranged weapon while mounted is halved: --2 instead of --4 if your mount is taking a double move, and --4 instead of --8 if your mount is running.
				
\subsection{Mounted Combat (Combat)}

				
You are adept at guiding your mount through combat.
				
\textbf{Prerequisite:} Ride 1 rank.
				
\textbf{Benefit:} Once per round when your mount is hit in combat, you may attempt a Ride check (as an immediate action) to negate the hit. The hit is negated if your Ride check result is greater than the opponent's attack roll.
				
\subsection{Natural Spell}

				
You can cast spells even while in a form that cannot normally cast spells.
				
\textbf{Prerequisites:} Wis 13, wild shape class feature.
				
\textbf{Benefit:} You can complete the verbal and somatic components of spells while using wild shape. You substitute various noises and gestures for the normal verbal and somatic components of a spell.

You can also use any material components or focuses you possess, even if such items are melded within your current form. This feat does not permit the use of magic items while you are in a form that could not ordinarily use them, and you do not gain the ability to speak while using wild shape.
				
\subsection{Nimble Moves}

				
You can move across a single obstacle with ease.
				
\textbf{Prerequisites:} Dex 13.
				
\textbf{Benefit:} Whenever you move, you may move through 5 feet of difficult terrain each round as if it were normal terrain. This feat allows you to take a 5-foot step into difficult terrain.
				
\subsection{Penetrating Strike (Combat)}

				
Your attacks are capable of penetrating the defenses of some creatures.
				
\textbf{Prerequisites:} Weapon Focus, base attack bonus +1, 12th-level fighter, proficiency with weapon.
				
\textbf{Benefit:} Your attacks with weapons selected with Weapon Focus ignore up to 5 points of damage reduction. This feat does not apply to damage reduction without a type (such as DR 10/---).
				
\subsection{Persuasive}

				
You are skilled at swaying attitudes and intimidating others into your way of thinking.
				
\textbf{Benefit:} You get a +2 bonus on Diplomacy and Intimidate skill checks. If you have 10 or more ranks in one of these skills, the bonus increases to +4 for that skill.
				
\subsection{Pinpoint Targeting (Combat)}

				
You can target the weak points in your opponent's armor.
				
\textbf{Prerequisites:} Dex 19, Improved Precise Shot, Point-Blank Shot, Precise Shot, base attack bonus +16.
				
\textbf{Benefit:} As a standard action, make a single ranged attack. The target does not gain any armor, natural armor, or shield bonuses to its Armor Class. You do not gain the benefit of this feat if you move this round.
				
\subsection{Point-Blank Shot (Combat)}

				
You are especially accurate when making ranged attacks against close targets.
				
\textbf{Benefit:} You get a +1 bonus on attack and damage rolls with ranged weapons at ranges of up to 30 feet.
				
\subsection{Power Attack (Combat)}

				
You can make exceptionally deadly melee attacks by sacrificing accuracy for strength.
				
\textbf{Prerequisites:} Str 13, base attack bonus +1.
				
\textbf{Benefit:} You can choose to take a --1 penalty on all melee attack rolls and combat maneuver checks to gain a +2 bonus on all melee damage rolls. This bonus to damage is increased by half (+50\%) if you are making an attack with a two-handed weapon, a one handed weapon using two hands, or a primary natural weapon that adds 1-1/2 times your Strength modifier on damage rolls. This bonus to damage is halved (--50\%) if you are making an attack with an off-hand weapon or secondary natural weapon. When your base attack bonus reaches +4, and every 4 points thereafter, the penalty increases by --1 and the bonus to damage increases by +2. You must choose to use this feat before making an attack roll, and its effects last until your next turn. The bonus damage does not apply to touch attacks or effects that do not deal hit point damage.
				
\subsection{Precise Shot (Combat)}

				
You are adept at firing ranged attacks into melee.
				
\textbf{Prerequisite:} Point-Blank Shot.
				
\textbf{Benefit:} You can shoot or throw ranged weapons at an opponent engaged in melee without taking the standard --4 penalty on your attack roll.
				
\subsection{Quick Draw (Combat)}

				
You can draw weapons faster than most.
				
\textbf{Prerequisite:} Base attack bonus +1.
				
\textbf{Benefit:} You can draw a weapon as a free action instead of as a move action. You can draw a hidden weapon (see the Sleight of Hand skill) as a move action.

A character who has selected this feat may throw weapons at his full normal rate of attacks (much like a character with a bow).

Alchemical items, potions, scrolls, and wands cannot be drawn quickly using this feat.
				
\textbf{Normal:} Without this feat, you may draw a weapon as a move action, or (if your base attack bonus is +1 or higher) as a free action as part of movement. Without this feat, you can draw a hidden weapon as a standard action.
				
\subsection{Quicken Spell (Metamagic)}

				
You can cast spells in a fraction of the normal time.
				
\textbf{Benefit:} Casting a quickened spell is a swift action. You can perform another action, even casting another spell, in the same round as you cast a quickened spell. A spell whose casting time is more than 1 round or 1 full-round action cannot be quickened.

A quickened spell uses up a spell slot four levels higher than the spell's actual level. Casting a quickened spell doesn't provoke an attack of opportunity.
				
\textbf{Special:} You can apply the effects of this feat to a spell cast spontaneously, so long as it has a casting time that is not more than 1 full-round action, without increasing the spell's casting time.
				
\subsection{Rapid Reload (Combat)}

				
Choose a type of crossbow (hand, light, or heavy). You can reload such weapons quickly.
				
\textbf{Prerequisite:} Weapon Proficiency (crossbow type chosen).
				
\textbf{Benefit:} The time required for you to reload your chosen type of crossbow is reduced to a free action (for a hand or light crossbow) or a move action (for a heavy crossbow). Reloading a crossbow still provokes an attack of opportunity.

If you have selected this feat for hand crossbow or light crossbow, you may fire that weapon as many times in a full-attack action as you could attack if you were using a bow.
				
\textbf{Normal:} A character without this feat needs a move action to reload a hand or light crossbow, or a full-round action to reload a heavy crossbow. 
				
\textbf{Special:} You can gain Rapid Reload multiple times. Each time you take the feat, it applies to a new type of crossbow.
				
\subsection{Rapid Shot (Combat)}

				
You can make an additional ranged attack.
				
\textbf{Prerequisites:} Dex 13, Point-Blank Shot.
				
\textbf{Benefit:} When making a full-attack action with a ranged weapon, you can fire one additional time this round at your highest bonus. All of your attack rolls take a --2 penalty when using Rapid Shot.
				
\subsection{Ride-By Attack (Combat)}

				
While mounted and charging, you can move, strike at a foe, and then continue moving.
				
\textbf{Prerequisites:} Ride 1 rank, Mounted Combat.
				
\textbf{Benefit:} When you are mounted and use the charge action, you may move and attack as if with a standard charge and then move again (continuing the straight line of the charge). Your total movement for the round can't exceed double your mounted speed. You and your mount do not provoke an attack of opportunity from the opponent that you attack.
				
\subsection{Run}

				
You are swift of foot.
				
\textbf{Benefit:} When running, you move five times your normal speed (if wearing medium, light, or no armor and carrying no more than a medium load) or four times your speed (if wearing heavy armor or carrying a heavy load). If you make a jump after a running start (see the Acrobatics skill description), you gain a +4 bonus on your Acrobatics check. While running, you retain your Dexterity bonus to your Armor Class.
				
\textbf{Normal:} You move four times your speed while running (if wearing medium, light, or no armor and carrying no more than a medium load) or three times your speed (if wearing heavy armor or carrying a heavy load), and you lose your Dexterity bonus to AC.
				
\subsection{Scorpion Style (Combat)}

				
You can perform an unarmed strike that greatly hampers your target's movement.
				
\textbf{Prerequisite:} Improved Unarmed Strike.
				
\textbf{Benefit:} To use this feat, you must make a single unarmed attack as a standard action. If this unarmed attack hits, you deal damage normally, and the target's base land speed is reduced to 5 feet for a number of rounds equal to your Wisdom modifier unless it makes a Fortitude saving throw (DC 10 + 1/2 your character level + your Wis modifier).
				
\subsection{Scribe Scroll (Item Creation)}

				
You can create magic scrolls.
				
\textbf{Prerequisite:} Caster level 1st.
				
\textbf{Benefit:} You can create a scroll of any spell that you know. Scribing a scroll takes 2 hours if its base price is 250 gp or less, otherwise scribing a scroll takes 1 day for each 1,000 gp in its base price. To scribe a scroll, you must use up raw materials costing half of this base price. See the magic item creation rules in Magic Items for more information.
				
\subsection{Selective Channeling}

				
You can choose whom to affect when you channel energy.
				
\textbf{Prerequisite:} Cha 13, channel energy class feature.
				
\textbf{Benefit:} When you channel energy, you can choose a number of targets in the area up to your Charisma modifier. These targets are not affected by your channeled energy.
				
\textbf{Normal:} All targets in a 30-foot burst are affected when you channel energy. You can only choose whether or not you are affected.
				
\subsection{Self-Sufficient}

				
You know how to get along in the wild and how to effectively treat wounds.
				
\textbf{Benefit:} You get a +2 bonus on all Heal checks and Survival checks. If you have 10 or more ranks in one of these skills, the bonus increases to +4 for that skill.
				
\subsection{Shatter Defenses (Combat)}

				
Your skill with your chosen weapon leaves opponents unable to defend themselves if you strike them when their defenses are already compromised.
				
\textbf{Prerequisites:} Weapon Focus, Dazzling Display, base attack bonus +6, proficiency with weapon.
				
\textbf{Benefit:} Any shaken, frightened, or panicked opponent hit by you this round is flat-footed to your attacks until the end of your next turn. This includes any additional attacks you make this round.
				
\subsection{Shield Focus (Combat)}

				
You are skilled at deflecting blows with your shield.
				
\textbf{Prerequisites:} Shield Proficiency, base attack bonus +1.
				
\textbf{Benefit:} Increase the AC bonus granted by any shield you are using by 1.
				
\subsection{Shield Master (Combat)}

				
Your mastery of the shield allows you to fight with it without hindrance.
				
\textbf{Prerequisites:} Improved Shield Bash, Shield Proficiency, Shield Slam, Two-Weapon Fighting, base attack bonus +11.
				
\textbf{Benefit:} You do not suffer any penalties on attack rolls made with a shield while you are wielding another weapon. Add your shield's enhancement bonus to attacks and damage rolls made with the shield as if it was a weapon enhancement bonus.
				
\subsection{Shield Proficiency (Combat)}

				
You are trained in how to properly use a shield.
				
\textbf{Benefit:} When you use a shield (except a tower shield), the shield's armor check penalty only applies to Strength- and Dexterity-based skills.
				
\textbf{Normal:} When you are using a shield with which you are not proficient, you take the shield's armor check penalty on attack rolls and on all skill checks that involve moving.
				
\textbf{Special:} Barbarians, bards, clerics, druids, fighters, paladins, and rangers all automatically have Shield Proficiency as a bonus feat. They need not select it.
				
\subsection{Shield Slam (Combat)}

				
In the right position, your shield can be used to send opponents flying.
				
\textbf{Prerequisites:} Improved Shield Bash, Shield Proficiency, Two-Weapon Fighting, base attack bonus +6.
				
\textbf{Benefit:} Any opponents hit by your shield bash are also hit with a free bull rush attack, substituting your attack roll for the combat maneuver check (see Combat). This bull rush does not provoke an attack of opportunity. Opponents who cannot move back due to a wall or other surface are knocked prone after moving the maximum possible distance. You may choose to move with your target if you are able to take a 5-foot step or to spend an action to move this turn.
				
\subsection{Shot on the Run (Combat)}

				
You can move, fire a ranged weapon, and move again before your foes can react.
				
\textbf{Prerequisites:} Dex 13, Dodge, Mobility, Point-Blank Shot, base attack bonus +4.
				
\textbf{Benefit:} As a full-round action, you can move up to your speed and make a single ranged attack at any point during your movement.
				
\textbf{Normal:} You cannot move before and after an attack with a ranged weapon.
				
\subsection{Sickening Critical (Combat, Critical)}

				
Your critical hits cause opponents to become sickened.
				
\textbf{Prerequisites:} Critical Focus, base attack bonus +11.
				
\textbf{Benefit:} Whenever you score a critical hit, your opponent becomes sickened for 1 minute. The effects of this feat do not stack. Additional hits instead add to the effect's duration.
				
\textbf{Special:} You can only apply the effects of one critical feat to a given critical hit unless you possess Critical Mastery.
				
\subsection{Silent Spell (Metamagic)}

				
You can cast your spells without making any sound.
				
\textbf{Benefit:} A silent spell can be cast with no verbal components. Spells without verbal components are not affected. A silent spell uses up a spell slot one level higher than the spell's actual level.
				
\textbf{Special:} Bard spells cannot be enhanced by this feat.
				
\subsection{Simple Weapon Proficiency (Combat)}

				
You are trained in the use of basic weapons.
				
\textbf{Benefit:} You make attack rolls with simple weapons without penalty.
				
\textbf{Normal:} When using a weapon with which you are not proficient, you take a --4 penalty on attack rolls.
				
\textbf{Special:} All characters except for druids, monks, and wizards are automatically proficient with all simple weapons. They need not select this feat.
				
\subsection{Skill Focus}

				
Choose a skill. You are particularly adept at that skill.
				
\textbf{Benefit:} You get a +3 bonus on all checks involving the chosen skill. If you have 10 or more ranks in that skill, this bonus increases to +6.
				
\textbf{Special:} You can gain this feat multiple times. Its effects do not stack. Each time you take the feat, it applies to a new skill.
				
\subsection{Snatch Arrows (Combat)}

				
Instead of knocking an arrow or ranged attack aside, you can catch it in mid-flight.
				
\textbf{Prerequisites:} Dex 15, Deflect Arrows, Improved Unarmed Strike.
				
\textbf{Benefit:} When using the Deflect Arrows feat you may choose to catch the weapon instead of just deflecting it. Thrown weapons can immediately be thrown back as an attack against the original attacker (even though it isn't your turn) or kept for later use.

You must have at least one hand free (holding nothing) to use this feat.
				
\subsection{Spell Focus}

				
Choose a school of magic. Any spells you cast of that school are more difficult to resist.
				
\textbf{Benefit:} Add +1 to the Difficulty Class for all saving throws against spells from the school of magic you select.
				
\textbf{Special:} You can gain this feat multiple times. Its effects do not stack. Each time you take the feat, it applies to a new school of magic.
				
\subsection{Spell Mastery}

				
You have mastered a small handful of spells, and can prepare these spells without referencing your spellbooks at all.
				
\textbf{Prerequisite:} 1st-level wizard
				
\textbf{Benefit:} Each time you take this feat, choose a number of spells that you already know equal to your Intelligence modifier. From that point on, you can prepare these spells without referring to a spellbook.
				
\textbf{Normal:} Without this feat, you must use a spellbook to prepare all your spells, except \textit{read magic.}
				
\subsection{Spell Penetration}

				
Your spells break through spell resistance more easily than most.
				
\textbf{Benefit:} You get a +2 bonus on caster level checks (1d20 + caster level) made to overcome a creature's spell resistance.
				
\subsection{Spellbreaker (Combat)}

				
You can strike at enemy spellcasters who fail to cast defensively when you threaten them.
				
\textbf{Prerequisites:} Disruptive, 10th-level fighter.
				
\textbf{Benefit:} Enemies in your threatened area that fail their checks to cast spells defensively provoke attacks of opportunity from you.
				
\textbf{Normal:} Enemies that fail to cast spells defensively do not provoke attacks of opportunity.
				
\subsection{Spirited Charge (Combat)}

				
Your mounted charge attacks deal a tremendous amount of damage.
				
\textbf{Prerequisites:} Ride 1 rank, Mounted Combat, Ride-By Attack.
				
\textbf{Benefit:} When mounted and using the charge action, you deal double damage with a melee weapon (or triple damage with a lance).
				
\subsection{Spring Attack (Combat)}

				
You can deftly move up to a foe, strike, and withdraw before he can react.
				
\textbf{Prerequisites:} Dex 13, Dodge, Mobility, base attack bonus +4.
				
\textbf{Benefit:} As a full-round action, you can move up to your speed and make a single melee attack without provoking any attacks of opportunity from the target of your attack. You can move both before and after the attack, but you must move at least 10 feet before the attack and the total distance that you move cannot be greater than your speed. You cannot use this ability to attack a foe that is adjacent to you at the start of your turn.
				
\textbf{Normal:} You cannot move before and after an attack.
				
\subsection{Staggering Critical (Combat, Critical)}

				
Your critical hits cause opponents to slow down.
				
\textbf{Prerequisites:} Critical Focus, base attack bonus +13.
				
\textbf{Benefit:} Whenever you score a critical hit, your opponent becomes staggered for 1d4+1 rounds. A successful Fortitude save reduces the duration to 1 round. The DC of this Fortitude save is equal to 10 + your base attack bonus. The effects of this feat do not stack. Additional hits instead add to the duration.
				
\textbf{Special:} You can only apply the effects of one critical feat to a given critical hit unless you possess Critical Mastery.
				
\subsection{Stand Still (Combat)}

				
You can stop foes that try to move past you.
				
\textbf{Prerequisites:} Combat Reflexes.
				
\textbf{Benefit:} When a foe provokes an attack of opportunity due to moving through your adjacent squares, you can make a combat maneuver check as your attack of opportunity. If successful, the enemy cannot move for the rest of his turn. An enemy can still take the rest of his action, but cannot move. This feat also applies to any creature that attempts to move from a square that is adjacent to you if such movement provokes an attack of opportunity.
				
\subsection{Stealthy}

				
You are good at avoiding unwanted attention and slipping out of bonds.
				
\textbf{Benefit:} You get a +2 bonus on all Escape Artist and Stealth skill checks. If you have 10 or more ranks in one of these skills, the bonus increases to +4 for that skill.
				
\subsection{Step Up (Combat)}

				
You can close the distance when a foe tries to move away.
				
\textbf{Prerequisite:} Base attack bonus +1.
				
\textbf{Benefit:} Whenever an adjacent foe attempts to take a 5-foot step away from you, you may also make a 5-foot step as an immediate action so long as you end up adjacent to the foe that triggered this ability. If you take this step, you cannot take a 5-foot step during your next turn. If you take an action to move during your next turn, subtract 5 feet from your total movement.
				
\subsection{Still Spell (Metamagic)}

				
You can cast spells without moving.
				
\textbf{Benefit:} A stilled spell can be cast with no somatic components. Spells without somatic components are not affected. A stilled spell uses up a spell slot one level higher than the spell's actual level.
				
\subsection{Strike Back (Combat)}

				
You can strike at foes that attack you using their superior reach, by targeting their limbs or weapons as they come at you.
				
\textbf{Prerequisite:} Base attack bonus +11.
				
\textbf{Benefit:} You can ready an action to make a melee attack against any foe that attacks you in melee, even if the foe is outside of your reach.
				
\subsection{Stunning Critical (Combat, Critical)}

				
Your critical hits cause opponents to become stunned.
				
\textbf{Prerequisites:} Critical Focus, Staggering Critical, base attack bonus +17.
				
\textbf{Benefit:} Whenever you score a critical hit, your opponent becomes stunned for 1d4 rounds. A successful Fortitude save reduces this to staggered for 1d4 rounds. The DC of this Fortitude save is equal to 10 + your base attack bonus. The effects of this feat do not stack. Additional hits instead add to the duration.
				
\textbf{Special:} You can only apply the effects of one critical feat to a given critical hit unless you possess Critical Mastery.
				
\subsection{Stunning Fist (Combat)}

				
You know just where to strike to temporarily stun a foe.
				
\textbf{Prerequisites:} Dex 13, Wis 13, Improved Unarmed Strike, base attack bonus +8.
				
\textbf{Benefit:} You must declare that you are using this feat before you make your attack roll (thus, a failed attack roll ruins the attempt). Stunning Fist forces a foe damaged by your unarmed attack to make a Fortitude saving throw (DC 10 + 1/2 your character level + your Wis modifier), in addition to dealing damage normally. A defender who fails this saving throw is stunned for 1 round (until just before your next turn). A stunned character drops everything held, can't take actions, loses any Dexterity bonus to AC, and takes a --2 penalty to AC. You may attempt a stunning attack once per day for every four levels you have attained (but see Special), and no more than once per round. Constructs, oozes, plants, undead, incorporeal creatures, and creatures immune to critical hits cannot be stunned.
				
\textbf{Special:} A monk receives Stunning Fist as a bonus feat at 1st level, even if he does not meet the prerequisites. A monk may attempt a stunning attack a number of times per day equal to his monk level, plus one more time per day for every four levels he has in classes other than monk.
				
\subsection{Throw Anything (Combat)}

				
You are used to throwing things you have on hand.
				
\textbf{Benefit:} You do not suffer any penalties for using an improvised ranged weapon. You receive a +1 circumstance bonus on attack rolls made with thrown splash weapons.
				
\textbf{Normal:} You take a --4 penalty on attack rolls made with an improvised weapon.
				
\subsection{Tiring Critical (Combat, Critical)}

				
Your critical hits cause opponents to become fatigued.
				
\textbf{Prerequisites:} Critical Focus, base attack bonus +13.
				
\textbf{Benefit:} Whenever you score a critical hit, your opponent becomes fatigued. This feat has no additional effect on a fatigued or exhausted creature.
				
\textbf{Special:} You can only apply the effects of one critical feat to a given critical hit unless you possess Critical Mastery.
				
\subsection{Toughness}

				
You have enhanced physical stamina.
				
\textbf{Benefit:} You gain +3 hit points. For every Hit Die you possess beyond 3, you gain an additional +1 hit point. If you have more than 3 Hit Dice, you gain +1 hit points whenever you gain a Hit Die (such as when you gain a level).
				
\subsection{Tower Shield Proficiency (Combat)}

				
You are trained in how to properly use a tower shield.
				
\textbf{Prerequisite:} Shield Proficiency.
				
\textbf{Benefit:} When you use a tower shield, the shield's armor check penalty only applies to Strength and Dexterity-based skills.
				
\textbf{Normal:} A character using a shield with which he is not proficient takes the shield's armor check penalty on attack rolls and on all skill checks that involve moving, including Ride.
				
\textbf{Special:} Fighters automatically have Tower Shield Proficiency as a bonus feat. They need not select it.
				
\subsection{Trample (Combat)}

				
While mounted, you can ride down opponents and trample them under your mount.
				
\textbf{Prerequisites:} Ride 1 rank, Mounted Combat.
				
\textbf{Benefit:} When you attempt to overrun an opponent while mounted, your target may not choose to avoid you. Your mount may make one hoof attack against any target you knock down, gaining the standard +4 bonus on attack rolls against prone targets.
				
\subsection{Turn Undead}

				
Calling upon higher powers, you cause undead to flee from the might of your unleashed divine energy.
				
\textbf{Prerequisites:} Channel positive energy class feature.
				
\textbf{Benefit:} You can, as a standard action, use one of your uses of channel positive energy to cause all undead within 30 feet of you to flee, as if panicked. Undead receive a Will save to negate the effect. The DC for this Will save is equal to 10 + 1/2 your cleric level + your Charisma modifier. Undead that fail their save flee for 1 minute. Intelligent undead receive a new saving throw each round to end the effect. If you use channel energy in this way, it has no other effect (it does not heal or harm nearby creatures).
				
\subsection{Two-Weapon Defense (Combat)}

				
You are skilled at defending yourself while dual-wielding.
				
\textbf{Prerequisites:} Dex 15, Two-Weapon Fighting.
				
\textbf{Benefit:} When wielding a double weapon or two weapons (not including natural weapons or unarmed strikes), you gain a +1 shield bonus to your AC.

When you are fighting defensively or using the total defense action, this shield bonus increases to +2.
				
\subsection{Two-Weapon Fighting (Combat)}

				
You can fight with a weapon wielded in each of your hands. You can make one extra attack each round with the secondary weapon.
				
\textbf{Prerequisite:} Dex 15.
				
\textbf{Benefit:} Your penalties on attack rolls for fighting with two weapons are reduced. The penalty for your primary hand lessens by 2 and the one for your off hand lessens by 6. See Two-Weapon Fighting in Combat.
				
\textbf{Normal:} If you wield a second weapon in your off hand, you can get one extra attack per round with that weapon. When fighting in this way you suffer a --6 penalty with your regular attack or attacks with your primary hand and a --10 penalty to the attack with your off hand. If your off-hand weapon is light, the penalties are reduced by 2 each. An unarmed strike is always considered light.
				
\subsection{Two-Weapon Rend (Combat)}

				
Striking with both of your weapons simultaneously, you can use them to deliver devastating wounds.
				
\textbf{Prerequisites:} Dex 17, Double Slice, Improved Two-Weapon Fighting, Two-Weapon Fighting, base attack bonus +11. 
				
\textbf{Benefit:} If you hit an opponent with both your primary hand and your off-hand weapon, you deal an additional 1d10 points of damage plus 1-1/2 times your Strength modifier. You can only deal this additional damage once each round.
				
\subsection{Unseat (Combat)}

				
You are skilled at unseating your mounted opponents.
				
\textbf{Prerequisites:} Str 13, Ride 1 rank, Mounted Combat, Power Attack, Improved Bull Rush, base attack bonus +1.
				
\textbf{Benefits:} When charging an opponent while mounted and wielding a lance, resolve the attack as normal. If it hits, you may immediately make a free bull rush attempt in addition to the normal damage. If successful, the target is knocked off his horse and lands prone in a space adjacent to his mount that is directly away from you.
				
\subsection{Vital Strike (Combat)}

				
You make a single attack that deals significantly more damage than normal.
				
\textbf{Prerequisites:} Base attack bonus +6.
				
\textbf{Benefit:} When you use the attack action, you can make one attack at your highest base attack bonus that deals additional damage. Roll the weapon's damage dice for the attack twice and add the results together before adding bonuses from Strength, weapon abilities (such as \textit{flaming}), precision based damage, and other damage bonuses. These extra weapon damage dice are not multiplied on a critical hit, but are added to the total.
				
\subsection{Weapon Finesse (Combat)}

				
You are trained in using your agility in melee combat, as opposed to brute strength.
				
\textbf{Benefit:} With a light weapon, rapier, whip, or spiked chain made for a creature of your size category, you may use your Dexterity modifier instead of your Strength modifier on attack rolls. If you carry a shield, its armor check penalty applies to your attack rolls.
				
\textbf{Special:} Natural weapons are considered light weapons.
				
\subsection{Weapon Focus (Combat)}

				
Choose one type of weapon. You can also choose unarmed strike or grapple (or ray, if you are a spellcaster) as your weapon for the purposes of this feat.
				
\textbf{Prerequisites:} Proficiency with selected weapon, base attack bonus +1.
				
\textbf{Benefit:} You gain a +1 bonus on all attack rolls you make using the selected weapon.
				
\textbf{Special:} You can gain this feat multiple times. Its effects do not stack. Each time you take the feat, it applies to a new type of weapon.
				
\subsection{Weapon Specialization (Combat)}

				
You are skilled at dealing damage with one weapon. Choose one type of weapon (including unarmed strike or grapple) for which you have already selected the Weapon Focus feat. You deal extra damage when using this weapon.
				
\textbf{Prerequisites:} Proficiency with selected weapon, Weapon Focus with selected weapon, fighter level 4th.
				
\textbf{Benefit:} You gain a +2 bonus on all damage rolls you make using the selected weapon.
				
\textbf{Special:} You can gain this feat multiple times. Its effects do not stack. Each time you take the feat, it applies to a new type of weapon.
				
\subsection{Whirlwind Attack (Combat)}

				
You can strike out at every foe within reach.
				
\textbf{Prerequisites:} Dex 13, Int 13, Combat Expertise, Dodge, Mobility, Spring Attack, base attack bonus +4.
				
\textbf{Benefit:} When you use the full-attack action, you can give up your regular attacks and instead make one melee attack at your highest base attack bonus against each opponent within reach. You must make a separate attack roll against each opponent.

When you use the Whirlwind Attack feat, you also forfeit any bonus or extra attacks granted by other feats, spells, or abilities.
				
\subsection{Widen Spell (Metamagic)}

				
You can cast your spells so that they occupy a larger space.
				
\textbf{Benefit:} You can alter a burst, emanation, or spread-shaped spell to increase its area. Any numeric measurements of the spell's area increase by 100\%. A widened spell uses up a spell slot three levels higher than the spell's actual level.

Spells that do not have an area of one of these four sorts are not affected by this feat.
				
\subsection{Wind Stance (Combat)}

				
Your erratic movements make it difficult for enemies to pinpoint your location.
				
\textbf{Prerequisites:} Dex 15, Dodge, base attack bonus +6.
				
\textbf{Benefit:} If you move more than 5 feet this turn, you gain 20\% concealment for 1 round against ranged attacks.
                
% </div class="stat-block-1">
		
