\chapter{Spells}
\section{Spell Lists}

\label{f0}        
An \(^{M}\) or \(^{F}\) appearing at the end of a spell's name in the spell lists denotes a spell with a material or focus component, respectively, that is not normally included in a spell component pouch.

\textbf{Order of Presentation}: In the spell lists and the short descriptions that follow them, the spells are presented in alphabetical order by name except for those belonging to certain spell chains. When a spell's name begins with \texttt{{}"{}}lesser,\texttt{{}"{}} \texttt{{}"{}}greater,\texttt{{}"{}} or \texttt{{}"{}}mass,\texttt{{}"{}} the spell is alphabetized under the second word of the spell name instead.

\textbf{Hit Dice}: The term \texttt{{}"{}}Hit Dice\texttt{{}"{}} is used synonymously with \texttt{{}"{}}character levels\texttt{{}"{}} for effects that affect a specific number of Hit Dice of creatures. Creatures with Hit Dice only from their race, not from classes, still have character levels equal to their Hit Dice.

\textbf{Caster Level}: A spell's power often depends on caster level, which is defined as the caster's class level for the purpose of casting a particular spell. A creature with no classes has a caster level equal to its Hit Dice unless otherwise specified. The word \texttt{{}"{}}level\texttt{{}"{}} in the short descriptions that follow always refers to caster level.

\textbf{Creatures and Characters}: The words \texttt{{}"{}}creature\texttt{{}"{}} and \texttt{{}"{}}character\texttt{{}"{}} are used synonymously in the short descriptions.

\section{Bard Spells}


\subsection{0-Level Bard Spells}


\textbf{Dancing Lights}: Creates torches or other lights.

\textbf{Daze}: Humanoid creature of 4 HD or less loses next action.

\textbf{Detect Magic}: Detects spells and magic items within 60 ft.

\textbf{Flare}: Dazzles one creature (--1 on attack rolls).

\textbf{Ghost Sound}: Figment sounds.

\textbf{Know Direction}: You discern north.

\textbf{Light}: Object shines like a torch.

\textbf{Lullaby}: Makes subject drowsy: --5 on Perception checks, --2 on Will saves against \textit{sleep}.

\textbf{Mage Hand}: 5-pound telekinesis.

\textbf{Mending}: Makes minor repairs on an object.

\textbf{Message}: Whisper conversation at distance.

\textbf{Open/Close}: Opens or closes small or light things.

\textbf{Prestidigitation}: Performs minor tricks.

\textbf{Read Magic}: Read scrolls and spellbooks.

\textbf{Resistance}: Subject gains +1 on saving throws.

\textbf{Summon Instrument}: Summons one musical instrument.

\subsection{1st-Level Bard Spells}


\textbf{Alarm}: Wards an area for 2 hours/level.

\textbf{Animate Rope}: Makes a rope move at your command.

\textbf{Cause Fear}: One creature of 5 HD or less flees for 1d4 rounds.

\textbf{Charm Person}: Makes one person your friend.

\textbf{Comprehend Languages}: You understand all languages.

\textbf{Confusion, Lesser}: One creature is \textit{confused }for 1 round.

\textbf{Cure Light Wounds}: Cures 1d8 damage + 1/level (max +5).

\textbf{Detect Secret Doors}: Reveals hidden doors within 60 ft.

\textbf{Disguise Self}: Changes your appearance.

\textbf{Erase}: Mundane or magical writing vanishes.

\textbf{Expeditious Retreat}: Your base land speed increases by 30 ft.

\textbf{Feather Fall}: Objects or creatures fall slowly.

\textbf{Grease}: Makes 10-ft. square or one object slippery.

\textbf{Hideous Laughter}: Subject loses actions for 1 round/ level.

\textbf{Hypnotism}: Fascinates 2d4 HD of creatures.

\textbf{Identify}: Gives +10 bonus to identify magic items.

\textbf{Magic Aura}: Alters object's magic aura.

\textbf{Magic Mouth}\(^{M}\): Objects speaks once when triggered.

\textbf{Obscure Object}: Masks object against \textit{scrying}.

\textbf{Remove Fear}: Suppresses fear or gives +4 on saves against fear for one subject + one per four levels.

\textbf{Silent Image}: Creates minor illusion of your design.

\textbf{Sleep}: Puts 4 HD of creatures into magical slumber.

\textbf{Summon Monster I}: Summons extraplanar creature to fight for you.

\textbf{Undetectable Alignment}: Conceals alignment for 24 hours.

\textbf{Unseen Servant}: Invisible force obeys your commands.

\textbf{Ventriloquism}: Throws voice for 1 min./level.

\subsection{2nd-Level Bard Spells}


\textbf{Alter Self}: Assume form of a Small or Medium humanoid.

\textbf{Animal Messenger}: Sends a Tiny animal to a specific place.

\textbf{Animal Trance}: Fascinates 2d6 HD of animals.

\textbf{Blindness/Deafness}: Makes subject blind or deaf.

\textbf{Blur}: Attacks miss subject 20\% of the time.

\textbf{Calm Emotions}: Calms creatures, negating emotion effects.

\textbf{Cat's Grace}: Subject gains +4 to Dex for 1 min./level.

\textbf{Cure Moderate Wounds}: Cures 2d8 damage + 1/level (max +10).

\textbf{Darkness}: 20-ft. radius of supernatural shadow.

\textbf{Daze Monster}: Living creature of 6 HD or less loses next action.

\textbf{Delay Poison}: Stops poison from harming target for 1 hour/level.

\textbf{Detect Thoughts}: Allows \texttt{{}"{}}listening\texttt{{}"{}} to surface thoughts.

\textbf{Eagle's Splendor}: Subject gains +4 to Cha for 1 min./level.

\textbf{Enthrall}: Captivates all within 100 ft. + 10 ft./level.

\textbf{Fox's Cunning}: Subject gains +4 to Int for 1 min./level.

\textbf{Glitterdust}: Blinds creatures, outlines invisible creatures.

\textbf{Heroism}: Gives +2 on attack rolls, saves, skill checks.

\textbf{Hold Person}: Paralyzes one humanoid for 1 round/level.

\textbf{Hypnotic Pattern}: Fascinates (2d4 + level) HD of creatures.

\textbf{Invisibility}: Subject is invisible for 1 min./level or until it attacks.

\textbf{Locate Object}: Senses direction toward object (specific or type).

\textbf{Minor Image}: As \textit{silent image}, plus some sound.

\textbf{Mirror Image}: Creates decoy duplicates of you (1d4 + 1 per three levels, max 8).

\textbf{Misdirection}: Misleads divinations for one creature or object.

\textbf{Pyrotechnics}: Turns fire into blinding light or choking smoke.

\textbf{Rage}: Gives +2 to Str and Con, +1 on Will saves, --2 to AC.

\textbf{Scare}: Panics creatures of less than 6 HD.

\textbf{Shatter}: Sonic vibration damages objects or crystalline creatures.

\textbf{Silence}: Negates sound in 20-ft. radius.

\textbf{Sound Burst}: Deals 1d8 sonic damage and may stun subjects.

\textbf{Suggestion}: Compels subject to follow stated course of action.

\textbf{Summon Monster II}: Summons extraplanar creature to fight for you.

\textbf{Summon Swarm}: Summons swarm of bats, rats, or spiders.

\textbf{Tongues}: Speak and understand any language.

\textbf{Whispering Wind}: Sends a short message 1 mile/level.

\subsection{3rd-Level Bard Spells}


\textbf{Blink}: You randomly vanish and reappear for 1 round/level.

\textbf{Charm Monster}: Makes monster believe it is your ally.

\textbf{Clairaudience/Clairvoyance}: Hear or see at a distance for 1 min./level.

\textbf{Confusion}: Subjects behave oddly for 1 round/level.

\textbf{Crushing Despair}: Subjects take --2 on attack rolls, damage rolls, saves, and checks.

\textbf{Cure Serious Wounds}: Cures 3d8 damage + 1/level (max +15).

\textbf{Daylight}: 60-ft. radius of bright light.

\textbf{Deep Slumber}: Puts 10 HD of creatures to sleep.

\textbf{Dispel Magic}: Cancels one magical spell or effect.

\textbf{Displacement}: Attacks miss subject 50\% of the time.

\textbf{Fear}: Subjects within cone flee for 1 round/level.

\textbf{Gaseous Form}: Subject becomes insubstantial and can fly slowly.

\textbf{Geas, Lesser}: Commands subject of 7 HD or less.

\textbf{Glibness}: You gain +20 bonus on Bluff checks, and your lies can escape magical discernment.

\textbf{Good Hope}: Subjects gain +2 on attack rolls, damage rolls, saves, and checks.

\textbf{Haste}: One creature/level moves faster, receives +1 on attack rolls, AC, and Reflex saves.

\textbf{Illusory Script}\(^{M}\): Only select creatures can read text.

\textbf{Invisibility Sphere}: Makes everyone within 10 feet invisible.

\textbf{Major Image}: As \textit{silent image}, plus sound, smell, and thermal effects.

\textbf{Phantom Steed}: Magic horse appears for 1 hour/level.

\textbf{Remove Curse}: Frees object or person from curse.

\textbf{Scrying}\(^{F}\): Spies on subject from a distance.

\textbf{Sculpt Sound}: Creates new sounds or changes existing ones into new sounds.

\textbf{Secret Page}: Changes one page to hide its real content.

\textbf{See Invisibility}: Reveals invisible creatures or objects.

\textbf{Sepia Snake Sigil}\(^{M}\): Creates a snake-shaped text symbol that immobilizes reader.

\textbf{Slow}: One subject/level takes only one action/round, --1 to AC, Reflex saves, and attack rolls.

\textbf{Speak with Animals}: You can communicate with animals.

\textbf{Summon Monster III}: Summons extraplanar creature to fight for you.

\textbf{Tiny Hut}: Creates shelter for 10 creatures.

\subsection{4th-Level Bard Spells}


\textbf{Break Enchantment}: Frees subjects from enchantments, transmutations, and curses.

\textbf{Cure Critical Wounds}: Cures 4d8 damage + 1/level (max +20).

\textbf{Detect Scrying}: Alerts you to magical eavesdropping.

\textbf{Dimension Door}: Teleports you a short distance.

\textbf{Dominate Person}: Controls humanoid telepathically.

\textbf{Freedom of Movement}: Subject moves normally despite impediments to restrict movement.

\textbf{Hallucinatory Terrain}: Makes one type of terrain appear like another (field as forest, or the like).

\textbf{Hold Monster}: As \textit{hold person}, but any creature.

\textbf{Invisibility, Greater}: As \textit{invisibility, }but subject can attack and stay invisible.

\textbf{Legend Lore}\(^{MF}\): Lets you learn tales about a person, place, or thing.

\textbf{Locate Creature}: Indicates direction to known creature.

\textbf{Modify Memory}: Changes 5 minutes of subject's memories.

\textbf{Neutralize Poison}: Immunizes subject against poison, detoxifies venom in or on subject.

\textbf{Rainbow Pattern}: Lights fascinate 24 HD of creatures.

\textbf{Repel Vermin}: Insects, spiders, and other vermin stay 10 ft. away.

\textbf{Secure Shelter}: Creates sturdy cottage.

\textbf{Shadow Conjuration}: Mimics conjuration below 4th level, but only 20\% real.

\textbf{Shout}: Deafens all within cone and deals 5d6 sonic damage.

\textbf{Speak with Plants}: You can talk to plants and plant creatures.

\textbf{Summon Monster IV}: Summons extraplanar creature to fight for you.

\textbf{Zone of Silence}: Keeps eavesdroppers from overhearing you.

\subsection{5th-Level Bard Spells}


\textbf{Cure Light Wounds, Mass}: Cures 1d8 damage + 1/level, affects 1 subject/level.

\textbf{Dispel Magic, Greater}: Works as \textit{dispel magic,} but can affect multiple targets.

\textbf{Dream}: Sends message to anyone sleeping.

\textbf{False Vision}\(^{M}\): Fools scrying with an illusion.

\textbf{Heroism, Greater}: Gives +4 bonus on attack rolls, saves, skill checks; immunity to fear; temporary hp.

\textbf{Mind Fog}: Subjects in fog get --10 to Wis and Will checks.

\textbf{Mirage Arcana}: As \textit{hallucinatory terrain, }plus structures.

\textbf{Mislead}: Turns you invisible and creates illusory double.

\textbf{Nightmare}: Sends vision dealing 1d10 damage, fatigue.

\textbf{Persistent Image}: As \textit{major image}, but no concentration required.

\textbf{Seeming}: Changes appearance of one person per two levels.

\textbf{Shadow Evocation}: Mimics evocation of lower than 5th level, but only 20\% real.

\textbf{Shadow Walk}: Step into shadow to travel rapidly.

\textbf{Song of Discord}: Forces targets to attack each other.

\textbf{Suggestion, Mass}: As \textit{suggestion, }affects subject/level.

\textbf{Summon Monster V}: Summons extraplanar creature to fight for you.

\subsection{6th-Level Bard Spells}


\textbf{Analyze Dweomer}\(^{F}\): Reveals magical aspects of subject.

\textbf{Animate Objects}: Objects attack your foes.

\textbf{Cat's Grace, Mass}: As \textit{cat's grace}, affects 1 subject/level.

\textbf{Charm Monster, Mass}: As \textit{charm monster, }but all within 30 ft.

\textbf{Cure Moderate Wounds, Mass}: Cures 2d8 damage + 1/level, affects 1 subject/level.

\textbf{Eagle's Splendor, Mass}: As \textit{eagle's splendor}, affects 1 subject/level.

\textbf{Eyebite}: Target becomes panicked, sickened, and/or comatose.

\textbf{Find the Path}: Shows most direct way to a location.

\textbf{Fox's Cunning, Mass}: As \textit{fox's cunning}, affects 1 subject/level.

\textbf{Geas/Quest}: As \textit{lesser geas, }but affects any creature.

\textbf{Heroes' Feast}: Food for one creature/level cures and grants combat bonuses\textit{.}

\textbf{Irresistible Dance}: Forces subject to dance.

\textbf{Permanent Image}: Permanent illusion, includes sight, sound, smell, and thermal effects.

\textbf{Programmed Image}\(^{M}\): As \textit{major image}, plus triggered by event.

\textbf{Project Image}\(^{M}\): Illusory double can talk and cast spells.

\textbf{Scrying, Greater}: As \textit{scrying}, but faster and longer.

\textbf{Shout, Greater}: Devastating yell deals 10d6 sonic damage; stuns creatures.

\textbf{Summon Monster VI}: Summons extraplanar creature to fight for you.

\textbf{Sympathetic Vibration}: Deals 2d10 damage/round to freestanding structure.

\textbf{Veil}: Changes appearance of group of creatures.

\section{Cleric Spells}


\subsection{0-Level Cleric Spells (Orisons)}


\textbf{Bleed}: Cause a stabilized creature to resume dying.

\textbf{Create Water}: Creates 2 gallons/level of pure water.

\textbf{Detect Magic}: Detects spells and magic items within 60 ft.

\textbf{Detect Poison}: Detects poison in one creature or object.

\textbf{Guidance}: +1 on one attack roll, saving throw, or skill check.

\textbf{Light}: Object shines like a torch.

\textbf{Mending}: Makes minor repairs on an object.

\textbf{Purify Food and Drink}: Purifies 1 cu. ft./level of food or water.

\textbf{Read Magic}: Read scrolls and spellbooks.

\textbf{Resistance}: Subject gains +1 on saving throws.

\textbf{Stabilize}: Cause a dying creature to stabilize.

\textbf{Virtue}: Subject gains 1 temporary hp.

\subsection{1st-Level Cleric Spells}


\textbf{Bane}: Enemies take --1 on attack rolls and saves against fear.

\textbf{Bless}: Allies gain +1 on attack rolls and saves against fear.

\textbf{Bless Water}\(^{M}\): Makes holy water.

\textbf{Cause Fear}: One creature of 5 HD or less flees for 1d4 rounds.

\textbf{Command}: One subject obeys selected command for 1 round.

\textbf{Comprehend Languages}: You understand all spoken and written languages.

\textbf{Cure Light Wounds}: Cures 1d8 damage + 1/level (max +5).

\textbf{Curse Water}\(^{M}\): Makes unholy water.

\textbf{Deathwatch}: Reveals how near death subjects within 30 ft. are.

\textbf{Detect Chaos/Evil/Good/Law}: Reveals creatures, spells, or objects of selected alignment.

\textbf{Detect Undead}: Reveals undead within 60 ft.

\textbf{Divine Favor}: You gain +1 per three levels on attack and damage rolls.

\textbf{Doom}: One subject takes --2 on attack rolls, damage rolls, saves, and checks.

\textbf{Endure Elements}: Exist comfortably in hot or cold regions.

\textbf{Entropic Shield}: Ranged attacks against you have 20\% miss chance.

\textbf{Hide from Undead}: Undead can't perceive one subject/level.

\textbf{Inflict Light Wounds}: Touch deals 1d8 damage +1/level (max +5).

\textbf{Magic Stone}: Three stones gain +1 on attack, deal 1d6 +1 damage.

\textbf{Magic Weapon}: Weapon gains +1 bonus.

\textbf{Obscuring Mist}: Fog surrounds you.

\textbf{Protection from Chaos/Evil/Good/Law}: +2 to AC and saves, plus additional protection against selected alignment.

\textbf{Remove Fear}: Suppresses fear or gives +4 on saves against fear for one subject + one per four levels.

\textbf{Sanctuary}: Opponents can't attack you, and you can't attack.

\textbf{Shield of Faith}: Aura grants +2 or higher deflection bonus.

\textbf{Summon Monster I}: Summons extraplanar creature to fight for you.

\subsection{2nd-Level Cleric Spells}


\textbf{Aid}: +1 on attack rolls and saves against fear, 1d8 temporary hp +1/level (max +10).

\textbf{Align Weapon}: Weapon becomes good, evil, lawful, or chaotic.

\textbf{Augury}\(^{MF}\): Learns whether an action will be good or bad.

\textbf{Bear's Endurance}: Subject gains +4 to Con for 1 min./level.

\textbf{Bull's Strength}: Subject gains +4 to Str for 1 min./level.

\textbf{Calm Emotions}: Calms creatures, negating emotion effects.

\textbf{Consecrate}\(^{M}\): Fills area with positive energy, weakening undead.

\textbf{Cure Moderate Wounds}: Cures 2d8 damage + 1/level (max +10).

\textbf{Darkness}: 20-ft. radius of supernatural shadow.

\textbf{Death Knell}: Kills dying creature; you gain 1d8 temporary hp, +2 to Str, and +1 caster level.

\textbf{Delay Poison}: Stops poison from harming target for 1 hour/level.

\textbf{Desecrate}\(^{M}\): Fills area with negative energy, making undead stronger.

\textbf{Eagle's Splendor}: Subject gains +4 to Cha for 1 min./level.

\textbf{Enthrall}: Captivates all within 100 ft. + 10 ft./level.

\textbf{Find Traps}: Notice traps as a rogue does.

\textbf{Gentle Repose}: Preserves one corpse.

\textbf{Hold Person}: Paralyzes one humanoid for 1 round/level.

\textbf{Inflict Moderate Wounds}: Touch attack, 2d8 damage + 1/level (max +10).

\textbf{Make Whole}: Repairs an object.

\textbf{Owl's Wisdom}: Subject gains +4 to Wis for 1 min./level.

\textbf{Remove Paralysis}: Frees creatures from paralysis or \textit{slow} effect.

\textbf{Resist Energy}: Ignores 10 (or more) points of damage/attack from specified energy type.

\textbf{Restoration, Lesser}: Dispels magical ability penalty or repairs 1d4 ability damage.

\textbf{Shatter}: Sonic vibration damages objects or crystalline creatures.

\textbf{Shield Other}\(^{F}\): You take half of subject's damage.

\textbf{Silence}: Negates sound in 20-ft. radius.

\textbf{Sound Burst}: Deals 1d8 sonic damage to subjects; may stun them.

\textbf{Spiritual Weapon}: Magic weapon attacks on its own.

\textbf{Status}: Monitors condition, position of allies.

\textbf{Summon Monster II}: Summons extraplanar creature to fight for you.

\textbf{Undetectable Alignment}: Conceals alignment for 24 hours.

\textbf{Zone of Truth}: Subjects within range cannot lie.

\subsection{3rd-Level Cleric Spells}


\textbf{Animate Dead}\(^{M}\): Creates undead skeletons and zombies.

\textbf{Bestow Curse}: --6 to an ability score; --4 on attack rolls, saves, and checks; or 50\% chance of losing each action.

\textbf{Blindness/Deafness}: Makes subject blinded or deafened.

\textbf{Contagion}: Infects subject with chosen disease.

\textbf{Continual Flame}\(^{M}\): Makes a permanent, heatless light.

\textbf{Create Food and Water}: Feeds three humans (or one horse)/level.

\textbf{Cure Serious Wounds}: Cures 3d8 damage + 1/level (max +15).

\textbf{Daylight}: 60-ft. radius of bright light.

\textbf{Deeper Darkness}: Object sheds supernatural shadow in 60-ft. radius.

\textbf{Dispel Magic}: Cancels one magical spell or effect.

\textbf{Glyph of Warding}\(^{M}\): Inscription harms those who pass it.

\textbf{Helping Hand}: Ghostly hand leads subject to you.

\textbf{Inflict Serious Wounds}: Touch attack, 3d8 damage + 1/level (max +15).

\textbf{Invisibility Purge}: Dispels invisibility within 5 ft./level.

\textbf{Locate Object}: Senses direction toward object (specific or type).

\textbf{Magic Circle against Chaos/Evil/Good/Law}: As \textit{protection }spells, but 10-ft. radius and 10 min./level.

\textbf{Magic Vestment}: Armor or shield gains +1 enhancement per four levels.

\textbf{Meld into Stone}: You and your gear merge with stone.

\textbf{Obscure Object}: Masks object against scrying.

\textbf{Prayer}: Allies get +1 bonus on most rolls, enemies --1 penalty.

\textbf{Protection from Energy}: Absorb 12 points/level of damage from one kind of energy.

\textbf{Remove Blindness/Deafness}: Cures normal or magical blindness or deafness.

\textbf{Remove Curse}: Frees object or person from curse.

\textbf{Remove Disease}: Cures all diseases affecting subject.

\textbf{Searing Light}: Ray deals 1d8/two levels damage (more against undead).

\textbf{Speak with Dead}: Corpse answers one question/two levels.

\textbf{Stone Shape}: Sculpts stone into any shape.

\textbf{Summon Monster III}: Summons extraplanar creature to fight for you.

\textbf{Water Breathing}: Subjects can breathe underwater.

\textbf{Water Walk}: Subject treads on water as if solid.

\textbf{Wind Wall}: Deflects arrows, smaller creatures, and gases.

\subsection{4th-Level Cleric Spells}


\textbf{Air Walk}: Subject treads on air as if solid (climb or descend at 45-degree angle).

\textbf{Chaos Hammer}: Harms and slows lawful creatures (1d8 damage/2 levels).

\textbf{Control Water}: Raises or lowers bodies of water.

\textbf{Cure Critical Wounds}: Cures 4d8 damage + 1/level (max +20).

\textbf{Death Ward}: Grants bonuses against death spells and negative energy.

\textbf{Dimensional Anchor}: Bars extradimensional movement.

\textbf{Discern Lies}: Reveals deliberate falsehoods.

\textbf{Dismissal}: Forces a creature to return to native plane.

\textbf{Divination}\(^{M}\): Provides useful advice for specific proposed actions.

\textbf{Divine Power}: You gain attack bonuses and 1 hp/level.

\textbf{Freedom of Movement}: Subject moves normally despite impediments to movement.

\textbf{Giant Vermin}: Turns centipedes, scorpions, or spiders into giant vermin.

\textbf{Holy Smite}: Harms and possibly blinds evil creatures (1d8 damage/2 levels).

\textbf{Imbue with Spell Ability}: Transfer spells to subject.

\textbf{Inflict Critical Wounds}: Touch attack, 4d8 damage + 1/level (max +20).

\textbf{Magic Weapon, Greater}: Weapon gains +1 bonus/four levels (max +5).

\textbf{Neutralize Poison}: Immunizes subject against poison, detoxifies venom in or on subject.

\textbf{Order's Wrath}: Harms and dazes chaotic creatures (1d8 damage/2 levels).

\textbf{Planar Ally, Lesser}\(^{M}\): Exchange services with a 6 HD extraplanar creature.

\textbf{Poison}: Touch deals 1d3 Con damage 1/round for 6 rounds.

\textbf{Repel Vermin}: Insects, spiders, and other vermin stay 10 ft. away.

\textbf{Restoration}\(^{M}\): Restores level and ability score drains.

\textbf{Sending}: Delivers short message anywhere, instantly.

\textbf{Spell Immunity}: Subject is immune to one spell per 4 levels.

\textbf{Summon Monster IV}: Summons extraplanar creature to fight for you.

\textbf{Tongues}: Speak and understand any language.

\textbf{Unholy Blight}: Harms and sickens good creatures (1d8 damage/2 levels).

\subsection{5th-Level Cleric Spells}


\textbf{Atonement}\(^{FM}\): Removes burden of misdeeds from subject and reverses magical alignment change.

\textbf{Break Enchantment}: Frees subjects from enchantments, transmutations, and curses.

\textbf{Breath of Life}: Cures 5d8 damage + 1/level and restores life to recently slain creatures.

\textbf{Command, Greater}: As \textit{command, }but affects one subject/level.

\textbf{Commune}\(^{M}\): Deity answers one yes-or-no question/level.

\textbf{Cure Light Wounds, Mass}: Cures 1d8 damage + 1/level, affects 1 subject/level.

\textbf{Dispel Chaos/Evil/Good/Law}: +4 bonus against attacks.

\textbf{Disrupting Weapon}: Melee weapon destroys undead.

\textbf{Flame Strike}: Smites foes with divine fire (1d6/level damage).

\textbf{Hallow}\(^{M}\): Designates location as holy.

\textbf{Inflict Light Wounds, Mass}: Deals 1d8 damage + 1/level, affects 1 subject/level.

\textbf{Insect Plague}: Wasp swarms attack creatures.

\textbf{Mark of Justice}: Designates action that triggers \textit{curse }on subject.

\textbf{Plane Shift}\(^{F}\): As many as 8 subjects travel to another plane.

\textbf{Raise Dead}\(^{M}\): Restores life to subject who died as long as one day/level ago.

\textbf{Righteous Might}: Your size increases, and you gain bonuses in combat.

\textbf{Scrying}\(^{F}\): Spies on subject from a distance.

\textbf{Slay Living}: Touch attack deals 12d6 + 1 per level.

\textbf{Spell Resistance}: Subject gains SR 12 + level.

\textbf{Summon Monster V}: Summons extraplanar creature to fight for you.

\textbf{Symbol of Pain}\(^{M}\): Triggered rune wracks nearby creatures with pain.

\textbf{Symbol of Sleep}\(^{M}\): Triggered rune puts nearby creatures into catatonic slumber.

\textbf{True Seeing}\(^{M}\): Lets you see all things as they really are.

\textbf{Unhallow}\(^{M}\): Designates location as unholy.

\textbf{Wall of Stone}: Creates a stone wall that can be shaped.

\subsection{6th-Level Cleric Spells}


\textbf{Animate Objects}: Objects attack your foes.

\textbf{Antilife Shell}: 10-ft.-radius field hedges out living creatures.

\textbf{Banishment}: Banishes 2 HD/level of extraplanar creatures.

\textbf{Bear's Endurance, Mass}: As \textit{bear's endurance}, affects 1 subject/level.

\textbf{Blade Barrier}: Wall of blades deals 1d6/level damage.

\textbf{Bull's Strength, Mass}: As \textit{bull's strength}, affects 1 subject/level.

\textbf{Create Undead}\(^{M}\): Create ghasts, ghouls, mohrgs, or mummies.

\textbf{Cure Moderate Wounds, Mass}: Cures 2d8 damage + 1/level, affects 1 subject/level.

\textbf{Dispel Magic, Greater}: As \textit{dispel magic}, but with multiple targets.

\textbf{Eagle's Splendor, Mass}: As \textit{eagle's splendor}, affects 1 subject/level.

\textbf{Find the Path}: Shows most direct way to a location.

\textbf{Forbiddance}\(^{M}\): Blocks planar travel, damages creatures of different alignment.

\textbf{Geas/Quest}: As \textit{lesser geas, }but affects any creature.

\textbf{Glyph of Warding, Greater}\(^{M}\): As \textit{glyph of warding}, but up to 10d8 damage or 6th-level spell.

\textbf{Harm}: Deals 10 points/level damage to target.

\textbf{Heal}: Cures 10 points/level damage, all diseases and mental conditions.

\textbf{Heroes' Feast}: Food for one creature/level cures and grants combat bonuses\textit{.}

\textbf{Inflict Moderate Wounds, Mass}: Deals 2d8 damage + 1/level, affects 1 subject/level.

\textbf{Owl's Wisdom, Mass}: As \textit{owl's wisdom}, affects 1 subject/level.

\textbf{Planar Ally}\(^{M}\): As \textit{lesser planar ally, }but up to 12 HD.

\textbf{Summon Monster VI}: Summons extraplanar creature to fight for you.

\textbf{Symbol of Fear}\(^{M}\): Triggered rune panics nearby creatures.

\textbf{Symbol of Persuasion}\(^{M}\): Triggered rune charms nearby creatures.

\textbf{Undeath to Death}\(^{M}\): Destroys 1d4 HD/level undead (max. 20d4).

\textbf{Wind Walk}: You and your allies turn vaporous and travel fast.

\textbf{Word of Recall}: Teleports you back to designated place.

\subsection{7th-Level Cleric Spells}


\textbf{Blasphemy}: Kills, paralyzes, weakens, or dazes nonevil subjects.

\textbf{Control Weather}: Changes weather in local area.

\textbf{Cure Serious Wounds, Mass}: Cures 3d8 damage + 1/level, affects 1 subject/level.

\textbf{Destruction}\(^{F}\): Kills subject and destroys remains.

\textbf{Dictum}: Kills, paralyzes, staggers, or deafens nonlawful targets.

\textbf{Ethereal Jaunt}: You become ethereal for 1 round/level.

\textbf{Holy Word}: Kills, paralyzes, blinds, or deafens nongood subjects.

\textbf{Inflict Serious Wounds, Mass}: Deals 3d8 damage + 1/level, affects 1 subject/level.

\textbf{Refuge}\(^{M}\): Alters item to transport its possessor to you.

\textbf{Regenerate}: Subject's severed limbs grow back, cures 4d8 damage +1/level (max +35).

\textbf{Repulsion}: Creatures can't approach you.

\textbf{Restoration, Greater}\(^{M}\): As \textit{restoration, }plus restores all levels and ability scores.

\textbf{Resurrection}\(^{M}\): Fully restore dead subject.

\textbf{Scrying, Greater}: As \textit{scrying}, but faster and longer.

\textbf{Summon Monster VII}: Summons extraplanar creature to fight for you.

\textbf{Symbol of Stunning}\(^{M}\): Triggered rune stuns nearby creatures.

\textbf{Symbol of Weakness}\(^{M}\): Triggered rune weakens nearby creatures.

\textbf{Word of Chaos}: Kills, confuses\textit{, }stuns, or deafens nonchaotic subjects.

\subsection{8th-Level Cleric Spells}


\textbf{Antimagic Field}: Negates magic within 10 ft.

\textbf{Cloak of Chaos}\(^{F}\): +4 to AC, +4 resistance, and SR 25 against lawful spells.

\textbf{Create Greater Undead}\(^{M}\): Create shadows, wraiths, spectres, or devourers.

\textbf{Cure Critical Wounds, Mass}: Cures 4d8 damage + 1/level for many creatures.

\textbf{Dimensional Lock}: Teleportation and interplanar travel blocked for 1 day/level.

\textbf{Discern Location}: Reveals exact location of creature or object.

\textbf{Earthquake}: Intense tremor shakes 80-ft. radius.

\textbf{Fire Storm}: Deals 1d6/level fire damage.

\textbf{Holy Aura}\(^{F}\): +4 to AC, +4 resistance, and SR 25 against evil spells.

\textbf{Inflict Critical Wounds, Mass}: Deals 4d8 damage + 1/level, affects 1 subject/level.

\textbf{Planar Ally, Greater}\(^{M}\): As \textit{lesser planar ally, }but up to 18 HD.

\textbf{Shield of Law}\(^{F}\): +4 to AC, +4 resistance, and SR 25 against chaotic spells.

\textbf{Spell Immunity, Greater}: As \textit{spell immunity, }but up to 8th-level spells.

\textbf{Summon Monster VIII}: Summons extraplanar creature to fight for you.

\textbf{Symbol of Death}\(^{M}\): Triggered rune kills nearby creatures.

\textbf{Symbol of Insanity}\(^{M}\): Triggered rune renders nearby creatures insane.

\textbf{Unholy Aura}\(^{F}\): +4 to AC, +4 resistance, and SR 25 against good spells.

\subsection{9th-Level Cleric Spells}


\textbf{Astral Projection}\(^{M}\): Projects you and others onto Astral Plane.

\textbf{Energy Drain}: Subject gains 2d4 negative levels.

\textbf{Etherealness}: Travel to Ethereal Plane with companions.

\textbf{Gate}\(^{M}\): Connects two planes for travel or summoning.

\textbf{Heal, Mass}: As \textit{heal}, but affects 1 subject/level.

\textbf{Implosion}: Inflict 10 damage/level to one creature/round.

\textbf{Miracle}\(^{M}\): Requests a deity's intercession.

\textbf{Soul Bind}\(^{F}\): Traps newly dead soul to prevent \textit{resurrection}.

\textbf{Storm of Vengeance}: Storm rains acid, lightning, and hail.

\textbf{Summon Monster IX}: Summons extraplanar creature to fight for you.

\textbf{True Resurrection}\(^{M}\): As \textit{resurrection}, plus remains aren't needed.

\section{Druid Spells}


\subsection{0-Level Druid Spells (Orisons)}


\textbf{Create Water}: Creates 2 gallons/level of pure water.

\textbf{Detect Magic}: Detects spells and magic items within 60 ft.

\textbf{Detect Poison}: Detects poison in one creature or object.

\textbf{Flare}: Dazzles one creature (--1 penalty on attack rolls).

\textbf{Guidance}: +1 on one attack roll, saving throw, or skill check.

\textbf{Know Direction}: You discern north.

\textbf{Light}: Object shines like a torch.

\textbf{Mending}: Makes minor repairs on an object.

\textbf{Purify Food and Drink}: Purifies 1 cu. ft./level of food or water.

\textbf{Read Magic}: Read scrolls and spellbooks.

\textbf{Resistance}: Subject gains +1 bonus on saving throws.

\textbf{Stabilize}: Cause a dying creature to stabilize.

\textbf{Virtue}: Subject gains 1 temporary hp.

\subsection{1st-Level Druid Spells}


\textbf{Calm Animals}: Calms 2d4 + level HD of animals.

\textbf{Charm Animal}: Makes one animal your friend.

\textbf{Cure Light Wounds}: Cures 1d8 damage + 1/level (max +5).

\textbf{Detect Animals or Plants}: Detects kinds of animals or plants.

\textbf{Detect Snares and Pits}: Reveals natural or primitive traps.

\textbf{Endure Elements}: Exist comfortably in hot or cold regions.

\textbf{Entangle}: Plants entangle everyone in 40-ft. radius.

\textbf{Faerie Fire}: Outlines subjects with light, canceling \textit{blur}, concealment, and the like.

\textbf{Goodberry}: 2d4 berries each cure 1 hp (max 8 hp/24 hours).

\textbf{Hide from Animals}: Animals can't perceive one subject/level.

\textbf{Jump}: Subject gets bonus on Acrobatics checks.

\textbf{Longstrider}: Your speed increases by 10 ft.

\textbf{Magic Fang}: One natural weapon of subject creature gets +1 on attack and damage rolls.

\textbf{Magic Stone}: Three stones gain +1 on attack rolls, deal 1d6+1 damage.

\textbf{Obscuring Mist}: Fog surrounds you.

\textbf{Pass without Trace}: One subject/level leaves no tracks.

\textbf{Produce Flame}: 1d6 damage + 1/level, touch or thrown.

\textbf{Shillelagh}: Cudgel or quarterstaff becomes +1 weapon (2d6 damage) for 1 min./level.

\textbf{Speak with Animals}: You can communicate with animals.

\textbf{Summon Nature's Ally I}: Summons creature to fight.

\subsection{2nd-Level Druid Spells}


\textbf{Animal Messenger}: Sends a Tiny animal to a specific place.

\textbf{Animal Trance}: Fascinates 2d6 HD of animals.

\textbf{Barkskin}: Grants +2 (or higher) enhancement to natural armor.

\textbf{Bear's Endurance}: Subject gains +4 to Con for 1 min./level.

\textbf{Bull's Strength}: Subject gains +4 to Str for 1 min./level.

\textbf{Cat's Grace}: Subject gains +4 to Dex for 1 min./level.

\textbf{Chill Metal}: Cold metal damages those who touch it.

\textbf{Delay Poison}: Stops poison from harming subject for 1 hour/level.

\textbf{Fire Trap}\(^{M}\): Opened object deals 1d4 + 1/level damage.

\textbf{Flame Blade}: Touch attack deals 1d8 + 1/two levels damage.

\textbf{Flaming Sphere}: Rolling ball of fire deals 3d6 fire damage.

\textbf{Fog Cloud}: Fog obscures vision.

\textbf{Gust of Wind}: Blows away or knocks down smaller creatures.

\textbf{Heat Metal}: Makes metal so hot it damages those who touch it.

\textbf{Hold Animal}: Paralyzes one animal for 1 round/level.

\textbf{Owl's Wisdom}: Subject gains +4 to Wis for 1 min./level.

\textbf{Reduce Animal}: Shrinks one willing animal.

\textbf{Resist Energy}: Ignores 10 or more points of damage per attack from specified energy type.

\textbf{Restoration, Lesser}: Dispels magical ability penalty or repairs 1d4 ability damage.

\textbf{Soften Earth and Stone}: Turns stone to clay, or dirt to sand or mud.

\textbf{Spider Climb}: Grants ability to walk on walls and ceilings.

\textbf{Summon Nature's Ally II}: Summons creature to fight.

\textbf{Summon Swarm}: Summons swarm of bats, rats, or spiders.

\textbf{Tree Shape}: You look exactly like a tree for 1 hour/level.

\textbf{Warp Wood}: Bends wood.

\textbf{Wood Shape}: Reshapes wooden objects to suit you.

\subsection{3rd-Level Druid Spells}


\textbf{Call Lightning}: Calls down lightning bolts (3d6 per bolt) from sky.

\textbf{Contagion}: Infects subject with chosen disease.

\textbf{Cure Moderate Wounds}: Cures 2d8 damage + 1/level (max +10).

\textbf{Daylight}: 60-ft. radius of bright light.

\textbf{Diminish Plants}: Reduces size or blights the growth of normal plants.

\textbf{Dominate Animal}: One animal obeys your silent mental commands and orders.

\textbf{Magic Fang, Greater}: One natural weapon gets + 1/four levels (max +5).

\textbf{Meld into Stone}: You and your gear merge with stone.

\textbf{Neutralize Poison}: Immunizes subject against poison, detoxifies venom in or on subject.

\textbf{Plant Growth}: Grows vegetation, improves crops.

\textbf{Poison}: Touch deals 1d3 Con damage 1/round for 6 rounds.

\textbf{Protection from Energy}: Absorbs 12 points/level of damage from one kind of energy.

\textbf{Quench}: Extinguishes fires.

\textbf{Remove Disease}: Cures all diseases affecting subject.

\textbf{Sleet Storm}: Hampers vision and movement.

\textbf{Snare}: Creates a magic booby trap.

\textbf{Speak with Plants}: You can talk to plants and plant creatures.

\textbf{Spike Growth}: Creatures in area take 1d4 damage, may be slowed\textit{.}

\textbf{Stone Shape}: Sculpts stone into any shape.

\textbf{Summon Nature's Ally III}: Summons creature to fight.

\textbf{Water Breathing}: Subjects can breathe underwater.

\textbf{Wind Wall}: Deflects arrows, smaller creatures, and gases.

\subsection{4th-Level Druid Spells}


\textbf{Air Walk}: Subject treads on air as if solid (climb or descend at 45-degree angle).

\textbf{Antiplant Shell}: Keeps animated plants at bay.

\textbf{Blight}: Withers one plant or deals 1d6/level damage to plant creature.

\textbf{Command Plants}: Sways the actions of plant creatures.

\textbf{Control Water}: Raises or lowers bodies of water.

\textbf{Cure Serious Wounds}: Cures 3d8 damage + 1/level (max +15).

\textbf{Dispel Magic}: Cancels one magical spell or effect.

\textbf{Flame Strike}: Smites foes with divine fire (1d6/level damage).

\textbf{Freedom of Movement}: Subject moves normally despite impediments to movement.

\textbf{Giant Vermin}: Turns centipedes, scorpions, or spiders into giant vermin.

\textbf{Ice Storm}: Hail deals 5d6 damage in cylinder 40 ft. across.

\textbf{Reincarnate}: Brings dead subject back in a random body.

\textbf{Repel Vermin}: Insects, spiders, and other vermin stay 10 ft. away.

\textbf{Rusting Grasp}: Your touch corrodes iron and alloys.

\textbf{Scrying}\(^{F}\): Spies on subject from a distance.

\textbf{Spike Stones}: Creatures in area take 1d8 damage, may also be slowed\textit{.}

\textbf{Summon Nature's Ally IV}: Summons creature to fight.

\subsection{5th-Level Druid Spells}


\textbf{Animal Growth}: One animal doubles in size.

\textbf{Atonement}\(^{FM}\): Removes burden of misdeeds from subject.

\textbf{Awaken}\(^{M}\): Animal or tree gains human intellect.

\textbf{Baleful Polymorph}: Transforms subject into harmless animal.

\textbf{Call Lightning Storm}: As \textit{call lightning, }but 5d6 damage per bolt.

\textbf{Commune with Nature}: Learn about terrain for 1 mile/level.

\textbf{Control Winds}: Changes wind direction and speed.

\textbf{Cure Critical Wounds}: Cures 4d8 damage + 1/level (max +20).

\textbf{Death Ward}: Grants bonuses against death spells and negative energy.

\textbf{Hallow}\(^{M}\): Designates location as holy.

\textbf{Insect Plague}: Wasp swarms attack creatures.

\textbf{Stoneskin}\(^{M}\): Grants DR 10/adamantine.

\textbf{Summon Nature's Ally V}: Summons creature to fight.

\textbf{Transmute Mud to Rock}: Transforms two 10-ft. cubes per level.

\textbf{Transmute Rock to Mud}: Transforms two 10-ft. cubes per level.

\textbf{Tree Stride}: Step from one tree to another far away.

\textbf{Unhallow}\(^{M}\): Designates location as unholy.

\textbf{Wall of Fire}: Deals 2d4 fire damage out to 10 ft. and 1d4 out to 20 ft. Passing through wall deals 2d6 damage +1/level.

\textbf{Wall of Thorns}: Thorns damage anyone who tries to pass.

\subsection{6th-Level Druid Spells}


\textbf{Antilife Shell}: 10-ft.-radius field hedges out living creatures.

\textbf{Bear's Endurance, Mass}: As \textit{bear's endurance}, affects 1 subject/level.

\textbf{Bull's Strength, Mass}: As \textit{bull's strength}, affects 1 subject/level.

\textbf{Cat's Grace, Mass}: As \textit{cat's grace}, affects one subject/level.

\textbf{Cure Light Wounds, Mass}: Cures 1d8 damage + 1/level, affects 1 subject/level.

\textbf{Dispel Magic, Greater}: As \textit{dispel magic}, but with multiple targets.

\textbf{Find the Path}: Shows most direct way to a location.

\textbf{Fire Seeds}: Acorns and berries become grenades and bombs.

\textbf{Ironwood}: Magic wood is as strong as steel.

\textbf{Liveoak}: Oak becomes treant guardian.

\textbf{Move Earth}: Digs trenches and builds hills.

\textbf{Owl's Wisdom, Mass}: As \textit{owl's wisdom}, affects 1 subject/level.

\textbf{Repel Wood}: Pushes away wooden objects.

\textbf{Spellstaff}: Stores one spell in wooden quarterstaff.

\textbf{Stone Tell}: Talk to natural or worked stone.

\textbf{Summon Nature's Ally VI}: Summons creature to fight.

\textbf{Transport via Plants}: Move instantly from one plant to another of the same kind.

\textbf{Wall of Stone}: Creates a stone wall that can be shaped.

\subsection{7th-Level Druid Spells}


\textbf{Animate Plants}: One or more plants animate and fight for you.

\textbf{Changestaff}: Your staff becomes a treant on command.

\textbf{Control Weather}: Changes weather in local area.

\textbf{Creeping Doom}: Swarms of centipedes attack at your command.

\textbf{Cure Moderate Wounds, Mass}: Cures 2d8 damage + 1/level, affects 1 subject/level.

\textbf{Fire Storm}: Deals 1d6/level fire damage.

\textbf{Heal}: Cures 10 points/level damage, all diseases and mental conditions.

\textbf{Scrying, Greater}: As \textit{scrying}, but faster and longer.

\textbf{Summon Nature's Ally VII}: Summons creature to fight.

\textbf{Sunbeam}: Beam blinds and deals 4d6 damage.

\textbf{Transmute Metal to Wood}: Metal within 40 ft. becomes wood.

\textbf{True Seeing}\(^{M}\): Lets you see all things as they really are.

\textbf{Wind Walk}: You and your allies turn vaporous and travel fast.

\subsection{8th-Level Druid Spells}


\textbf{Animal Shapes}: One ally/level polymorphs into chosen animal.

\textbf{Control Plants}: Controls actions of one or more plant creatures.

\textbf{Cure Serious Wounds, Mass}: Cures 3d8 damage + 1/level, affects 1 subject/level.

\textbf{Earthquake}: Intense tremor shakes 80-ft.-radius.

\textbf{Finger of Death}: Deals 10 damage/level to one subject.

\textbf{Repel Metal or Stone}: Pushes away metal and stone.

\textbf{Reverse Gravity}: Objects and creatures fall upward.

\textbf{Summon Nature's Ally VIII}: Summons creature to fight.

\textbf{Sunburst}: Blinds all within 10 ft., deals 6d6 damage.

\textbf{Whirlwind}: Cyclone deals damage and can pick up creatures.

\textbf{Word of Recall}: Teleports you back to designated place.

\subsection{9th-Level Druid Spells}


\textbf{Antipathy}: Object or location affected by spell repels certain creatures.

\textbf{Cure Critical Wounds, Mass}: Cures 4d8 damage + 1/level for many creatures.

\textbf{Elemental Swarm}: Summons multiple elementals.

\textbf{Foresight}: \texttt{{}"{}}Sixth sense\texttt{{}"{}} warns of impending danger.

\textbf{Regenerate}: Subject's severed limbs grow back, cures 4d8 damage +1/level (max +35).

\textbf{Shambler}: Creates 1d4+2 shambling mounds to fight for you.

\textbf{Shapechange}\(^{F}\): Transforms you into certain creatures, and you can change forms once per round.

\textbf{Storm of Vengeance}: Storm rains acid, lightning, and hail.

\textbf{Summon Nature's Ally IX}: Summons creature to fight.

\textbf{Sympathy}\(^{M}\): Object or location attracts certain creatures.

\section{Paladin Spells}


\subsection{1st-Level Paladin Spells}


\textbf{Bless}: Allies gain +1 on attack rolls and saves against fear.

\textbf{Bless Water}\(^{M}\): Makes holy water.

\textbf{Bless Weapon}: Weapon strikes true against evil foes.

\textbf{Create Water}: Creates 2 gallons/level of pure water.

\textbf{Cure Light Wounds}: Cures 1d8 damage + 1/level (max +5).

\textbf{Detect Poison}: Detects poison in one creature or small object.

\textbf{Detect Undead}: Reveals undead within 60 ft.

\textbf{Divine Favor}: You gain +1 per three levels on attack and damage rolls.

\textbf{Endure Elements}: Exist comfortably in hot or cold regions.

\textbf{Magic Weapon}: Weapon gains +1 bonus.

\textbf{Protection from Chaos/Evil}: +2 to AC and saves, plus additional protection against selected alignment.

\textbf{Read Magic}: Read scrolls and spellbooks.

\textbf{Resistance}: Subject gains +1 on saving throws.

\textbf{Restoration, Lesser}: Dispels magical ability penalty or repairs 1d4 ability damage.

\textbf{Virtue}: Subject gains 1 temporary hp.

\subsection{2nd-Level Paladin Spells}


\textbf{Bull's Strength}: Subject gains +4 to Str for 1 min./level.

\textbf{Delay Poison}: Stops poison from harming subject for 1 hour/level.

\textbf{Eagle's Splendor}: Subject gains +4 to Cha for 1 min./level.

\textbf{Owl's Wisdom}: Subject gains +4 to Wis for 1 min./level.

\textbf{Remove Paralysis}: Frees one or more creatures from paralysis or slow effect.

\textbf{Resist Energy}: Ignores 10 or more points of damage per attack from specified energy type.

\textbf{Shield Other}\(^{F}\): You take half of subject's damage.

\textbf{Undetectable Alignment}: Conceals alignment from magical detection for 24 hours.

\textbf{Zone of Truth}: Subjects within range cannot lie.

\subsection{3rd-Level Paladin Spells}


\textbf{Cure Moderate Wounds}: Cures 2d8 damage +1/level (max. +10).

\textbf{Daylight}: 60-ft. radius of bright light.

\textbf{Discern Lies}: Reveals deliberate falsehoods.

\textbf{Dispel Magic}: Cancels one magical spell or effect.

\textbf{Heal Mount}: As \textit{heal }on horse or other special mount.

\textbf{Magic Circle against Chaos/Evil}: As \textit{protection from chaos}, but 10-ft. radius and 10 min./level.

\textbf{Magic Weapon, Greater}: Weapon gains +1 bonus/four levels (max +5).

\textbf{Prayer}: Allies get +1 bonus on most rolls, enemies --1 penalty.

\textbf{Remove Blindness/Deafness}: Cures normal or magical blindness or deafness.

\textbf{Remove Curse}: Frees object or person from curse.

\subsection{4th-Level Paladin Spells}


\textbf{Break Enchantment}: Frees subjects from enchantments, transmutations, and curses.

\textbf{Cure Serious Wounds}: Cures 3d8 damage + 1/level (max +15).

\textbf{Death Ward}: Grants bonuses against death spells and negative energy.

\textbf{Dispel Chaos}: +4 bonus against attacks by chaotic creatures.

\textbf{Dispel Evil}: +4 bonus against attacks by evil creatures.

\textbf{Holy Sword}: Weapon becomes +5, deals +2d6 damage against evil.

\textbf{Mark of Justice}: Designates action that triggers \textit{curse }on subject.

\textbf{Neutralize Poison}: Immunizes subject against poison, detoxifies venom in or on subject.

\textbf{Restoration}\(^{M}\): Restores level and ability score drains.

\section{Ranger Spells}


\subsection{1st-Level Ranger Spells}


\textbf{Alarm}: Wards an area for 2 hours/level.

\textbf{Animal Messenger}: Sends a Tiny animal to a specific place.

\textbf{Calm Animals}: Calms 2d4 + level HD of animals.

\textbf{Charm Animal}: Makes one animal your friend.

\textbf{Delay Poison}: Stops poison from harming subject for 1 hour/level.

\textbf{Detect Animals or Plants}: Detects kinds of animals or plants.

\textbf{Detect Poison}: Detects poison in one creature or object.

\textbf{Detect Snares and Pits}: Reveals natural or primitive traps.

\textbf{Endure Elements}: Exist comfortably in hot or cold regions.

\textbf{Entangle}: Plants entangle everyone in 40-ft. radius.

\textbf{Hide from Animals}: Animals can't perceive one subject/level.

\textbf{Jump}: Subject gets bonus on Acrobatics checks.

\textbf{Longstrider}: Your base speed increases by 10 ft.

\textbf{Magic Fang}: One natural weapon of subject creature gets +1 on attack and damage rolls.

\textbf{Pass without Trace}: One subject/level leaves no tracks.

\textbf{Read Magic}: Read scrolls and spellbooks.

\textbf{Resist Energy}: Ignores 10 (or more) points of damage/attack from specified energy type.

\textbf{Speak with Animals}: You can communicate with animals.

\textbf{Summon Nature's Ally I}: Summons creature to fight for you.

\subsection{2nd-Level Ranger Spells}


\textbf{Barkskin}: Grants +2 (or higher) enhancement to natural armor.

\textbf{Bear's Endurance}: Subject gains +4 to Con for 1 min./level.

\textbf{Cat's Grace}: Subject gains +4 to Dex for 1 min./level.

\textbf{Cure Light Wounds}: Cures 1d8 damage + 1/level (max +5).

\textbf{Hold Animal}: Paralyzes one animal for 1 round/level.

\textbf{Owl's Wisdom}: Subject gains +4 to Wis for 1 min./level.

\textbf{Protection from Energy}: Absorbs 12 points/level of damage from one kind of energy.

\textbf{Snare}: Creates a magic booby trap.

\textbf{Speak with Plants}: You can talk to plants and plant creatures.

\textbf{Spike Growth}: Creatures in area take 1d4 damage, may be slowed\textit{.}

\textbf{Summon Nature's Ally II}: Summons creature to fight for you.

\textbf{Wind Wall}: Deflects arrows, smaller creatures, and gases.

\subsection{3rd-Level Ranger Spells}


\textbf{Command Plants}: Sway the actions of plant creatures.

\textbf{Cure Moderate Wounds}: Cures 2d8 damage +1/level (max. +10).

\textbf{Darkvision}: See 60 ft. in total darkness.

\textbf{Diminish Plants}: Reduces size or blights growth of normal plants.

\textbf{Magic Fang, Greater}: One natural weapon gets + 1/four levels (max. +5).

\textbf{Neutralize Poison}: Immunizes subject against poison, detoxifies venom in or on subject.

\textbf{Plant Growth}: Grows vegetation, improves crops.

\textbf{Reduce Animal}: Shrinks one willing animal.

\textbf{Remove Disease}: Cures all diseases affecting subject.

\textbf{Repel Vermin}: Insects, spiders, and other vermin stay 10 ft. away.

\textbf{Summon Nature's Ally III}: Summons creature to fight for you.

\textbf{Tree Shape}: You look exactly like a tree for 1 hour/level.

\textbf{Water Walk}: Subject treads on water as if solid.

\subsection{4th-Level Ranger Spells}


\textbf{Animal Growth}: One animal doubles in size.

\textbf{Commune with Nature}: Learn about terrain for 1 mile/level.

\textbf{Cure Serious Wounds}: Cures 3d8 damage + 1/level (max +15).

\textbf{Freedom of Movement}: Subject moves normally despite impediments to movement.

\textbf{Nondetection}\(^{M}\): Hides subject from divination, scrying.

\textbf{Summon Nature's Ally IV}: Summons creature to fight for you.

\textbf{Tree Stride}: Step from one tree to another far away.

\section{Sorcerer/wizard Spells}


\subsection{0-Level Sorcerer/Wizard Spells (Cantrips)}


\subsection{Abjuration}


\textbf{Resistance}: Subject gains +1 on saving throws.

\subsection{Conjuration}


\textbf{Acid Splash}: Orb deals 1d3 acid damage.

\subsection{Divination}


\textbf{Detect Magic}: Detects all spells and magic items within 60 ft.

\textbf{Detect Poison}: Detects poison in one creature or small object.

\textbf{Read Magic}: Read scrolls and spellbooks.

\subsection{Enchantment}


\textbf{Daze}: A single humanoid creature with 4 HD or less loses its next action.

\subsection{Evocation}


\textbf{Dancing Lights}: Creates torches or other lights.

\textbf{Flare}: Dazzles one creature (--1 on attack rolls).

\textbf{Light}: Object shines like a torch.

\textbf{Ray of Frost}: Ray deals 1d3 cold damage.

\subsection{Illusion}


\textbf{Ghost Sound}: Figment sounds.

\subsection{Necromancy}


\textbf{Bleed}: Cause a stabilized creature to resume dying.

\textbf{Disrupt Undead}: Deals 1d6 damage to one undead.

\textbf{Touch of Fatigue}: Touch attack fatigues target.

\subsection{Transmutation}


\textbf{Mage Hand}: 5-pound telekinesis.

\textbf{Mending}: Makes minor repairs on an object.

\textbf{Message}: Whisper conversation at distance.

\textbf{Open/Close}: Opens or closes small or light things.

\subsection{Universal}


\textbf{Arcane Mark}: Inscribes a personal rune on an object or creature (visible or invisible).

\textbf{Prestidigitation}: Performs minor tricks.

\subsection{1st-Level Sorcerer/Wizard Spells}


\subsection{Abjuration}


\textbf{Alarm}: Wards an area for 2 hours/level.

\textbf{Endure Elements}: Exist comfortably in hot or cold regions.

\textbf{Hold Portal}: Holds door shut.

\textbf{Protection from Chaos/Evil/Good/Law}: +2 to AC and saves, plus additional protection against selected alignment.

\textbf{Shield}: Invisible disc gives +4 to AC, blocks \textit{magic missiles.}

\subsection{Conjuration}


\textbf{Grease}: Makes 10-ft. square or one object slippery.

\textbf{Mage Armor}: Gives subject +4 armor bonus.

\textbf{Mount}: Summons riding horse for 2 hours/level.

\textbf{Obscuring Mist}: Fog surrounds you.

\textbf{Summon Monster I}: Summons extraplanar creature to fight for you.

\textbf{Unseen Servant}: Invisible force obeys your commands.

\subsection{Divination}


\textbf{Comprehend Languages}: You understand all spoken and written languages.

\textbf{Detect Secret Doors}: Reveals hidden doors within 60 ft.

\textbf{Detect Undead}: Reveals undead within 60 ft.

\textbf{Identify}: Gives +10 bonus to identify magic items.

\textbf{True Strike}: +20 on your next attack roll.

\subsection{Enchantment}


\textbf{Charm Person}: Makes one person your friend.

\textbf{Hypnotism}: Fascinates 2d4 HD of creatures.

\textbf{Sleep}: Puts 4 HD of creatures into magical slumber.

\subsection{Evocation}


\textbf{Burning Hands}: 1d4/level fire damage (max 5d4).

\textbf{Floating Disk}: Creates 3-ft.-diameter horizontal disk that holds 100 lbs./level.

\textbf{Magic Missile}: 1d4+1 damage; +1 missile per two levels above 1st (max 5).

\textbf{Shocking Grasp}: Touch delivers 1d6/level electricity damage (max 5d6).

\subsection{Illusion}


\textbf{Color Spray}: Knocks unconscious, blinds, and/or stuns weak creatures.

\textbf{Disguise Self}: Changes your appearance.

\textbf{Magic Aura}: Alters object's magic aura.

\textbf{Silent Image}: Creates minor illusion of your design.

\textbf{Ventriloquism}: Throws voice for 1 min./level.

\subsection{Necromancy}


\textbf{Cause Fear}: One creature of 5 HD or less flees for 1d4 rounds.

\textbf{Chill Touch}: One touch/level deals 1d6 damage and possibly 1 Str damage.

\textbf{Ray of Enfeeblement}: Ray causes 1d6 Str penalty + 1 per 2 levels.

\subsection{Transmutation}


\textbf{Animate Rope}: Makes a rope move at your command.

\textbf{Enlarge Person}: Humanoid creature doubles in size.

\textbf{Erase}: Mundane or magical writing vanishes.

\textbf{Expeditious Retreat}: Your base speed increases by 30 ft.

\textbf{Feather Fall}: Objects or creatures fall slowly.

\textbf{Jump}: Subject gets bonus on Acrobatics checks.

\textbf{Magic Weapon}: Weapon gains +1 bonus.

\textbf{Reduce Person}: Humanoid creature halves in size.

\subsection{2nd-Level Sorcerer/Wizard Spells}


\subsection{Abjuration}


\textbf{Arcane Lock}\(^{M}\): Magically locks a portal or chest.

\textbf{Obscure Object}: Masks object against scrying.

\textbf{Protection from Arrows}: Subject gains DR 10/magic against ranged attacks.

\textbf{Resist Energy}: Ignores first 10 (or more) points of damage per attack from specified energy type.

\subsection{Conjuration}


\textbf{Acid Arrow}: Ranged touch attack; 2d4 damage for 1 round + 1 round/three levels.

\textbf{Fog Cloud}: Fog obscures vision.

\textbf{Glitterdust}: Blinds creatures, outlines invisible creatures.

\textbf{Summon Monster II}: Summons extraplanar creature to fight for you.

\textbf{Summon Swarm}: Summons swarm of bats, rats, or spiders.

\textbf{Web}: Fills 20-ft.-radius spread with sticky spiderwebs that can grapple foes and impair movement.

\subsection{Divination}


\textbf{Detect Thoughts}: Allows \texttt{{}"{}}listening\texttt{{}"{}} to surface thoughts.

\textbf{Locate Object}: Senses direction toward object (specific or type).

\textbf{See Invisibility}: Reveals invisible creatures or objects.

\subsection{Enchantment}


\textbf{Daze Monster}: Living creature of 6 HD or less loses its next action.

\textbf{Hideous Laughter}: Subject loses actions for 1 round/level.

\textbf{Touch of Idiocy}: Subject takes 1d6 penalty to Int, Wis, and Cha.

\subsection{Evocation}


\textbf{Continual Flame}\(^{M}\): Makes a permanent, heatless light.

\textbf{Darkness}: 20-ft. radius of supernatural shadow.

\textbf{Flaming Sphere}: Rolling ball of fire deals 3d6 fire damage.

\textbf{Gust of Wind}: Blows away or knocks down smaller creatures.

\textbf{Scorching Ray}: Ranged touch attack deals 4d6 fire damage, + 1 ray/four levels (max 3).

\textbf{Shatter}: Sonic energy damages objects or crystalline creatures.

\subsection{Illusion}


\textbf{Blur}: Attacks miss subject 20\% of the time.

\textbf{Hypnotic Pattern}: Fascinates 2d4 + level HD of creatures.

\textbf{Invisibility}: Subject is invisible for 1 min./level or until it attacks.

\textbf{Magic Mouth}\(^{M}\): Object speaks once when triggered.

\textbf{Minor Image}: As \textit{silent image}, plus some sound.

\textbf{Mirror Image}: Creates decoy duplicates of you.

\textbf{Misdirection}: Misleads divinations for 1 creature or object.

\textbf{Phantom Trap}\(^{M}\): Makes item seem trapped.

\subsection{Necromancy}


\textbf{Blindness/Deafness}: Makes subject blinded or deafened.

\textbf{Command Undead}: Undead creature obeys your commands.

\textbf{False Life}: Gain 1d10 temporary hp + 1/level (max +10).

\textbf{Ghoul Touch}: Paralyzes one subject, which exudes stench that makes those nearby sickened.

\textbf{Scare}: Frightens creatures of less than 6 HD.

\textbf{Spectral Hand}: Creates disembodied glowing hand to deliver touch attacks.

\subsection{Transmutation}


\textbf{Alter Self}: Assume form of a Small or Medium humanoid.

\textbf{Bear's Endurance}: Subject gains +4 to Con for 1 min./level.

\textbf{Bull's Strength}: Subject gains +4 to Str for 1 min./level.

\textbf{Cat's Grace}: Subject gains +4 to Dex for 1 min./level.

\textbf{Darkvision}: See 60 ft. in total darkness.

\textbf{Eagle's Splendor}: Subject gains +4 to Cha for 1 min./level.

\textbf{Fox's Cunning}: Subject gains +4 to Int for 1 min./level.

\textbf{Knock}: Opens locked or magically sealed door.

\textbf{Levitate}: Subject moves up and down at your direction.

\textbf{Make Whole}: Repairs an object.

\textbf{Owl's Wisdom}: Subject gains +4 to Wis for 1 min./level.

\textbf{Pyrotechnics}: Turns fire into blinding light or thick smoke.

\textbf{Rope Trick}: As many as eight creatures hide in extradimensional space.

\textbf{Spider Climb}: Grants ability to walk on walls and ceilings.

\textbf{Whispering Wind}: Sends a short message 1 mile/level.

\subsection{3rd-Level Sorcerer/Wizard Spells}


\subsection{Abjuration}


\textbf{Dispel Magic}: Cancels one magical spell or effect.

\textbf{Explosive Runes}: Deals 6d6 damage when read.

\textbf{Magic Circle against Chaos/Evil/Good/Law}: As \textit{protection }spells, but 10-ft. radius and 10 min./level.

\textbf{Nondetection}\(^{M}\): Hides subject from divination, scrying.

\textbf{Protection from Energy}: Absorbs 12 points/level of damage from one kind of energy.

\subsection{Conjuration}


\textbf{Phantom Steed}: Magic horse appears for 1 hour/level.

\textbf{Sepia Snake Sigil}\(^{M}\): Creates text symbol that immobilizes reader.

\textbf{Sleet Storm}: Hampers vision and movement.

\textbf{Stinking Cloud}: Nauseating vapors, 1 round/level.

\textbf{Summon Monster III}: Summons extraplanar creature to fight for you.

\subsection{Divination}


\textbf{Arcane Sight}: Magical auras become visible to you.

\textbf{Clairaudience/Clairvoyance}: Hear or see at a distance for 1 min./level.

\textbf{Tongues}: Speak and understand any language.

\subsection{Enchantment}


\textbf{Deep Slumber}: Puts 10 HD of creatures to sleep.

\textbf{Heroism}: Gives +2 bonus on attack rolls, saves, skill checks.

\textbf{Hold Person}: Paralyzes one humanoid for 1 round/level.

\textbf{Rage}: Gives +2 to Str and Con, +1 on Will saves, --2 to AC.

\textbf{Suggestion}: Compels a subject to follow stated course of action.

\subsection{Evocation}


\textbf{Daylight}: 60-ft. radius of bright light.

\textbf{Fireball}: 1d6 damage per level, 20-ft. radius.

\textbf{Lightning Bolt}: Electricity deals 1d6/level damage.

\textbf{Tiny Hut}: Creates shelter for 10 creatures.

\textbf{Wind Wall}: Deflects arrows, smaller creatures, and gases.

\subsection{Illusion}


\textbf{Displacement}: Attacks miss subject 50\% of the time.

\textbf{Illusory Script}\(^{M}\): Only select creatures can read text.

\textbf{Invisibility Sphere}: Makes everyone within 10 ft. invisible.

\textbf{Major Image}: As \textit{silent image}, plus sound, smell and thermal effects.

\subsection{Necromancy}


\textbf{Gentle Repose}: Preserves one corpse.

\textbf{Halt Undead}: Immobilizes undead for 1 round/level.

\textbf{Ray of Exhaustion}: Ray makes subject exhausted.

\textbf{Vampiric Touch}: Touch deals 1d6 damage per two levels; caster gains damage as temporary hp.

\subsection{Transmutation}


\textbf{Beast Shape I}: You take the form and some of the powers of a Small or Medium animal.

\textbf{Blink}: You randomly vanish and reappear for 1 round per level.

\textbf{Flame Arrow}: Arrows deal +1d6 fire damage.

\textbf{Fly}: Subject flies at speed of 60 ft.

\textbf{Gaseous Form}: Subject becomes insubstantial and can fly slowly.

\textbf{Haste}: One creature/level moves faster, +1 on attack rolls, AC, and Reflex saves.

\textbf{Keen Edge}: Doubles normal weapon's threat range.

\textbf{Magic Weapon, Greater}: Weapon gains +1 bonus/four levels (max +5).

\textbf{Secret Page}: Changes one page to hide its real content.

\textbf{Shrink Item}: Object shrinks to one-sixteenth size.

\textbf{Slow}: One subject/level takes only one action/round, --1 to AC, Reflex saves, and attack rolls.

\textbf{Water Breathing}: Subjects can breathe underwater.

\subsection{4th-Level Sorcerer/Wizard Spells}


\subsection{Abjuration}


\textbf{Dimensional Anchor}: Bars extradimensional movement.

\textbf{Fire Trap}\(^{M}\): Opened object deals 1d4 damage + 1/level.

\textbf{Globe of Invulnerability, Lesser}: Stops 1st- through 3rd-level spell effects.

\textbf{Remove Curse}: Frees object or person from curse.

\textbf{Stoneskin}\(^{M}\): Grants DR 10/adamantine.

\subsection{Conjuration}


\textbf{Black Tentacles}: Tentacles grapple all creatures within a 20-ft. spread.

\textbf{Dimension Door}: Teleports you a short distance.

\textbf{Minor Creation}: Creates one cloth or wood object.

\textbf{Secure Shelter}: Creates sturdy cottage.

\textbf{Solid Fog}: Blocks vision and slows movement.

\textbf{Summon Monster IV}: Summons extraplanar creature to fight for you.

\subsection{Divination}


\textbf{Arcane Eye}: Invisible floating eye moves 30 ft./round.

\textbf{Detect Scrying}: Alerts you to magical eavesdropping

\textbf{Locate Creature}: Indicates direction to familiar creature.

\textbf{Scrying}\(^{F}\): Spies on subject from a distance.

\subsection{Enchantment}


\textbf{Charm Monster}: Makes monster believe it is your ally.

\textbf{Confusion}: Subjects behave oddly for 1 round/level.

\textbf{Crushing Despair}: Subjects take --2 on attack rolls, damage rolls, saves, and checks.

\textbf{Geas, Lesser}: Commands subject of 7 HD or less.

\subsection{Evocation}


\textbf{Fire Shield}: Creatures attacking you take fire damage; you're protected from heat or cold.

\textbf{Ice Storm}: Hail deals 5d6 damage in cylinder 40 ft. across.

\textbf{Resilient Sphere}: Force globe protects but traps one subject.

\textbf{Shout}: Deafens all within cone and deals 5d6 sonic damage.

\textbf{Wall of Fire}: Deals 2d4 fire damage out to 10 ft. and 1d4 out to 20 ft. Passing through wall deals 2d6 damage + 1/level.

\textbf{Wall of Ice}: \textit{Ice plane }creates wall or \textit{hemisphere }creates dome.

\subsection{Illusion}


\textbf{Hallucinatory Terrain}: Makes one type of terrain appear like another (field as forest, or the like).

\textbf{Illusory Wall}: Wall, floor, or ceiling looks real, but anything can pass through.

\textbf{Invisibility, Greater}: As \textit{invisibility, }but subject can attack and stay invisible.

\textbf{Phantasmal Killer}: Fearsome illusion kills subject or deals 3d6 damage.

\textbf{Rainbow Pattern}: Lights fascinate 24 HD of creatures.

\textbf{Shadow Conjuration}: Mimics conjuration below 4th level, but only 20\% real.

\subsection{Necromancy}


\textbf{Animate Dead}\(^{M}\): Creates undead skeletons and zombies out of corpses.

\textbf{Bestow Curse}: --6 to an ability score; --4 on attack rolls, saves, and checks; or 50\% chance of losing each action.

\textbf{Contagion}: Infects subject with chosen disease.

\textbf{Enervation}: Subject gains 1d4 negative levels.

\textbf{Fear}: Subjects within cone flee for 1 round/level.

\subsection{Transmutation}


\textbf{Beast Shape II}: You take the form and some of the powers of a Tiny or Large animal.

\textbf{Elemental Body I}: Turns you into a Small elemental.

\textbf{Enlarge Person, Mass}: 1 humanoid creature/level doubles in size.

\textbf{Mnemonic Enhancer}\(^{F}\): \textit{Wizard only. }Prepare extra spells or retain one just cast.

\textbf{Reduce Person, Mass}: As \textit{reduce person}, but affects 1 humanoid creature/level.

\textbf{Stone Shape}: Sculpts stone into any shape.

\subsection{5th-Level Sorcerer/Wizard Spells}


\subsection{Abjuration}


\textbf{Break Enchantment}: Frees subjects from enchantments, transmutations, and curses.

\textbf{Dismissal}: Forces a creature to return to its native plane.

\textbf{Mage's Private Sanctum}: Prevents anyone from viewing or scrying an area for 24 hours.

\subsection{Conjuration}


\textbf{Cloudkill}: Kills 3 HD or less; 4--6 HD save or die, 6+ HD take Con damage.

\textbf{Mage's Faithful Hound}: Phantom dog can guard a location and attack intruders.

\textbf{Major Creation}: As \textit{minor creation, }plus stone and metal.

\textbf{Planar Binding, Lesser}: Traps extraplanar creature of 6 HD or less until it performs a task.

\textbf{Secret Chest}\(^{F}\): Hides expensive chest on Ethereal Plane; you retrieve it at will.

\textbf{Summon Monster V}: Summons extraplanar creature to fight for you.

\textbf{Teleport}: Instantly transports you as far as 100 miles per level.

\textbf{Wall of Stone}: Creates a stone wall that can be shaped.

\subsection{Divination}


\textbf{Contact Other Plane}: Lets you ask question of extraplanar entity.

\textbf{Prying Eyes}: 1d4 + 1/level floating eyes scout for you.

\textbf{Telepathic Bond}: Link lets allies communicate.

\subsection{Enchantment}


\textbf{Dominate Person}: Controls humanoid telepathically.

\textbf{Feeblemind}: Subject's Int and Cha drop to 1.

\textbf{Hold Monster}: As \textit{hold person}, but any creature.

\textbf{Mind Fog}: Subjects in fog get --10 to Wis and Will checks.

\textbf{Symbol of Sleep}\(^{M}\): Triggered rune puts nearby creatures into catatonic slumber.

\subsection{Evocation}


\textbf{Cone of Cold}: 1d6/level cold damage.

\textbf{Interposing Hand}: Hand provides cover against 1 opponent.

\textbf{Sending}: Delivers short message anywhere, instantly.

\textbf{Wall of Force}: Wall is immune to damage.

\subsection{Illusion}


\textbf{Dream}: Sends message to anyone sleeping.

\textbf{False Vision}\(^{M}\): Fools scrying with an illusion.

\textbf{Mirage Arcana}: As \textit{hallucinatory terrain, }plus structures.

\textbf{Nightmare}: Sends vision dealing 1d10 damage, fatigue.

\textbf{Persistent Image}: As \textit{major image}, but with no concentration required.

\textbf{Seeming}: Changes appearance of 1 person per 2 levels.

\textbf{Shadow Evocation}: Mimics evocation below 5th level, but only 20\% real.

\subsection{Necromancy}


\textbf{Blight}: Withers one plant or deals 1d6/level damage to plant creature.

\textbf{Magic Jar}\(^{F}\): Enables possession of another creature.

\textbf{Symbol of Pain}\(^{M}\): Triggered rune wracks creatures with pain.

\textbf{Waves of Fatigue}: Several targets become fatigued.

\subsection{Transmutation}


\textbf{Animal Growth}: One animal doubles in size.

\textbf{Baleful Polymorph}: Turns subject into harmless animal.

\textbf{Beast Shape III}: You take the form of a Diminutive or Huge animal, or Small or Medium magical beast.

\textbf{Elemental Body II}: Turns you into a Medium elemental.

\textbf{Fabricate}\(^{M}\): Transforms raw materials into finished items.

\textbf{Overland Flight}: You fly at a speed of 40 ft. and can hustle over long distances.

\textbf{Passwall}: Creates passage through wood or stone wall.

\textbf{Plant Shape I}: Turns you into a Small or Medium plant.

\textbf{Polymorph}: Gives one willing subject a new form.

\textbf{Telekinesis}: Moves object, attacks creature, or hurls object or creature.

\textbf{Transmute Mud to Rock}: Transforms two 10-ft. cubes per level.

\textbf{Transmute Rock to Mud}: Transforms two 10-ft. cubes per level.

\subsection{Universal}


\textbf{Permanency}\(^{M}\): Makes certain spells permanent.

\subsection{6th-Level Sorcerer/Wizard Spells}


\subsection{Abjuration}


\textbf{Antimagic Field}: Negates magic within 10 ft.

\textbf{Dispel Magic, Greater}: As \textit{dispel magic}, but with multiple targets.

\textbf{Globe of Invulnerability}: As \textit{lesser globe of invulnerability}, plus 4th-level spell effects.

\textbf{Guards and Wards}: Array of magic effects protect area.

\textbf{Repulsion}\(^{F}\): Creatures can't approach you.

\subsection{Conjuration}


\textbf{Acid Fog}: Fog deals acid damage.

\textbf{Planar Binding}: As \textit{lesser planar binding, }but up to 12 HD.

\textbf{Summon Monster VI}: Summons extraplanar creature to fight for you.

\textbf{Wall of Iron}\(^{M}\): 30 hp/four levels; can topple onto foes.

\subsection{Divination}


\textbf{Analyze Dweomer}\(^{F}\): Reveals magical aspects of subject.

\textbf{Legend Lore}\(^{MF}\): Lets you learn tales about a person, place, or thing.

\textbf{True Seeing}\(^{M}\): Lets you see all things as they really are.

\subsection{Enchantment}


\textbf{Geas/Quest}: As \textit{lesser geas, }but affects any creature.

\textbf{Heroism, Greater}: Gives +4 bonus on attack rolls, saves, skill checks; immunity to fear; temporary hp.

\textbf{Suggestion, Mass}: As \textit{suggestion, }affects 1 subject/level.

\textbf{Symbol of Persuasion}\(^{M}\): Triggered rune charms creatures.

\subsection{Evocation}


\textbf{Chain Lightning}: 1d6/level damage and 1 secondary bolt/level.

\textbf{Contingency}\(^{F}\): Sets trigger condition for another spell.

\textbf{Forceful Hand}: Hand pushes creatures away.

\textbf{Freezing Sphere}: Freezes water or deals cold damage.

\subsection{Illusion}


\textbf{Mislead}: Turns you invisible and creates illusory double.

\textbf{Permanent Image}: Permanent illusion, includes sight, sound, smell, and thermal effects.

\textbf{Programmed Image}\(^{M}\): As \textit{major image}, but triggered by event.

\textbf{Shadow Walk}: Step into shadow to travel rapidly.

\textbf{Veil}: Changes appearance of a group of creatures.

\subsection{Necromancy}


\textbf{Circle of Death}\(^{M}\): Kills 1d4/level HD of creatures.

\textbf{Create Undead}\(^{M}\): Raises ghouls, ghasts, mummies, or mohrgs from physical remains.

\textbf{Eyebite}: Target becomes panicked, sickened, and comatose.

\textbf{Symbol of Fear}\(^{M}\): Triggered rune panics nearby creatures.

\textbf{Undeath to Death}\(^{M}\): Destroys 1d4/level HD of undead (max. 20d4).

\subsection{Transmutation}


\textbf{Bear's Endurance, Mass}: As \textit{bear's endurance}, affects one subject/level.

\textbf{Beast Shape IV}: You take the form of a Diminutive to Huge animal or a Tiny to Large magical beast.

\textbf{Bull's Strength, Mass}: As \textit{bull's strength}, affects 1 subject per level.

\textbf{Cat's Grace, Mass}: As \textit{cat's grace}, affects 1 subject/level.

\textbf{Control Water}: Raises or lowers bodies of water.

\textbf{Disintegrate}: Reduces one creature or object to dust.

\textbf{Eagle's Splendor, Mass}: As \textit{eagle's splendor}, 1 subject/level.

\textbf{Elemental Body III}: Turns you into a Large elemental.

\textbf{Flesh to Stone}: Turns subject creature into statue.

\textbf{Form of the Dragon I}: Turns you into a Medium dragon.

\textbf{Fox's Cunning, Mass}: As \textit{fox's cunning}, affects 1 subject/ level.

\textbf{Mage's Lucubration}: \textit{Wizard only. }Recalls spell of 5th level or lower.

\textbf{Move Earth}: Digs trenches and builds hills.

\textbf{Owl's Wisdom, Mass}: As \textit{owl's wisdom}, affects 1 subject/level.

\textbf{Plant Shape II}: Turns you into a Large plant creature.

\textbf{Stone to Flesh}: Restores petrified creature.

\textbf{Transformation}\(^{M}\): You gain combat bonuses.

\subsection{7th-Level Sorcerer/Wizard Spells}


\subsection{Abjuration}


\textbf{Banishment}: Banishes 2 HD/level of extraplanar creatures.

\textbf{Sequester}: Subject is invisible to sight and scrying; renders creature comatose.

\textbf{Spell Turning}\(^{M}\): Reflect 1d4+6 spell levels back at caster.

\subsection{Conjuration}


\textbf{Instant Summons}\(^{M}\): Prepared object appears in your hand.

\textbf{Mage's Magnificent Mansion}\(^{F}\): Door leads to extradimensional mansion.

\textbf{Phase Door}: Creates an invisible passage through a barrier.

\textbf{Plane Shift}\(^{F}\): As many as eight subjects travel to another plane.

\textbf{Summon Monster VII}: Summons extraplanar creature to fight for you.

\textbf{Teleport, Greater}: As \textit{teleport, }but no range limit and no off-target arrival.

\textbf{Teleport Object}: As \textit{teleport, }but affects a touched object.

\subsection{Divination}


\textbf{Arcane Sight, Greater}: As \textit{arcane sight}, but also reveals magic effects on creatures and objects.

\textbf{Scrying, Greater}: As \textit{scrying}, but faster and longer.

\textbf{Vision}\(^{M}\): As \textit{legend lore}, but quicker.

\subsection{Enchantment}


\textbf{Hold Person, Mass}: As \textit{hold person}, but all within 30 ft.

\textbf{Insanity}: Subject suffers continuous \textit{confusion}.

\textbf{Power Word Blind}: Blinds creature with 200 hp or less.

\textbf{Symbol of Stunning}\(^{M}\): Triggered rune stuns creatures.

\subsection{Evocation}


\textbf{Delayed Blast Fireball}: 1d6/level fire damage; you can postpone blast for up to 5 rounds.

\textbf{Forcecage}\(^{M}\): Cube or cage of force imprisons all inside.

\textbf{Grasping Hand}: Hand provides cover, pushes, or grapples.

\textbf{Mage's Sword}\(^{F}\): Floating magic blade strikes opponents.

\textbf{Prismatic Spray}: Rays hit subjects with variety of effects.

\subsection{Illusion}


\textbf{Invisibility, Mass}: As \textit{invisibility}, but affects all in range.

\textbf{Project Image}: Illusory double can talk and cast spells.

\textbf{Shadow Conjuration, Greater}: As \textit{shadow conjuration, }but up to 6th level and 60\% real.

\textbf{Simulacrum}\(^{M}\): Creates partially real double of a creature.

\subsection{Necromancy}


\textbf{Control Undead}: Undead don't attack you while under your command.

\textbf{Finger of Death}: Deals 10 damage/level to one subject.

\textbf{Symbol of Weakness}\(^{M}\): Triggered rune weakens creatures.

\textbf{Waves of Exhaustion}: Several targets become exhausted.

\subsection{Transmutation}


\textbf{Control Weather}: Changes weather in local area.

\textbf{Elemental Body IV}: Turns you into a Huge elemental.

\textbf{Ethereal Jaunt}: You become ethereal for 1 round/level.

\textbf{Form of the Dragon II}: Turns you into a Large dragon.

\textbf{Giant Form I}: Turns you into a Large giant.

\textbf{Plant Shape III}: Turns you into a Huge plant.

\textbf{Polymorph, Greater}: Gives one willing subject a new, more powerful form.

\textbf{Reverse Gravity}: Objects and creatures fall upward.

\textbf{Statue}: Subject can become a statue at will.

\subsection{Universal}


\textbf{Limited Wish}\(^{M}\): Alters reality (within limits).

\subsection{8th-Level Sorcerer/Wizard Spells}


\subsection{Abjuration}


\textbf{Dimensional Lock}: Teleportation and interplanar travel blocked for 1 day/level.

\textbf{Mind Blank}: Subject is protected from mental/emotional magic and scrying.

\textbf{Prismatic Wall}: Wall's colors have array of effects.

\textbf{Protection from Spells}\(^{MF}\): Confers +8 resistance bonus.

\subsection{Conjuration}


\textbf{Incendiary Cloud}: Cloud deals 6d6 fire damage/round.

\textbf{Maze}: Traps subject in extradimensional maze.

\textbf{Planar Binding, Greater}: As \textit{lesser planar binding, }but up to 18 HD.

\textbf{Summon Monster VIII}: Summons extraplanar creature to fight for you.

\textbf{Trap the Soul}\(^{M}\): Imprisons subject within gem.

\subsection{Divination}


\textbf{Discern Location}: Reveals exact location of creature or object.

\textbf{Moment of Prescience}: You gain +1/level insight bonus on single attack roll, check, or save.

\textbf{Prying Eyes, Greater}: As \textit{prying eyes}, but eyes have \textit{true seeing}.

\subsection{Enchantment}


\textbf{Antipathy}: Object or location affected by spell repels certain creatures.

\textbf{Binding}\(^{M}\): Utilizes an array of techniques to imprison a creature.

\textbf{Charm Monster, Mass}: As \textit{charm monster, }but all within 30 ft.

\textbf{Demand}: As \textit{sending, }plus you can send \textit{suggestion.}

\textbf{Irresistible Dance}: Forces subject to dance.

\textbf{Power Word Stun}: Stuns creature with 150 hp or less.

\textbf{Symbol of Insanity}\(^{M}\): Triggered rune renders nearby creatures insane.

\textbf{Sympathy}\(^{M}\): Object or location attracts certain creatures.

\subsection{Evocation}


\textbf{Clenched Fist}: Large hand provides cover, pushes, or attacks your foes.

\textbf{Polar Ray}: Ranged touch attack deals 1d6/level cold damage and 1d4 points of Dexterity drain.

\textbf{Shout, Greater}: Devastating yell deals 10d6 sonic damage; stuns creatures.

\textbf{Sunburst}: Blinds all within 10 ft., deals 6d6 damage.

\textbf{Telekinetic Sphere}: As\textit{ resilient sphere, }but you move the sphere telekinetically.

\subsection{Illusion}


\textbf{Scintillating Pattern}: Twisting colors \textit{confuse, }stun, or render unconscious.

\textbf{Screen}: Illusion hides area from vision and scrying.

\textbf{Shadow Evocation, Greater}: As \textit{shadow evocation, }but up to 7th level and 60\% real.

\subsection{Necromancy}


\textbf{Clone}\(^{MF}\): Duplicate awakens when original dies.

\textbf{Create Greater Undead}\(^{M}\): Creates shadows, wraiths, spectres, or devourers.

\textbf{Horrid Wilting}: Deals 1d6/level damage within 30 ft.

\textbf{Symbol of Death}\(^{M}\): Triggered rune kills nearby creatures.

\subsection{Transmutation}


\textbf{Form of the Dragon III}: Turns you into a Huge dragon.

\textbf{Giant Form II}: Turns you into a Huge giant.

\textbf{Iron Body}: Your body becomes living iron.

\textbf{Polymorph Any Object}: Changes a subject into anything else.

\textbf{Temporal Stasis}\(^{M}\): Puts subject into suspended animation.

\subsection{9th-Level Sorcerer/Wizard Spells}


\subsection{Abjuration}


\textbf{Freedom}: Releases creature from \textit{imprisonment.}

\textbf{Imprisonment}: Entombs subject beneath the earth.

\textbf{Mage's Disjunction}: Dispels magic, disenchants magic items.

\textbf{Prismatic Sphere}: As \textit{prismatic wall}, but surrounds on all sides.

\subsection{Conjuration}


\textbf{Gate}\(^{M}\): Connects two planes for travel or summoning.

\textbf{Refuge}\(^{M}\): Alters item to transport its possessor to your abode.

\textbf{Summon Monster IX}: Summons extraplanar creature to fight for you.

\textbf{Teleportation Circle}\(^{M}\): Teleports creatures inside circle.

\subsection{Divination}


\textbf{Foresight}: \texttt{{}"{}}Sixth sense\texttt{{}"{}} warns of impending danger.

\subsection{Enchantment}


\textbf{Dominate Monster}: As \textit{dominate person, }but any creature.

\textbf{Hold Monster, Mass}: As \textit{hold monster, }but all within 30 ft.

\textbf{Power Word Kill}: Kills one creature with 100 hp or less.

\subsection{Evocation}


\textbf{Crushing Hand}: Large hand provides cover, pushes, or crushes your foes.

\textbf{Meteor Swarm}: Four exploding spheres each deal 6d6 fire damage.

\subsection{Illusion}


\textbf{Shades}: As \textit{shadow conjuration}, but up to 8th level and 80\% real.

\textbf{Weird}: As \textit{phantasmal killer, }but affects all within 30 ft.

\subsection{Necromancy}


\textbf{Astral Projection}\(^{M}\): Projects you and companions onto Astral Plane.

\textbf{Energy Drain}: Subject gains 2d4 negative levels.

\textbf{Soul Bind}\(^{F}\): Traps newly dead soul to prevent \textit{resurrection}.

\textbf{Wail of the Banshee}: Deals 10 damage/level to 1 creature/level.

\subsection{Transmutation}


\textbf{Etherealness}: Travel to Ethereal Plane with companions.

\textbf{Shapechange}\(^{F}\): Transforms you into certain creatures, and lets you change forms once per round.

\textbf{Time Stop}: You act freely for 1d4+1 rounds.

\subsection{Universal}


\textbf{Wish}\(^{M}\): As \textit{limited wish, }but with fewer limits.
        
