\chapter{Prestige Classes}

\label{f0}				
Prestige classes allow characters to become truly exceptional, gaining powers beyond the ken of their peers. Unlike the core classes\textit{, }characters must meet specific requirements before they can take their first level of a prestige class. If a character does not meet the requirements for a prestige class before gaining any benefits of that level, that character cannot take that prestige class. Characters that take levels in prestige classes do not gain any favored class bonuses for those levels.
				
\textbf{Arcane Archer}: An arcane spellcaster who draws upon ancient elven traditions to infuse his arrows with potent magical power.
				
\textbf{Arcane Trickster}: A troublemaker and a scoundrel who uses arcane magic to enhance her thievery and trickery.
				
\textbf{Assassin}: A remorseless murderer who kills for money and the sheer thrill of death-dealing.
				
\textbf{Dragon Disciple}: An arcane spellcaster who has embraced his latent draconic heritage and, over the course of training and devotion, undergoes a partial transformation into a dragon.
				
\textbf{Duelist}: A swashbuckling swordfighter who relies upon grace, poise, and acrobatics to win the day.
				
\textbf{Eldritch Knight}: An arcane spellcaster who augments his magical skills with combat to create a deadly combination of weapons and magic.
				
\textbf{Loremaster}: A spellcaster who devotes his life to research and rumination upon the mysteries of the world.
				
\textbf{Mystic Theurge}: Equally devoted to divine and arcane magic, the mystic theurge combines both magical traditions into one incredibly diverse class.
				
\textbf{Pathfinder Chronicler}: An explorer at heart, the Pathfinder chronicler travels to distant, exotic lands to expand her knowledge of the world. 
				
\textbf{Shadowdancer}: A mysterious adventurer who walks the boundaries between the real world and the realm of shadows, and who can command shadows to do her bidding.
				
\subsection{Definitions of Terms}

				
Here are definitions of some terms used in this section.
				
\textbf{Core Class}: One of the standard eleven classes found in Classes.
				
\textbf{Caster Level}: Generally equal to the number of class levels (see below) in a spellcasting class. Some prestige classes add caster levels to an existing class.
				
\textbf{Character Level}: The total level of the character, which is the sum of all class levels held by that character.
				
\textbf{Class Level}: The level of a character in a particular class. For a character with levels in only one class, class level and character level are the same.
	        	
\input{classes-prestige/arcane-archer}
\input{classes-prestige/arcane-trickster}
\input{classes-prestige/assassin}
\input{classes-prestige/dragon-disciple}
\input{classes-prestige/duelist}
\input{classes-prestige/eldritch-knight}
\input{classes-prestige/loremaster}
\input{classes-prestige/mystic-theurge}
\input{classes-prestige/pathfinder-chronicler}
\input{classes-prestige/shadowdancer}
