\spellentry{Command}
\textbf{School} enchantment (compulsion) [language-dependent, mind-affecting]; \textbf{Level} cleric 1\\
\textbf{Casting Time} 1 standard action\\
\textbf{Components} V\\
\textbf{Range} close (25 ft. + 5 ft./2 levels)\\
\textbf{Target} one living creature\\
\textbf{Duration} 1 round\\
\textbf{Saving Throw }Will negates; \textbf{Spell Resistance} yes\\
You give the subject a single command, which it obeys to the best of its ability at its earliest opportunity. You may select from the following options.\\
\textit{Approach}: On its turn, the subject moves toward you as quickly and directly as possible for 1 round. The creature may do nothing but move during its turn, and it provokes attacks of opportunity for this movement as normal.\\
\textit{Drop}: On its turn, the subject drops whatever it is holding. It can't pick up any dropped item until its next turn.\\
\textit{Fall}: On its turn, the subject falls to the ground and remains prone for 1 round. It may act normally while prone but takes any appropriate penalties.\\
\textit{Flee}: On its turn, the subject moves away from you as quickly as possible for 1 round. It may do nothing but move during its turn, and it provokes attacks of opportunity for this movement as normal.\\
\textit{Halt}: The subject stands in place for 1 round. It may not take any actions but is not considered helpless.\\
If the subject can't carry out your command on its next turn, the spell automatically fails.\\
\textbf{Command, Greater}\\
\textbf{School} enchantment (compulsion) [language-dependent, mind-affecting]; \textbf{Level} cleric 5 \\
\textbf{Targets} one creature/level, no two of which can be more than 30 ft. apart\\
\textbf{Duration} 1 round/level\\
This spell functions like \textit{command, }except that up to one creature per level may be affected, and the activities continue beyond 1 round. At the start of each commanded creature's action after the first, it gets another Will save to attempt to break free from the spell. Each creature must receive the same command.\\
