\spellentry{Dictum}
\textbf{School} evocation [lawful, sonic]; \textbf{Level} cleric 7\\
\textbf{Casting Time} 1 standard action\\
\textbf{Components} V\\
\textbf{Range} 40 ft.\\
\textbf{Area} nonlawful creatures in a 40-ft.-radius spread centered on you\\
\textbf{Duration} instantaneous\\
\textbf{Saving Throw} none or Will negates; see text; \textbf{Spell Resistance} yes\\
Any nonlawful creature within the area of a \textit{dictum }spell suffers the following ill effects.\\
The effects are cumulative and concurrent. A successful Will save reduces or eliminates these effects. Creatures affected by multiple effects make only one save and apply the result to all the effects.\\
\textit{Deafened}: The creature is deafened for 1d4 rounds. Save negates.\\
\textit{Staggered}: The creature is staggered for 2d4 rounds. Save reduces the staggered effect to 1d4 rounds.\\
\textit{Paralyzed}: The creature is paralyzed and helpless for 1d10 minutes. Save reduces the paralyzed effect to 1 round.\\
\textit{Killed}: Living creatures die. Undead creatures are destroyed. Save negates. If the save is successful, the creature instead takes 3d6 points of damage + 1 point per caster level (maximum +25).\\
Furthermore, if you are on your home plane when you cast this spell, nonlawful extraplanar creatures within the area are instantly banished back to their home planes. Creatures so banished cannot return for at least 24 hours. This effect takes place regardless of whether the creatures hear the \textit{dictum }or not\textit{. }The banishment effect allows a Will save (at a --4 penalty) to negate.\\
Creatures whose Hit Dice exceed your caster level are unaffected by \textit{dictum.}\\
