\spellentry{Elemental Body I}
\textbf{School }transmutation (polymorph); \textbf{Level }sorcerer/wizard 4\\
\textbf{Casting Time }1 standard action\\
\textbf{Components }V, S, M (the element you plan to assume)\\
\textbf{Range }personal\\
\textbf{Target }you\\
\textbf{Duration }1 min/level (D)\\
When you cast this spell, you can assume the form of a Small air, earth, fire, or water elemental. The abilities you gain depend upon the type of elemental into which you change. Elemental abilities based on size, such as burn, vortex, and whirlwind, use the size of the elemental you transform into to determine their effect.\\
\textit{Air elemental}: If the form you take is that of a Small air elemental, you gain a +2 size bonus to your Dexterity and a +2 natural armor bonus. You also gain fly 60 feet (perfect), darkvision 60 feet, and the ability to create a whirlwind.\\
\textit{Earth elemental}: If the form you take is that of a Small earth elemental, you gain a +2 size bonus to your Strength and a +4 natural armor bonus. You also gain darkvision 60 feet, and the ability to earth glide.\\
\textit{Fire elemental}: If the form you take is that of a Small fire elemental, you gain a +2 size bonus to your Dexterity and a +2 natural armor bonus. You gain darkvision 60 feet, resist fire 20, vulnerability to cold, and the burn ability.\\
\textit{Water elemental}: If the form you take is that of a Small water elemental, you gain a +2 size bonus to your Constitution and a +4 natural armor bonus. You also gain swim 60 feet, darkvision 60 feet, the ability to create a vortex, and the ability to breathe water.\\
\textbf{Elemental Body II}\\
\textbf{School }transmutation (polymorph); \textbf{Level }sorcerer/wizard 5\\
This spell functions as \textit{elemental body I, }except that it also allows you to assume the form of a Medium air, earth, fire, or water elemental. The abilities you gain depend upon the elemental.\\
\textit{Air elemental}: As \textit{elemental body I} except that you gain a +4 size bonus to your Dexterity and a +3 natural armor bonus. \\
\textit{Earth elemental}: As \textit{elemental body I} except that you gain a +4 size bonus to your Strength and a +5 natural armor bonus. \\
\textit{Fire elemental}: As \textit{elemental body I} except that you gain a +4 size bonus to your Dexterity and a +3 natural armor bonus.\\
\textit{Water elemental}: As \textit{elemental body I} except that you gain a +4 size bonus to your Constitution and a +5 natural armor bonus. \\
\textbf{Elemental Body III}\\
\textbf{School }transmutation (polymorph); \textbf{Level }sorcerer/wizard 6\\
This spell functions as \textit{elemental body II, }except that it also allows you to assume the form of a Large air, earth, fire, or water elemental. The abilities you gain depend upon the type of elemental into which you change. You are also immune to bleed damage, critical hits, and sneak attacks while in elemental form.\\
\textit{Air elemental}: As \textit{elemental body I} except that you gain a +2 size bonus to your Strength, +4 size bonus to your Dexterity, and a +4 natural armor bonus. \\
\textit{Earth elemental}: As \textit{elemental body I} except that you gain a +6 size bonus to your Strength, a --2 penalty on your Dexterity, a +2 size bonus to your Constitution, and a +6 natural armor bonus. \\
\textit{Fire elemental}: As \textit{elemental body I} except that you gain a +4 size bonus to your Dexterity, a +2 size bonus to your Constitution, and a +4 natural armor bonus.\\
\textit{Water elemental}: As \textit{elemental body I} except that you gain a +2 size bonus to your Strength, a --2 penalty on your Dexterity, a +6 size bonus to your Constitution, and a +6 natural armor bonus. \\
\textbf{Elemental Body IV}\\
\textbf{School }transmutation (polymorph); \textbf{Level }sorcerer/wizard 7\\
This spell functions as \textit{elemental body III, }except that it also allows you to assume the form of a Huge air, earth, fire, or water elemental. The abilities you gain depend upon the type of elemental into which you change. You are also immune to bleed damage, critical hits, and sneak attacks while in elemental form and gain DR 5/—.\\
\textit{Air elemental}: As \textit{elemental body I} except that you gain a +4 size bonus to your Strength, +6 size bonus to your Dexterity, and a +4 natural armor bonus. You also gain fly 120 feet (perfect).\\
\textit{Earth elemental}: As \textit{elemental body I} except that you gain a +8 size bonus to your Strength, a --2 penalty on your Dexterity, a +4 size bonus to your Constitution, and a +6 natural armor bonus. \\
\textit{Fire elemental}: As \textit{elemental body I} except that you gain a +6 size bonus to your Dexterity, a +4 size bonus to your Constitution, and a +4 natural armor bonus.\\
\textit{Water elemental}: As \textit{elemental body I} except that you gain a +4 size bonus to your Strength, a --2 penalty on your Dexterity, a +8 size bonus to your Constitution, and a +6 natural armor bonus. You also gain swim 120 feet.\\
