\spellentry{Forcecage}
\textbf{School }evocation [force]; \textbf{Level }sorcerer/wizard 7\\
\textbf{Casting Time }1 standard action\\
\textbf{Components }V, S, M (ruby dust worth 500 gp)\\
\textbf{Range }close (25 ft. + 5 ft./2 levels)\\
\textbf{Area }barred cage (20-ft. cube) or windowless cell (10-ft. cube)\\
\textbf{Duration }1 round/level (D)\\
\textbf{Saving Throw }Reflex negates; \textbf{Spell Resistance }no\\
This spell creates an immobile, invisible cubical prison composed of either bars of force or solid walls of force (your choice).\\
Creatures within the area are caught and contained unless they are too big to fit inside, in which case the spell automatically fails. Teleportation and other forms of astral travel provide a means of escape, but the force walls or bars extend into the Ethereal Plane, blocking ethereal travel.\\
Like a \textit{wall of force}, a \textit{forcecage} resists \textit{dispel magic, }although a \textit{mage's disjunction} still functions\textit{. }The walls of a \textit{forcecage }can be damaged by spells as normal, except for \textit{disintegrate, }which automatically destroys it. The walls of a \textit{forcecage} can be damaged by weapons and supernatural abilities, but they have a Hardness of 30 and a number of hit points equal to 20 per caster level. Contact with a \textit{sphere of annihilation }or \textit{rod of cancellation} instantly destroys a \textit{forcecage.}\\
\textit{Barred Cage}: This version of the spell produces a 20-foot cube made of bands of force (similar to a \textit{wall of force }spell) for bars. The bands are a half-inch wide, with half-inch gaps between them. Any creature capable of passing through such a small space can escape; others are confined within the barred cage. You can't attack a creature in a barred cage with a weapon unless the weapon can fit between the gaps. Even against such weapons (including arrows and similar ranged attacks), a creature in the barred cage has cover. All spells and breath weapons can pass through the gaps in the bars.\\
\textit{Windowless Cell}: This version of the spell produces a 10-foot cube with no way in and no way out. Solid walls of force form its six sides.\\
