\spellentry{Form of the Dragon I}
\textbf{School }transmutation (polymorph); \textbf{Level }sorcerer/wizard 6\\
\textbf{Casting Time }1 standard action\\
\textbf{Components }V, S, M (a scale of the dragon type you plan to assume)\\
\textbf{Range }personal\\
\textbf{Target }you\\
\textbf{Duration }1 min./level (D)\\
\textbf{Save }see text; \textbf{Spell Resistance} no\\
You become a Medium chromatic or metallic dragon. You gain a +4 size bonus to Strength, a +2 size bonus to Constitution, a +4 natural armor bonus, fly 60 feet (poor), darkvision 60 feet, a breath weapon, and resistance to one element. You also gain one bite (1d8), two claws (1d6), and two wing attacks (1d4). Your breath weapon and resistance depend on the type of dragon. You can only use the breath weapon once per casting of this spell. All breath weapons deal 6d8 points of damage and allow a Reflex save for half damage. In addition, some of the dragon types grant additional abilities, as noted below.\\
\textit{Black dragon}: 60-foot line of acid, resist acid 20, swim 60 feet\\
\textit{Blue dragon}: 60-foot line of electricity, resist electricity 20, burrow 20 feet\\
\textit{Green dragon}: 30-foot cone of acid, resist acid 20, swim 40 feet\\
\textit{Red dragon}: 30-foot cone of fire, resist fire 30, vulnerability to cold\\
\textit{White dragon}: 30-foot cone of cold, resist cold 20, swim 60 feet, vulnerability to fire\\
\textit{Brass dragon}: 60-foot line of fire, resist fire 20, burrow 30 feet, vulnerability to cold\\
\textit{Bronze dragon}: 60-foot line of electricity, resist electricity 20, swim 60 feet\\
\textit{Copper dragon}: 60-foot line of acid, resist acid 20, \textit{spider climb} (always active)\\
\textit{Gold dragon}: 30-foot cone of fire, resist fire 20, swim 60 feet\\
\textit{Silver dragon}: 30-foot cone of cold, resist cold 30, vulnerability to fire\\
\textbf{Form of the Dragon II}\\
\textbf{School }transmutation (polymorph); \textbf{Level }sorcerer/wizard 7\\
This spell functions as \textit{form of the dragon I }except that it also allows you to assume the form of a Large chromatic or metallic dragon. You gain the following abilities: a +6 size bonus to Strength, a +4 size bonus to Constitution, a +6 natural armor bonus, fly 90 feet (poor), darkvision 60 feet, a breath weapon, DR 5/magic, and resistance to one element. You also gain one bite (2d6), two claws (1d8), two wing attacks (1d6), and one tail slap attack (1d8). You can only use the breath weapon twice per casting of this spell, and you must wait 1d4 rounds between uses. All breath weapons deal 8d8 points of damage and allow a Reflex save for half damage. Line breath weapons increase to 80-foot lines and cones increase to 40-foot cones.\\
\textbf{Form of the Dragon III}\\
\textbf{School }transmutation (polymorph); \textbf{Level }sorcerer/wizard 8\\
This spell functions as \textit{form of the dragon II }save that it also allows you to take the form of a Huge chromatic or metallic dragon. You gain the following abilities: a +10 size bonus to Strength, a +8 size bonus to Constitution, a +8 natural armor bonus, fly 120 feet (poor), blindsense 60 feet, darkvision 120 feet, a breath weapon, DR 10/magic, frightful presence (DC equal to the DC for this spell), and immunity to one element (of the same type \textit{form of the dragon I }grants resistance to). You also gain one bite (2d8), two claws (2d6), two wing attacks (1d8), and one tail slap attack (2d6). You can use the breath weapon as often as you like, but you must wait 1d4 rounds between uses. All breath weapons deal 12d8 points of damage and allow a Reflex save for half damage. Line breath weapons increase to 100-foot lines and cones increase to 50-foot cones.\\
