\spellentry{Freezing Sphere}
\textbf{School} evocation [cold]; \textbf{Level} sorcerer/wizard 6\\
\textbf{Casting Time} 1 standard action\\
\textbf{Components} V, S, F (a small crystal sphere)\\
\textbf{Range} long (400 ft. + 40 ft./level)\\
\textbf{Target, Effect, or Area} see text\\
\textbf{Duration} instantaneous or 1 round/level; see text\\
\textbf{Saving Throw} Reflex half; see text; \textbf{Spell Resistance} yes\\
\textit{Freezing sphere }creates a frigid globe of cold energy that streaks from your fingertips to the location you select, where it explodes in a 40-foot-radius burst, dealing 1d6 points of cold damage per caster level (maximum 15d6) to each creature in the area. A creature of the water subtype instead takes 1d8 points of cold damage per caster level (maximum 15d8) and is staggered for 1d4 rounds.\\
If the \textit{freezing sphere }strikes a body of water or a liquid that is principally water (not including water-based creatures), it freezes the liquid to a depth of 6 inches in a 40-foot radius. This ice lasts for 1 round per caster level. Creatures that were swimming on the surface of a targeted body of water become trapped in the ice. Attempting to break free is a full-round action. A trapped creature must make a DC 25 Strength check or a DC 25 Escape Artist check to do so.\\
You can refrain from firing the globe after completing the spell, if you wish. Treat this as a touch spell for which you are holding the charge. You can hold the charge for as long as 1 round per level, at the end of which time the \textit{freezing sphere }bursts centered on you (and you receive no saving throw to resist its effect). Firing the globe in a later round is a standard action.\\
