\spellentry{Geas, Lesser}
\textbf{School} enchantment (compulsion) [language-dependent, mind-affecting]; \textbf{Level} bard 3, sorcerer/wizard 4\\
\textbf{Casting Time} 1 round\\
\textbf{Components} V\\
\textbf{Range} close (25 ft. + 5 ft./2 levels)\\
\textbf{Target} one living creature with 7 HD or less\\
\textbf{Duration} 1 day/level or until discharged (D)\\
\textbf{Saving Throw }Will negates; \textbf{Spell Resistance} yes\\
A \textit{lesser geas }places a magical command on a creature to carry out some service or to refrain from some action or course of activity, as desired by you. The creature must have 7 or fewer HD and be able to understand you. While a \textit{geas }cannot compel a creature to kill itself or perform acts that would result in certain death, it can cause almost any other course of activity.\\
The \textit{geased }creature must follow the given instructions until the \textit{geas }is completed, no matter how long it takes.\\
If the instructions involve some open-ended task that the recipient cannot complete through his own actions, the spell remains in effect for a maximum of 1 day per caster level. A clever recipient can subvert some instructions.\\
If the subject is prevented from obeying the \textit{lesser geas }for 24 hours, it takes a –2 penalty to each of its ability scores. Each day, another –2 penalty accumulates, up to a total of –8. No ability score can be reduced to less than 1 by this effect. The ability score penalties are removed 24 hours after the subject resumes obeying the \textit{lesser geas.}\\
A \textit{lesser geas }(and all ability score penalties) can be ended by \textit{break enchantment}, \textit{limited wish}, \textit{remove curse}, \textit{miracle}, or \textit{wish}. \textit{Dispel magic }does not affect a \textit{lesser geas}.\\
\textbf{Geas/Quest}\\
\textbf{School }enchantment (compulsion) [language-dependent, mind-affecting]; \textbf{Level }bard 6, cleric 6, sorcerer/wizard 6\\
\textbf{Casting Time }10 minutes\\
\textbf{Target }one living creature\\
\textbf{Saving Throw} none; \textbf{Spell Resistance }yes\\
This spell functions similarly to \textit{lesser geas, }except that it affects a creature of any HD and allows no saving throw.\\
If the subject is prevented from obeying the \textit{geas/quest }for 24 hours, it takes a –3 penalty to each of its ability scores. Each day, another –3 penalty accumulates, up to a total of –12. No ability score can be reduced to less than 1 by this effect. The ability score penalties are removed 24 hours after the subject resumes obeying the \textit{geas/quest.}\\
A \textit{remove curse }spell ends a \textit{geas/quest }spell only if its caster level is at least two higher than your caster level. \textit{Break enchantment }does not end a \textit{geas/quest}, but \textit{limited wish}, \textit{miracle}, and \textit{wish }do.\\
Bards, sorcerers, and wizards usually refer to this spell as \textit{geas, }while clerics call the same spell \textit{quest}.\\
