\spellentry{Illusory Script}
\textbf{School} illusion (phantasm) [mind-affecting]; \textbf{Level} bard 3, sorcerer/wizard 3\\
\textbf{Casting Time} 1 minute per page\\
\textbf{Components} V, S, M (lead-based ink worth 50 gp)\\
\textbf{Range} touch\\
\textbf{Target} one touched object weighing no more than 10 lbs.\\
\textbf{Duration} one day/level (D)\\
\textbf{Saving Throw }Will negates; see text; \textbf{Spell Resistance} yes\\
You write instructions or other information on parchment, paper, or any suitable writing material. The \textit{illusory script }appears to be some form of foreign or magical writing. Only the person (or people) designated by you at the time of the casting can read the writing; it's unintelligible to any other character\textit{.}\\
Any unauthorized creature attempting to read the script triggers a potent illusory effect and must make a saving throw. A successful saving throw means the creature can look away with only a mild sense of disorientation. Failure means the creature is subject to a \textit{suggestion} implanted in the script by you at the time the \textit{illusory script} spell was cast. The \textit{suggestion} lasts only 30 minutes. Typical \textit{suggestions} include "Close the book and leave," "Forget the existence of this note," and so forth. If successfully dispelled by \textit{dispel magic, }the \textit{illusory script }and its secret message disappear. The hidden message can be read by a combination of the \textit{true seeing }spell with the \textit{read magic }or \textit{comprehend languages }spell.\\
