\spellentry{Phantom Steed}
\textbf{School} conjuration (creation); \textbf{Level} bard 3, sorcerer/wizard 3\\
\textbf{Casting Time} 10 minutes\\
\textbf{Components} V, S\\
\textbf{Range} 0 ft.\\
\textbf{Effect} one quasi-real, horselike creature\\
\textbf{Duration} 1 hour/level (D)\\
\textbf{Saving Throw} none; \textbf{Spell Resistance} no\\
You conjure a Large, quasi-real, horselike creature (the exact coloration can be customized as you wish). It can be ridden only by you or by the one person for whom you specifically created the mount. A phantom steed has a black head and body, gray mane and tail, and smoke-colored, insubstantial hooves that make no sound. It has what seems to be a saddle, bit, and bridle. It does not fight, but animals shun it and refuse to attack it.\\
The mount is AC 18 (--1 size, +4 natural armor, +5 Dex) and 7 hit points + 1 hit point per caster level. If it loses all its hit points, the phantom steed disappears. A phantom steed has a speed of 20 feet per two caster levels, to a maximum of 100 feet at 10th level. It can bear its rider's weight plus up to 10 pounds per caster level.\\
These mounts gain certain powers according to caster level. A mount's abilities include those of mounts of lower caster levels. \\
\textit{8th Level}: The mount can ride over sandy, muddy, or even swampy ground without difficulty or decrease in speed.\\
\textit{10th Level}: The mount can use \textit{water walk }at will (as the spell, no action required to activate this ability).\\
\textit{12th Level}: The mount can use \textit{air walk }at will (as the spell, no action required to activate this ability) for up to 1 round at a time, after which it falls to the ground.\\
\textit{14th Level}: The mount can fly at its speed with a bonus on Fly skill checks equal to your caster level.\\
