\spellentry{Pyrotechnics}
\textbf{School} transmutation; \textbf{Level} bard 2, sorcerer/wizard 2\\
\textbf{Casting Time} 1 standard action\\
\textbf{Components} V, S, M (one fire source)\\
\textbf{Range} long (400 ft. + 40 ft./level)\\
\textbf{Target} one fire source, up to a 20-ft. cube\\
\textbf{Duration} 1d4+1 rounds, or 1d4+1 rounds after creatures leave the smoke cloud; see text\\
\textbf{Saving Throw }Will negates or Fortitude negates; see text; \textbf{Spell Resistance} yes or no; see text\\
\textit{Pyrotechnics }turns a fire into a burst of blinding fireworks or a thick cloud of choking smoke, depending on your choice. The spell uses one fire source, which is immediately extinguished. A fire so large that it exceeds a 20-foot cube is only partly extinguished. Magical fires are not extinguished, although a fire-based creature used as a source takes 1 point of damage per caster level.\\
\textit{Fireworks}: The fireworks are a flashing, fiery, momentary burst of glowing, colored aerial lights. This effect causes creatures within 120 feet of the fire source to become blinded for 1d4+1 rounds (Will negates). These creatures must have line of sight to the fire to be affected. Spell resistance can prevent blindness.\\
\textit{Smoke Cloud}: A stream of smoke billows out from the fire, forming a choking cloud that spreads 20 feet in all directions and lasts for 1 round per caster level. All sight, even darkvision, is ineffective in or through the cloud. All within the cloud take –4 penalties to Strength and Dexterity (Fortitude negates). These effects last for 1d4+1 rounds after the cloud dissipates or after the creature leaves the area of the cloud. Spell resistance does not apply.\\
