\spellentry{Secret Page}
\textbf{School} transmutation; \textbf{Level} bard 3, sorcerer/wizard 3\\
\textbf{Casting Time} 10 minutes\\
\textbf{Components} V, S, M (powdered herring scales and a vial of will-o'-wisp essence)\\
\textbf{Range} touch\\
\textbf{Target} page touched, up to 3 sq. ft. in size\\
\textbf{Duration} permanent\\
\textbf{Saving Throw} none; \textbf{Spell Resistance} no\\
S\textit{ecret page }alters the contents of a page so that it appears to be something entirely different. The text of a spell can be changed to show another spell of equal or lower level known by the caster. This spell cannot be used to change a spell contained on a scroll, but it can be used to hide a scroll. \textit{Explosive runes }or \textit{sepia snake sigil }can be cast upon the \textit{secret page.}\\
A \textit{comprehend languages }spell alone cannot reveal a \textit{secret page}'s contents. You are able to reveal the original contents by speaking a special word. You can then peruse the actual page and return it to its \textit{secret page }form at will. You can also remove the spell by double repetition of the special word. A \textit{detect magic }spell reveals dim magic on the page in question but does not reveal its true contents. \textit{True seeing }reveals the presence of the hidden material but does not reveal the contents unless cast in combination with \textit{comprehend languages. }A \textit{secret page }spell can be dispelled, and the hidden writings can be destroyed by means of an \textit{erase }spell.\\
