\spellentry{Suggestion}
\textbf{School} enchantment (compulsion) [language-dependent, mind-affecting]; \textbf{Level} bard 2, sorcerer/wizard 3\\
\textbf{Casting Time} 1 standard action\\
\textbf{Components} V, M (a snake's tongue and a honeycomb)\\
\textbf{Range} close (25 ft. + 5 ft./2 levels)\\
\textbf{Target} one living creature\\
\textbf{Duration} 1 hour/level or until completed\\
\textbf{Saving Throw }Will negates; \textbf{Spell Resistance} yes\\
You influence the actions of the target creature by suggesting a course of activity (limited to a sentence or two). The \textit{suggestion }must be worded in such a manner as to make the activity sound reasonable. Asking the creature to do some obviously harmful act automatically negates the effect of the spell. \\
The suggested course of activity can continue for the entire duration. If the suggested activity can be completed in a shorter time, the spell ends when the subject finishes what it was asked to do. You can instead specify conditions that will trigger a special activity during the duration. If the condition is not met before the spell duration expires, the activity is not performed.\\
A very reasonable \textit{suggestion }causes the save to be made with a penalty (such as --1 or --2).\\
\textbf{Suggestion, Mass}\\
\textbf{School} enchantment (compulsion) [language-dependent, mind-affecting]; \textbf{Level} bard 5, sorcerer/wizard 6\\
\textbf{Range }medium (100 ft. + 10 ft./level)\\
\textbf{Targets} one creature/level, no two of which can be more than 30 ft. apart\\
This spell functions like \textit{suggestion}, except that it can affect more creatures. The same \textit{suggestion }applies to all these creatures.\\
