\spellentry{Transmute Rock to Mud}
\textbf{School} transmutation [earth]; \textbf{Level} druid 5, sorcerer/wizard 5\\
\textbf{Casting Time} 1 standard action\\
\textbf{Components} V, S, M/DF (clay and water)\\
\textbf{Range }medium (100 ft. + 10 ft./level)\\
\textbf{Area} up to two 10-ft. cubes/level (S)\\
\textbf{Duration} permanent; see text\\
\textbf{Saving Throw} see text; \textbf{Spell Resistance} no\\
This spell turns natural, uncut, or unworked rock of any sort into an equal volume of mud. Magical stone is not affected by the spell. The depth of the mud created cannot exceed 10 feet. A creature unable to levitate, fly, or otherwise free itself from the mud sinks until hip- or chest-deep, reducing its speed to 5 feet and causing a –2 penalty on attack rolls and AC. Brush or similar material thrown atop the mud can support creatures able to climb on top of it. Creatures large enough to walk on the bottom can wade through the area at a speed of 5 feet.\\
If \textit{transmute rock to mud }is cast upon the ceiling of a cavern or tunnel, the mud falls to the floor and spreads out in a pool at a depth of 5 feet. The falling mud and the ensuing cave-in deal 8d6 points of bludgeoning damage to anyone caught directly beneath the targeted area, or half damage to those who succeed on Reflex saves.\\
Castles and large stone buildings are generally immune to the effect of the spell, since \textit{transmute rock to mud }can't affect worked stone and doesn't reach deep enough to undermine such buildings' foundations. However, small buildings or structures often rest upon foundations shallow enough to be damaged or even partially toppled by this spell.\\
The mud remains until a successful \textit{dispel magic }or \textit{transmute mud to rock }spell restores its substance—but not necessarily its form. Evaporation turns the mud to normal dirt over a period of days. The exact time depends on exposure to the sun, wind, and normal drainage.\\
\textit{Transmute rock to mud }counters and dispels \textit{transmute mud to rock}.\\
